### 2020-02-20 Paper_SPHfP_feb2020_prb.tex
include manuscript and figures

%%   This file is part of the APS files in the REVTeX 4 distribution.
%%   Version 4.1r of REVTeX, August 2010
%%
%%
%%   Copyright (c) 2001, 2009, 2010 The American Physical Society.
%%
%
\documentclass[aps,prb,reprint,groupedaddress]{revtex4-1}

%\documentclass[aps,pra,twocolumn,groupedaddress]{revtex4-1}
%\documentclass[aps,prb,reprint,superscriptaddress]{revtex4-1}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsmath} 
\usepackage{graphicx}
\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

%\def\clau#1{\textcolor{ForestGreen}{#1}}
%\def\minor#1{\textcolor{green}{#1}}

\begin{document}

%Title of paper
\title{The role of $4f$ electrons in the stopping power of hafnium}
%Authors and affiliations

%---------------------------Argentina
\author{C. C. Montanari}
\email{mclaudia@iafe.uba.ar}
\affiliation{Instituto de
Astronom\'{\i}a y F\'{\i}sica del Espacio, Consejo Nacional de Investigaciones Cient\'{\i}ficas y
T\'{e}cnicas - Universidad de Buenos Aires, Pabell\'on IAFE, 1428 Buenos Aires, Argentina}%
%\affiliation{Universidad de Tres de Febrero, Sáenz Peña, Buenos Aires, Argentina.}
%------------------------------
\author{A. M. P. Mendez}
\affiliation{Instituto de
Astronom\'{\i}a y F\'{\i}sica del Espacio, Consejo Nacional de Investigaciones Cient\'{\i}ficas y
T\'{e}cnicas - Universidad de Buenos Aires, Pabell\'on IAFE, 1428 Buenos Aires, Argentina}%
%-------------------------------
\author{D. M. Mitnik}
\affiliation{Instituto de
Astronom\'{\i}a y F\'{\i}sica del Espacio, Consejo Nacional de Investigaciones Cient\'{\i}ficas y
T\'{e}cnicas - Universidad de Buenos Aires, Pabell\'on IAFE, 1428 Buenos Aires, Argentina}%
%\affiliation{Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales, Departamento de F\'{\i}sica, Ciudad Universitaria, 1428 Buenos Aires, Argentina.}
%----------------------------------------------
\author{J. E. Miraglia}
\affiliation{Instituto de
Astronom\'{\i}a y F\'{\i}sica del Espacio, Consejo Nacional de Investigaciones Cient\'{\i}ficas y
T\'{e}cnicas - Universidad de Buenos Aires, Pabell\'on IAFE, 1428 Buenos Aires, Argentina}%
%---------------------------Chile

\author{P.A. Miranda}
\affiliation{Departmento de F\'isica, Facultad de Ciencias Naturales, Matem\'atica y del Medio Ambiente. Universidad Tecnol\'ogica Metropolitana. 7800002, Chile}
%----------------------------
\author{R. Correa}
\affiliation{Departmento de F\'isica, Facultad de Ciencias Naturales, Matem\'atica y del Medio Ambiente. Universidad Tecnol\'ogica Metropolitana. 7800002, Chile}
\author{J. Wachter}
\affiliation{Departmento de F\'isica, Facultad de Ciencias Naturales, Matem\'atica y del Medio Ambiente. Universidad Tecnol\'ogica Metropolitana. 7800002, Chile}
\author{M. Aguilera}
\affiliation{Departmento de F\'isica, Facultad de Ciencias Naturales, Matem\'atica y del Medio Ambiente. Universidad Tecnol\'ogica Metropolitana. 7800002, Chile}
%-------------------------Lisboa
\author{E. Alves}
\affiliation{Centro de Ci\^{e}ncias e Tecnologias Nucleares, Instituto Superior T\'ecnico, Universidade de Lisboa,
2696-953 Sacav\'{e}m, Portugal}
\affiliation{Instituto de Plasma e Fus\~{a}o Nuclear, Instituto Superior T\'ecnico, Universidade de Lisboa, 2696-953 Sacav\'{e}m, Portugal}
%---------------------------
\author{N. Catarino}
\affiliation{Centro de Ci\^{e}ncias e Tecnologias Nucleares, Instituto Superior T\'ecnico, Universidade de Lisboa,
2696-953 Sacav\'{e}m, Portugal}
%---------------------------
\author{R.C. da Silva}
\affiliation{Centro de Ci\^{e}ncias e Tecnologias Nucleares, Instituto Superior T\'ecnico, Universidade de Lisboa,
2696-953 Sacav\'{e}m, Portugal}

\date{\today}

%=============================================================================================================================================================================================
%=============================================================================================================================================================================================
\begin{abstract}
 The stopping power of protons through Hf foil has been studied both experimentally and theoretically. The measurements were performed at the Laboratory of Accelerators and X-Ray Diffraction in Lisbon, by using the transmission method on self-supporting stopping material, for (0.6-2.5) MeV protons. The theoretical developments involved fully relativistic atomic structure calculations for %this Z=72 target 
 Hf, which required the solution of the Dirac equation. The shell-wise local plasma approximation (SLPA) was employed to describe the energy transferred to the bound 1s-4f electrons, and the outer four electrons were considered as a free electron gas (FEG). %We contemplated two different models for the FEG: the screened potential with cusp condition (SPCC) for energies below that of the plasmon excitation, and the dielectric formalism, for energies around the stopping maximum and above. 
 We found the relativistic description of the 4f-shell and the screening between $4f$ and $5p$ electrons are decisive around the stopping maximum. %for energies above 40 keV.
 Present theoretical and experimental results are in very good agreement in the energy region of the new measurements. However, our theoretical stopping cross sections show %interesting 
 substantial differences with the semi-empirical values by SRIM2013 and ICRU-49 at intermediate to low energies. Our calculations suggest the stopping maximum %around 40 eV$/(10^{15}$ atom/cm$^2)$ at 70 keV, which is
 higher and shifted to lower energies than previous predictions. Future measurements below 100 keV would be decisive for a better knowledge of the stopping maximum and below.
\end{abstract}

% insert suggested PACS numbers in braces on next line
\pacs{34.50.Bw} \keywords{Stopping power, Hafnium, Relativistic atomic structure, Transmission method}

\maketitle

%\linenumbers
%=============================================================================================================================================================================================
%=============================================================================================================================================================================================
\section{Introduction}
\label{intro}

For impact energies above a few keV/amu, mono-energetic charged particles penetrating a foil of any material lose their energy through a series of consecutive inelastic collisions, mainly with target electrons \cite{Chu01,Sigmund}. The information given by the energy loss process is essential not only to have a better knowledge of the physics behind the fundamental interactions, but also because it plays an important role in many applied fields such as materials science, nuclear physics, ionic implantation and radiotherapy \cite{Sigmund,Schardt}. Moreover, experimental values on ion mean energy loss per unit path $S(E)$ are of crucial relevance in order to check the reliability of semi-empirical models and to determine some key parameters 
%included in theoretical approaches of stopping power %\cite{Diwan,Damache01,Damache02,Damache03,Damache04}. 
\cite{Diwan,Damache04,Damache02}. However, very often, the experimental data available is rather scarce which is particularly troublesome when the material under study corresponds to an element of low occurrence on the Earth's upper crust, such as Hf.

So far, only one experimental work has been published regarding the stopping power cross section of pure hafnium for protons \cite{Sirotinin}, while more attention has been recently given to studies involving hafnium oxide due to its practical use \cite{Abril,Behar,Primetzhofer}. It is well known that significant attention has been paid in recent years to transition metaloxides such as HfO$_2$ for their potential as alternative gate dielectrics to replace SiO$_2$ for the future generation of nano-electronics with less than 45 nm gate length \cite{Choi,Robertson}. In this regard, some important physical properties of the above mentioned metal-oxide films depend on its thickness, which is often measured by using Rutherford Backscattering Spectrometry \cite{Alfassi01, Tesmer01}, a methodology that relies heavily on the determination of the stopping power of ion beams in the material of interest.

In this study we report stopping power cross section experimental values over the incident energy range (0.6-2.5) MeV for protons crossing self-supported Hf thin-film by using the transmission method. Our aim is not only to upgrade stopping power data compilations \cite{HPaul03,mondim17} but also to provide useful information about the processes governing the slowing down of protons in multielectronic targets. In the rare earth metals, the 4f electrons play a relevant role in the stopping power, being the first shell of bound electrons below the conduction band. As already noted \cite{Roth17}, the FEG shows unexpected %a peculiar
behaviour in these elements that states doubts on the proper description. In the case of Hf we found the contribution of the 4f-shell decisive even at impact energies around the stopping maximum, as we will show later in this work.

The theoretical approach used in this work requires 
the relativistic wave functions and binding energies of Hf, and considering 4 electrons per atom in the free electron gas (FEG) \cite{mendez2019}. The shell-wise local plasma approximation (SLPA) \cite{mon13} was employed to describe the energy transferred to the bound 1s-4f electrons, and two different models for the FEG: the screened potential with cusp condition model (SPCC) \cite{mon17}, which is non-linear binary formalism%,  for energies below that of the plasmon excitation
; and the Mermin-Lindhard dielectric formalism (ML) \cite{Mermin}, for energies around the stopping maximum and above.  On the other hand, Hf has the extra interest of the filled $4f$-subshell (with 14 electrons) as the main contribution below the FEG, causing the stopping cross sections to be very sensitive to a good description of this shell. The screening among the $4f$ and $5p$ electrons has been considered and found to play a major role within the SLPA calculations.

The experimental details and data are given in section \ref{experiment}, while the theoretical method is explained in section \ref{theory}. Present theoretical and experimental values are finally compared to the available experimental values \cite{Sirotinin},  the ICRU-49 calculations \cite{ICRU49} and the SRIM-2013 package \cite{Ziegler01}. Conclusions and discusions are given in section \ref{conclusion}.

%=============================================================================================================================================================================================

\begin{figure}[!t]
\centering
\includegraphics[width=9cm]{Fig01.eps}
\caption{RBS spectrum for $E_{avg} = 921,1$ keV protons on hafnium sample which is subsequently used to determine the energy loss in the foil.}
\label{F01}
\end{figure}

%=============================================================================================================================================================================================

\section{Experimental arrangements}
\label{experiment}
\subsection{Accelerator and scattering Chamber}
The procedure used in this work to obtain stopping power data is essentially the same as described in reference \cite{Miranda01}. However, this time the measurements were made at the IST/LATR
(Laboratory of Accelerators and X-Ray Diffraction) in Lisbon. This facility uses a 2.5 MV Van de Graaff accelerator to deliver $^1$H$^+$ primary ion beams through a series of electrostatic lenses and collimators onto a thin Au/SiO$_2$ sample which is used as a scattering center. This sample was placed in the center of a RBS/C scattering chamber where a high vacuum (pressure of
$\sim$10$^{-6}$ Torr) was maintained during the measurements. The beam current on sample was kept at around 5.0 nA in order to attain sufficient statistics in each particle spectrum.
By using a beam spot of about 1.0 mm in diameter, a solid angle of 11.4 msr was attained. The overall energy resolution (FWHM) of the detection system was about 15 keV relative to 5.486 MeV alpha particles from a $^{241}$Am source.

%---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Target}
The stopping material under analysis was a hafnium foil with nominal thickness of 1.0 $\mu$m and 99.95\% purity which was supplied by Lebow Company \cite{Lebow}. However, a more precise thickness value was achieved by measuring the energy loss of alpha particles coming from a calibrated ($^{239}$Pu, $^{241}$Am, $^{244}$Cm) source. From the alpha spectra with and without the Hf foil interposed, the characteristic energy shift $\delta$E was measured and then combined with the stopping power for 5.486 MeV alphas on hafnium (55.69 eV/$10^{15}$ at/cm$^2$) found in Ref.\cite{Ziegler01} in order to obtain an areal density of (4,13$\pm$0.21)$\times 10^{19}$ at/cm$^2$ which corresponds to a thickness of 0.920$\pm$0.046 $\mu$m.
%---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Energy loss measurement}
Once the beam impinges on the Au/SiO$_2$ sample, protons are backscattered towards a Si surface barrier detector located at 140$^{\circ}$ relative to the initial beam direction. Fig.\ref{F01} shows two particle spectra where ion energies $E_2$ and $E_1$ are associated to a placed and removed hafnium sample, respectively. Both energy distributions were fitted by gaussian functions to obtain the mean energy and width (FWHM) of the peaks \cite{Sun01}, and from the difference between these two peak positions in the spectrum, the total energy loss $\Delta E = (E_1 - E_2)$ in the foil was calculated. As established in previous studies \cite{Miranda01, Damache02}, the experimental stopping powers cross sections $\varepsilon (E) $ are determined at some mean energy $E_{avg}$ by measuring the ion energy losses $\Delta E$ within the investigated Hf foil which has a mean thickness denoted by $\Delta x$. In this way, only when the energy loss fraction $\Delta E/E_{avg}$ across the Hf foil is not exceeding 20\%
it is possible to define the stopping cross section by \cite{Raisanen01,Schulz01}:
\begin{equation}\label{eq:stcross}
 \varepsilon (E) = \frac{S(E)}{N} = -\frac{dE}{N dx} \approx -\frac{\Delta E}{N \Delta x},
\end{equation}
where $N$ denotes the atomic number density (atoms cm$^{-3}$) of the material under study. When this condition was not fulfilled, a small correction to the mean energies $E_{avg}$ was applied in order to account for the nonlinear dependence on ion energy of stopping powers \cite{Chilton,Rajatora}.

\section{Theoretical Method} \label{theory}
The energy loss of ions in metal targets responds to different physical mechanisms, depending on the impact ion velocity. At low velocities the binary collisions are responsible of the loss of energy by the ion. The main contribution is the ionization of electrons of the metal conduction band, which is well approximated by a free electron gas (FEG) of Fermi velocity $v_F$. Instead, above certain velocity (i.e. $v\geq 1.5 \ v_F$) , not only binary but also collective excitations (plasmons) are possible \cite{mon17}. At high energies, not only the FEG contributes to the stopping power but also the bound electrons. The method used in this work combines a description for the interaction with the valence (or conduction) electrons as a FEG, and a different one for the interaction with the bound electrons.

We used the SPCC model \cite{mon17} to describe the stopping power of low velocity charged particles in the FEG. It is a non-perturbative binary collisional model, thus valid at energies below that of plasmon excitations.  The SPCC \cite{mon17} is based on a screened central potential with cusp condition of the electronic density close to the projectile. It proved to give a good description of the induced electron density even for negative projectiles \cite{mon17}, and reproduces the low velocity proton-antiproton differences in the stopping power (Barkas effect). The SPCC formalism only depends on the Wigner-Seitz radio, $r_S$, which is a measure of the electronic density of the FEG. For metals of well-known value of $r_S$, the SPCC describes very well the low energy experimental stopping data \cite{mon17}, agreeing with the DFT results by Echenique and coworkers \cite{eche81,nagy89} at $v=0$.

%In the case of 
Hafnium (Z=72, [Xe] $4f^{14} \ 6s^2 \ 5d^2$), belongs to the first groups of transition metals, with  4 electrons as FEG (theoretical $r_S=2.14$ a.u.) and $1s-4f$ electrons bound. We checked the $r_S$ value with the experimental one obtained from measured energy loss function by Lynch \textit{et al} \cite{lynch75}. The experimental plasmon energy of Hf is $\hbar\omega_P \approx 15.8$ eV, with a width at half maximum $\delta \approx 4.4$ eV, and $r_S \approx 2.07$ a.u. \cite{lynch75}. Following \cite{mon17}, the difference of less than $5\%$ between theoretical and experimental $r_S$ assess Hf as a canonical target.

Above certain impact velocity the plasmon contribution is important (i.e. around and above the stopping maximum). An interesting value for our analysis is the minimum impact velocity to excite
plasmons, $v_P$. In the dielectric formalism this value can be obtained as $v_P\ \approx \ v_F[1+(3 \pi\ v_F)^{-1/2} ]$ \cite{suppression}. To describe the energy loss considering collective and binary excitation we resort to the ML dielectric formalism \cite{Mermin}, which is a linear response, perturbative approximation, so it depends on the square of the ion charge. In this formalism, the response of target electrons to the ion passage is described through the quantum dielectric function, depending on the characteristic $r_S$ and $\delta$ parameters of the FEG.

%----------------------------------------------------------------------
\begin{figure*}[!t]
\centering
\includegraphics[width=11.cm]{Hf_bindener_bar.eps}
% \includegraphics[width=13.cm]{Hf_bindener_ratio.eps}
\caption{(a) Binding energies of Hf. Non relativistic and relativistic calculations are given with filled symbols.  Experimental measurements for solids\cite{williams1995} are depicted with hollow circles. (b) Corresponding relative errors respect to experimental data.}
\label{Binding_E}
\end{figure*}
%------------------------------------------------------------------------

For the stopping power due to bound electrons, the SLPA \cite{mon17,mon13} is employed. It is worth to mention that the only inputs for the SLPA are the space-dependent densities of each shell in the ground state, and their binding energies. Collective processes and screening among electrons are included. Hafnium is a relativistic target, therefore the wave functions and binding energies must be obtained as solutions of Dirac equation  \cite{mendez2019}.

Figure \ref{Binding_E} (a) shows non relativistic and our relativistic binding energies $E_{nl\pm}$, where $\pm=j\pm1/2$.
We compare our results with experimental measurements on solid Hf \cite{williams1995} from $1s$ to $4f_{\pm}$. We notice that not only the deep shells require relativistic calculations, but also the outer $5p$ and $4f$ shells. Figure \ref{Binding_E} (b) displays the relative error of non relativistic and relativistic energies with respect to the experimental data in logarithmic scale. This figure shows very clearly the disability of non-relativistic calculations to describe the experimental data, which surprisingly worsens from the inner to the outer shells. 
%It is clear in this figure the disability of non-relativistic binding energies to describe the experimental data, which surprisingly worsens from the inner to the outer-shells.
%We used here the theoretical binding energies and wave functions of Hf calculated in \cite{mendez2019}. 
%The present theoretical description of Hf considers a FEG of $r_S=2.14$ and 1s-4f electrons as bound.

%The relativistic results for binding energies and wave functions in \cite{mendez2019} 
Our relativistic binding energies present spin-orbit split. However, in total stopping power where the initial state of the excited electron is not measured, the quantum uncertainty in energy $\Delta E$ melts this split. The criteria $\Delta E\Delta t \geq \hbar /2$ merges the energies  $E_{nl+} - E_{nl-}$ for sufficiently small values of $\Delta t$ (the collisional mean-time). In fact, at sufficiently high impact velocity we can expect all target electrons to respond together to the ion passage \cite{lindhard53,chu72}. As in \cite{mon09}, the collisional time was estimated as follows $\Delta t \approx \langle r_i\rangle/v$, with $\langle r_i\rangle$ and $v$ being the orbital mean radio and impact velocity, respectively.

%--------------------------------------------------------------------------------
\begin{figure*}[!t]
\centering
\includegraphics[width=13.cm]{Fig02.eps}
\caption{Theoretical total stopping cross sections of protons in afnium adding FEG and bound $1s$-$4f$ contributions, as mentioned in the inset. The vertical grey dashed-line indicates the energy above which plasmon excitation is possible. %The FEG calculations employ the SPCC model at low energies (no plasmon excitation) and the ML formalism (includes plasmon excitation). 
The SLPA calculations for $1s$-$4f$ bound electrons with and without $5p-4f$ screening give different total stopping values, displayed with solid and dotted lines, respectively.}
\label{slpa4f}
\end{figure*}
%========================================================================================

The SLPA calculates the contribution of each subshell of bound electrons to the total stopping cross sections. We found that for every sub-shell of Hf, at the impact energy at which this sub-shell began to contribute, the spin-orbit split was unresolved.  Then the $nl$-electrons should be considered together, responding to the ion passage as a single gas of electrons with density $\delta_{nl}(r)$ and a mean binding energy $E_{nl}$. This is important within the SLPA calculations because of the screening among electrons of the \textit{same} binding energy. For example, the $4f_{-}$ and $4f_{+}$ of Hf can only be resolved for impact energies $E<0.05$ keV, but the contribution of $4f$ to the total stopping is negligible if $E<40$ keV. Moreover, the $5p$ and $4f$ electrons of Hf are very close in energy ( $\Delta E_{5p-4f} \approx 1 a.u.$ \cite{mendez2019}) %merges them 
and react together for impact energies $E>40$ keV (inter-shell screening).

%================================================================================
\begin{figure*}[!t]
\centering
\includegraphics[width=13.0cm]{Fig03.eps}
\caption{Stopping power cross section of hafnium for protons. Symbols: solid circles, present values; open circles, previous data \cite{Sirotinin}. Curves: Black solid-line, present full theoretical results with $4f$-$5p$ screening; violet dashed-line, values from ICRU-49 \cite{ICRU49}, and green dotted line, SRIM-2013 \cite{Ziegler01}.}
\label{F03}
\end{figure*}
%=============================================================================================================================================================================================

In figure \ref{slpa4f} we display the present theoretical stopping cross section of Hf for protons with and without the $5p-4f$ screening. We added the FEG and the bound electron contributions as explained above. The minimum energy for plasmon excitation was estimated as $37$ keV. We used the non-perturbative SPCC model for impact energies $E \leq 37$ keV, and the perturbative ML calculation above this energy. %The range of validity of the perturbative model below 100 keV is arguable, however the asymetric $Z_{ion}/Z_{target}$ relation is clear. 
%The importance of the 4f-5p screening is signed with arrows in this figure. 
Clearly, considering $5p-4f$ electrons as a single group of 20 electrons with screening among them gives lower stopping values than the addition of the separate $5p$ and $4f$ contributions. This is a shell correction that can only be considered within a many electron model such as the SLPA.

%=============================================================================================================================================================================================
\section{Analysis of the results and discussion}
\label{discussion}

The present data and theoretical results are displayed in table \ref{table01}, and figure \ref{F03}. As can be seen in Table~\ref{table01}, an overall relative uncertainty of around 5$\%$ was achieved for the experimental stopping power values, mainly due to the uncertainty in the hafnium foil thickness. %A good agreement in the whole energy range is found between our new data and the previous measurements made by Sirotinin \textit{et al.} \cite{Sirotinin}, as depicted in Fig.~\ref{F03}. Also shown in the same figure are our theoretical stopping cross sections together with the semi-empirical SRIM-2013 curve \cite{Ziegler01} and the suggested values by the ICRU-49 report \cite{ICRU49}.

%-----------------TABLE--------------------------------------------------------------------------
\begin{table*}[!t]
\centering
\caption{Stopping power values S$_{exp}$ of hafnium for protons measured in this work. $\Delta$E/E values are also shown.
\label{table01}}

\vspace{0.2cm}

\begin{ruledtabular}
\begin{tabular}{ccc|ccc|ccc} %\hline\hline
E$_{avg}$	&S$_{exp}$			&$\Delta$E/E	&E$_{avg}$	&S$_{exp}$			&$\Delta$E/E	&E$_{avg}$	&S$_{exp}$			&$\Delta$E/E\\
keV		&eV/(10$^{15}$ at/cm$^2$)	&$\%$		&keV		&eV/(10$^{15}$ at/cm$^2$)	&$\%$		&keV		&eV/(10$^{15}$ at/cm$^2$)	&$\%$\\ \hline
516,6	&	25,8	$\pm$	1,3	&	20,5	&	1170,3	&	18,25	$\pm$	0,91	&	6,4	&	1813,4	&	15,10	$\pm$	0,76	&	3,4	\\
567,8	&	24,8	$\pm$	1,2	&	17,9	&	1220,0	&	18,08	$\pm$	0,90	&	6,1	&	1862,7	&	14,79	$\pm$	0,74	&	3,3	\\
618,8	&	23,9	$\pm$	1,2	&	15,8	&	1269,6	&	17,57	$\pm$	0,88	&	5,7	&	1912,0	&	14,21	$\pm$	0,71	&	3,0	\\
669,6	&	23,2	$\pm$	1,2	&	14,2	&	1319,2	&	17,32	$\pm$	0,87	&	5,4	&	1961,2	&	14,46	$\pm$	0,72	&	3,0	\\
720,1	&	22,5	$\pm$	1,1	&	12,8	&	1368,8	&	17,15	$\pm$	0,86	&	5,1	&	2010,4	&	14,34	$\pm$	0,72	&	2,9	\\
770,5	&	21,8	$\pm$	1,1	&	11,6	&	1418,3	&	16,69	$\pm$	0,83	&	4,8	&	2059,6	&	13,76	$\pm$	0,69	&	2,7	\\
820,8	&	21,3	$\pm$	1,1	&	10,7	&	1467,8	&	16,43	$\pm$	0,82	&	4,6	&	2108,8	&	13,78	$\pm$	0,69	&	2,7	\\
871,0	&	20,8	$\pm$	1,0	&	9,8	&	1517,2	&	16,13	$\pm$	0,81	&	4,4	&	2158,0	&	13,70	$\pm$	0,69	&	2,6	\\
921,1	&	20,3	$\pm$	1,0	&	9,1	&	1566,7	&	16,04	$\pm$	0,80	&	4,2	&	2206,5	&	13,33	$\pm$	0,67	&	2,5	\\
971,1	&	19,9	$\pm$	1,0	&	8,4	&	1616,0	&	15,77	$\pm$	0,79	&	4,0	&	2256,4	&	13,27	$\pm$	0,66	&	2,4	\\
1021,0	&	19,33	$\pm$	0,97	&	7,8	&	1665,4	&	15,51	$\pm$	0,78	&	3,8	&	2305,5	&	13,07	$\pm$	0,65	&	2,3	\\
1070,8	&	19,03	$\pm$	0,95	&	7,3	&	1714,8	&	15,46	$\pm$	0,77	&	3,7	&	2354,7	&	12,91	$\pm$	0,65	&	2,2	\\
1120,6	&	18,73	$\pm$	0,94	&	6,9	&	1764,1	&	14,93	$\pm$	0,75	&	3,5	&	2403,8	&	12,61	$\pm$	0,63	&	2,2	\\ \\ %\hline
\end{tabular}
\end{ruledtabular}
\end{table*}

There is good agreement between the present theoretical results, the new data in Table~\ref{table01}, and also with previous data by Sirotinin \cite{Sirotinin}, except for the lowest energy measurement at 80 keV. It is interesting that our full theoretical curve differs from SRIM-2013 one for impact energies below 100 keV. We obtain a stopping maximum around 40 $10^{-15}$ eV cm$^2$/atom  at 65 keV. Instead, following the up-to-now only set of data \cite{Sirotinin}, SRIM-2013 suggests a lower stopping maximum %of 37.4 $10^{-15}$ eV cm$^2$/atom 
at impact energy of 115 kev. It is worth to mention that similar theoretical results using the experimental $r_S=2.07$ a.u. instead of the theoretical $r_S=2.14$ a.u. give the stopping maximum at the same impact energy but $4\%$ higher. Future experiments for
impact energies below 100 keV would be important to complete this study.

%=============================================================================================================================================================================================
%=============================================================================================================================================================================================
\section{Conclusion}
\label{conclusion}
In this work, we have used the transmission method to experimentally determine stopping power cross section values for (0.6-2.5) MeV protons incident on self-supporting Hf foils with an overall uncertainty of around 5\%. Additionally, we calculated values extracted from the theoretical framework that involved the relativistic wave functions and binding energies of Hf, and considered 4 electrons per atom in the free electron gas. The shell-wise local plasma approximation was employed to describe the energy transferred to the bound 1s-4f electrons, and two different models for the FEG: the screened potential with cusp condition (SPCC model) for energies below that of the plasmon excitation, and the Mermin-Lindhard dielectric formalism, for energies around the stopping maximum and above. Present theoretical stopping cross sections cover an extensive energy range from 1 keV/amu-10 meV/amu.

%Our results have been confronted to both experimental values found in the literature and semi-empirical results. 
At high impact energies, the new stopping  measurements are in good agreement with our theoretical results, and also with already published experimental data and semi-empirical values calculated by means of SRIM-2013 and ICRU-49.  However, we call the attention that around the stopping maximum and at lower impact energies the difference between our full-theoretical results and SRIM is substantial. To our knowledge, these are the first theoretical calculations of stopping in Hf taking into account the atomic relativistic effect.
%for energies around and below the stopping maximum, the present full-theoretical results clearly differ from semiempirical SRIM values, which are based on previous data. 
Future experiments for impact energies below 100 keV would be important to complete this study.
%=======================================================================================================================================
\begin{acknowledgments}
This work was funded by VRAC Grant Number L1-17 of Universidad Tecnol\'ogica Metropolitana, Chile; also by the Consejo Nacional de Investigaciones Científicas y Técnicas (CONICET), the Agencia Nacional de Promoción Científica y Tecnológica (ANPCyT), and Universidad de Buenos Aires (UBA), of Argentina.

The authors gratefully acknowledge the invaluable contribution of F. Baptista from ITN/IST-UTL, Sacav\'{e}m, Portugal for his constant availability. 
\end{acknowledgments}

%=======================================================================================================================================
%  references
\begin{thebibliography}{00}
%============================================================================================================================
%Fundamentals of Stopping Power and Energy Loss processes
\bibitem{Chu01} W.K. Chu, J.W. Mayer, M.A. Nicolet (Eds.) Backscattering Spectrometry, Academic Press, New York, 1978.
\bibitem{Sigmund} P. Sigmund, ''Particle Penetration and Radiation Effects. General Aspects and Stopping of Swift Point Charges``.
Springer Series in Solid-State Sciences, Vol. 151, Springer, Berlin (2006).
\bibitem{Schardt} D. Schardt, T. Els\"asser, D. Schulz-Ertner, Rev. Mod. Phys. \textbf{82},  383-425 (2010).
\bibitem{Diwan} P.K. Diwan, S. Kumar, Nucl. Instr. and Meth. B \textbf{359}, 78-84 (2015).
%\bibitem{Damache01} D. Moussa, S. Damache and S. Ouichaoui, Nucl. Instr. and Meth. \textbf{B} 343,  44-47 (2015).
\bibitem{Damache02} D. Moussa, S. Damache and S. Ouichaoui, Nucl. Instr. and Meth. B \textbf{268}, 1754-1758 (2010); \textbf{343},  44-47 (2015).
%\bibitem{Damache03} S. Damache, S. Ouichaoui, D. Moussa and A. Dib, Nucl. Instr. and Meth. \textbf{B} 249,  22-25 (2006).
\bibitem{Damache04} S. Damache, S. Ouichaoui, A. Belhout, A. Medouni and I. Toumert, Nucl. Instr. and Meth. B \textbf{225}, 449-463 (2004).
%============================================================================================================================
%% Use of Hafnium and Stopping Power of Hafnium
\bibitem{Sirotinin} E.I. Sirotinin, A.F. Tulinov, V.A. Khodyrev, V.N. Mizgulin, Nucl. Instr. and Meth. B \textbf{4}, 337-345 (1984).
\bibitem{Abril} I. Abril, M. Behar, R. Garc\'a Molina, R.C. Fadanelli, L.C.C.M. Nagamine, P.L. Grande, L. Sch\"unemann, C.D. Denton, N.R. Arista, E.B. Saitovich,
Eur. Phys. J. D \text{54}, 65-70 (2009).
\bibitem{Behar} M. Behar, R.C. Fadanelli, I. Abril, R. García-Molina, C.D. Denton, L.C.C.M. Nagamine, N.R. Arista, Phys. Rev. A \textbf{80},  062901 (2009).
\bibitem{Primetzhofer} D. Primetzhofer, Nucl. Instr. and Meth. B \textbf{320}, 100-103 (2014) .
%============================================================================================================================
% Why Hafnium is so important:
\bibitem{Choi} J.H. Choi, Y. Mao, J.P. Chang, Mat. Sci. Eng. R \textbf{72}, 97-136 (2011) .
\bibitem{Robertson} J. Robertson, R.M. Wallace, Mat. Sci. Eng. R \textbf{88}, 1-41 (2015) .
%============================================================================================================================
%Importance of Stopping Power for Ion Beam Analysis
\bibitem{Alfassi01} Z.B. Alfassi, "Non-destructive elemental analysis", Blackwell Science Ltd. (2001).
\bibitem{Tesmer01} J.R. Tesmer, M. Nastasi, J.C. Barbour, C.J. Maggiore and J.W. Mayer, ''Handbook of Modern Ion Beam Material Analysis``, Materials Research Society (1995).
\bibitem{HPaul03} https://www-nds.iaea.org/stopping/.
\bibitem{mondim17} C.C. Montanari, and P. Dimitriou, Nucl. Instr. and Meth. B \textbf{408},  50-55 (2017).
\bibitem{Roth17} D. Roth, B. Bruckner, M. V. Moro, S. Gruber, D. Goebl, J. I. Juaristi, M. Alducin,R. Steinberger, J. Duchoslav, D. Primetzhofer, and P. Bauer1, Phys. Rev. Lett \textbf{118}, 103401 (2017)

%============================================================================================================================
%Semi-empirical and Theoretical approach

\bibitem{mendez2019} A.M.P. Mendez, C.C. Montanari, D.M. Mitnik, %\textit{Relativistic atomic structure calculations of heavy targets for inelastic collisions}, 
Nucl. Instrum. Meth. B \textbf{460}, 114-118 (2019) .
\bibitem{williams1995}
G. Williams in http://xdb.lbl.gov/Section1/Sec\_1-1.html
\bibitem{mon13} C.C. Montanari and J.E. Miraglia, Adv. Quant. Chem. \textbf{65}, edited by Dz. Belkic (Elsevier, Amsterdam, 2013), Chap. 7, pp. 165-201.
\bibitem{mon17} C.C. Montanari, and J.E. Miraglia, Phys. Rev. A \textbf{96}, 012707 (2017).
\bibitem{Mermin} N.D. Mermin, Phys. Rev. B \textbf{1}, 2362 (1970).
\bibitem{ICRU49} ICRU report 49, Stopping Powers and Ranges for Protons and Alpha Particles, International Commission on
Radiation Units and Measurements, 1993.
\bibitem{Ziegler01} J.F. Ziegler, J.P. Biersack, SRIM2013, Computer Program and Manual. Available from www.srim.org.
%============================================================================================================================
% Experimental
\bibitem{Miranda01} P.A. Miranda, A. Sep\'ulveda, J.R. Morales, T. Rodriguez, E. Burgos, H. Fern\'andez, Nucl. Instr. and Meth. B \textbf{318}, 292-296  (2014).
\bibitem{Lebow} Lebow Company. 5960 Mandarin Ave. Goleta CA, 93117, USA.

%============================================================================================================================
% Transmission Method
\bibitem{Sun01} G. Sun, M. D\"{o}belli, A.M. M\"{u}ller, M. Stocker, M. Suter and L. Wacker, Nucl. Instr. and Meth. B \textbf{256}, 586-590 (2007).
\bibitem{Raisanen01} J. Raisanen, U. Watjen, A.J.M. Plompen, F. Munnik, Nucl. Instr. and Meth. \textbf{B} 118 (1996) 1-6.
\bibitem{Schulz01} F. Schulz and J. Shchuchinzky, Nucl. Instr. and Meth. B \textbf{12},  90-94 (1985).
\bibitem{Chilton} A.B. Chilton, J.N. Cooper and J.C. Harris, Phys. Rev. \textbf{93}, 413-418  (1954).
\bibitem{Rajatora} M. Rajatora, K. Vakevainen, T. Ahlgre, E. Rauhala, J. Raisanen, K. Rakennus, Nucl. Instr. and Meth. B \textbf{119}, 457-462 (1996).
%============================================================================================================================
%X Theoretical Approach (Montanari)
\bibitem{eche81} P. M. Echenique, R. M. Nieminen and R. H. Ritchie, Sol. State Comm. \textbf{37}, 779-781 (1981).
\bibitem{nagy89} I. Nagy, A. Arnau, P. M. Echenique, and E. Zaremba, Phys. Rev. B \textbf{40}, 11986 (1989).
% \bibitem{Vs94} J. E. Valdes, J. C. Eckardt, G. H. Lantschner, and N. R. Arista, Phys. Rev. A \textbf{49}, 1083 (1994).
\bibitem{lynch75} D. W. Lynch, C. G. Olson, and J. H. Weaver, Phys. Rev. B \textbf{11}, 3617-3624 (1975).
\bibitem{suppression} C. C. Montanari, J. E. Miraglia, and N. R. Arista, Phys. Rev. A \textbf{62}, 052902 (2000).
\bibitem{lindhard53} J. Lindhard, M. Scharff,  Mat. Fys. Medd. Dan. Vid. Selsk  \textbf{27}, 1 (1953).
\bibitem{chu72} W. K. Chu, D. Powers, Rev. Lett. A \textbf{40}, 23 (1972).
\bibitem{mon09} C.C. Montanari, D.M. Mitnik, C.D. Archubi, and J.E. Miraglia, Phys. Rev. A \textbf{80}, 012901 (2009).

\end{thebibliography}

\end{document}

### 2020-02-21 hafnio_ale.tex
intro and experiment revised

--- 2020-02-20 Paper_SPHfP_feb2020_prb.tex
@@
 The measurements were performed at the Laboratory of Accelerators and X-Ray Diffraction in Lisbon, by using the transmission method on self-supporting stopping material, for (0.6-2.5) MeV protons.
-The theoretical developments involved fully relativistic atomic structure calculations for %this Z=72 target
- Hf, which required the solution of the Dirac equation.
-The shell-wise local plasma approximation (SLPA) was employed to describe the energy transferred to the bound 1s-4f electrons, and the outer four electrons were considered as a free electron gas (FEG).
-%We contemplated two different models for the FEG: the screened potential with cusp condition (SPCC) for energies below that of the plasmon excitation, and the dielectric formalism, for energies around the stopping maximum and above.
- We found the relativistic description of the 4f-shell and the screening between $4f$ and $5p$ electrons are decisive around the stopping maximum. %for energies above 40 keV.
- Present theoretical and experimental results are in very good agreement in the energy region of the new measurements. However, our theoretical stopping cross sections show %interesting 
- substantial differences with the semi-empirical values by SRIM2013 and ICRU-49 at intermediate to low energies.
-Our calculations suggest the stopping maximum %around 40 eV$/(10^{15}$ atom/cm$^2)$ at 70 keV, which is
- higher and shifted to lower energies than previous predictions. Future measurements below 100 keV would be decisive for a better knowledge of the stopping maximum and below.
+The theoretical developments involved fully relativistic atomic structure calculations for Hf, which required the solution of the Dirac equation. 
+The shell-wise local plasma approximation (SLPA) was used to describe the energy transferred to the bound 1s-4f electrons, and the outer four electrons were considered as a free electron gas (FEG). 
+ We found the relativistic description of the 4f-shell, and the screening between $4f$ and $5p$ electrons are decisive around the stopping maximum. 
+The present theoretical and experimental results have very good agreement in the energy region of the new measurements. 
+However, our theoretical stopping cross sections show substantial differences with the semi-empirical models, such as SRIM2013 and ICRU-49, at intermediate to low energies. 
+Our calculations suggest the stopping maximum to be larger, and shifted to lower energies, respect to these semi-empirical models.
+Future measurements below 100 keV would be required to validate our predictions.
 \end{abstract}
@@
 For impact energies above a few keV/amu, mono-energetic charged particles penetrating a foil of any material lose their energy through a series of consecutive inelastic collisions, mainly with target electrons \cite{Chu01,Sigmund}.
-The information given by the energy loss process is essential not only to have a better knowledge of the physics behind the fundamental interactions, but also because it plays an important role in many applied fields such as materials science, nuclear physics, ionic implantation and radiotherapy \cite{Sigmund,Schardt}.
-Moreover, experimental values on ion mean energy loss per unit path $S(E)$ are of crucial relevance in order to check the reliability of semi-empirical models and to determine some key parameters
-%included in theoretical approaches of stopping power %\cite{Diwan,Damache01,Damache02,Damache03,Damache04}. 
-\cite{Diwan,Damache04,Damache02}.
-However, very often, the experimental data available is rather scarce which is particularly troublesome when the material under study corresponds to an element of low occurrence on the Earth's upper crust, such as Hf.
+The information given by the energy loss process is essential not only to have a better knowledge of the phenomena behind the fundamental interactions but also because it plays an important role in many applied fields such as materials science, nuclear physics, ionic implantation, and radiotherapy \cite{Sigmund,Schardt}.
+Experimental data in ion mean energy loss per unit path $S(E)$ is of crucial relevance to check the reliability of semi-empirical models and to determine some key parameters \cite{Diwan,Damache04,Damache02}.
+However, often, the experimental data available is rather scarce, which is troublesome when the material under study corresponds to an element of low occurrence on the Earth's upper crust, such as Hf.
 
-So far, only one experimental work has been published regarding the stopping power cross section of pure hafnium for protons \cite{Sirotinin}, while more attention has been recently given to studies involving hafnium oxide due to its practical use \cite{Abril,Behar,Primetzhofer}.
-It is well known that significant attention has been paid in recent years to transition metaloxides such as HfO$_2$ for their potential as alternative gate dielectrics to replace SiO$_2$ for the future generation of nano-electronics with less than 45 nm gate length \cite{Choi,Robertson}.
-In this regard, some important physical properties of the above mentioned metal-oxide films depend on its thickness, which is often measured by using Rutherford Backscattering Spectrometry \cite{Alfassi01, Tesmer01}, a methodology that relies heavily on the determination of the stopping power of ion beams in the material of interest.
+So far, only one experimental work has been published regarding the stopping power cross section of pure hafnium for protons \cite{Sirotinin}, while more attention has been recently given to studies involving hafnium oxide because of its practical use \cite{Abril,Behar,Primetzhofer}.
+It is well known that significant attention has been paid in recent years to transition metal-oxides such as HfO$_2$ because of their potential as alternative gate dielectrics to replace SiO$_2$ for the future generation of nano-electronics with less than 45 nm gate length \cite{Choi,Robertson}.
+Some important physical properties of the above mentioned metal-oxide films depend on its thickness, which is often measured by using Rutherford Backscattering Spectrometry \cite{Alfassi01, Tesmer01}, a method that relies heavily on the determination of the stopping power of ion beams in the material of interest.
 
-In this study we report stopping power cross section experimental values over the incident energy range (0.6-2.5) MeV for protons crossing self-supported Hf thin-film by using the transmission method.
-Our aim is not only to upgrade stopping power data compilations \cite{HPaul03,mondim17} but also to provide useful information about the processes governing the slowing down of protons in multielectronic targets.
-In the rare earth metals, the 4f electrons play a relevant role in the stopping power, being the first shell of bound electrons below the conduction band.
-As already noted \cite{Roth17}, the FEG shows unexpected %a peculiar
-behaviour in these elements that states doubts on the proper description.
-In the case of Hf we found the contribution of the 4f-shell decisive even at impact energies around the stopping maximum, as we will show later in this work.
+In this study, we report experimental stopping power cross section values over the incident energy range (0.6-2.5) MeV for protons crossing self-supported Hf thin-film by using the transmission method.
+We aim not only to upgrade stopping power data compilations \cite{HPaul03,mondim17} but also to provide useful information about the processes governing the slowing down of protons in multi-electronic targets.
+In the rare earth metals, the $4f$ electrons play a relevant role in the stopping power, being the first shell of bound electrons below the conduction band.
+As already noted \cite{Roth17}, the free electron gas (FEG) shows unexpected behavior in these elements, which casts doubts on its proper description.
+With Hf, we found the contribution of the 4f-shell decisive even at impact energies around the stopping maximum, as shown later.
 
-The theoretical approach used in this work requires 
-the relativistic wave functions and binding energies of Hf, and considering 4 electrons per atom in the free electron gas (FEG) \cite{mendez2019}.
-The shell-wise local plasma approximation (SLPA) \cite{mon13} was employed to describe the energy transferred to the bound 1s-4f electrons, and two different models for the FEG: the screened potential with cusp condition model (SPCC) \cite{mon17}, which is non-linear binary formalism%,  for energies below that of the plasmon excitation
-; and the Mermin-Lindhard dielectric formalism (ML) \cite{Mermin}, for energies around the stopping maximum and above.
+The theoretical approach implemented in this work uses the shell-wise local plasma approximation (SLPA) \cite{mon13} to describe the energy transferred to the bounds 1s-4f electrons, and two different models for the FEG: the screened potential with cusp condition model (SPCC) \cite{mon17}, which is non-linear binary formalism; and the Mermin-Lindhard dielectric formalism (ML) \cite{Mermin}, for energies around the stopping maximum and above.
+Our model requires the relativistic description of the wave functions and binding energies of Hf \cite{mendez2019}, where we consider 4 electrons per atom in the FEG. 
+The 14 electrons occupying the 4f subshell of Hf makes up the main contribution to the stopping below the FEG, which causes the cross sections to be very sensitive to a good description of this shell.
+The screening between the 4f and 5p electrons has been considered, and it has been found to play a major role within the SLPA predictions.
+
+%The theoretical approach used in this work requires the relativistic wave functions and binding energies of Hf, and considering 4 electrons per atom in the free electron gas (FEG) \cite{mendez2019}.
+The shell-wise local plasma approximation (SLPA) \cite{mon13} was employed to describe the energy transferred to the bound 1s-4f electrons, and two different models for the FEG: the screened potential with cusp condition model (SPCC) \cite{mon17}, which is non-linear binary formalism; and the Mermin-Lindhard dielectric formalism (ML) \cite{Mermin}, for energies around the stopping maximum and above.
 On the other hand, Hf has the extra interest of the filled $4f$-subshell (with 14 electrons) as the main contribution below the FEG, causing the stopping cross sections to be very sensitive to a good description of this shell.
@@
 
-The experimental details and data are given in section \ref{experiment}, while the theoretical method is explained in section \ref{theory}.
-Present theoretical and experimental values are finally compared to the available experimental values \cite{Sirotinin},  the ICRU-49 calculations \cite{ICRU49} and the SRIM-2013 package \cite{Ziegler01}.
-Conclusions and discusions are given in section \ref{conclusion}.
+The experimental details and data are given in Section \ref{experiment}, while the theoretical method is explained in Section \ref{theory}.
+Finally, our theoretical approach and experimental measurements are compared to the available experimental values \cite{Sirotinin}, the ICRU-49 calculations \cite{ICRU49}, and the SRIM-2013 package \cite{Ziegler01}.
+Conclusions and discussions are given in Section \ref{conclusion}.
 
@@
 \includegraphics[width=9cm]{Fig01.eps}
-\caption{RBS spectrum for $E_{avg} = 921,1$ keV protons on hafnium sample which is subsequently used to determine the energy loss in the foil.}
+\caption{RBS spectrum for $E_{\mathrm{avg}} = 921.1$ keV protons on hafnium sample, which is subsequently used to determine the energy loss in the foil.}
 \label{F01}
@@
 \subsection{Accelerator and scattering Chamber}
-The procedure used in this work to obtain stopping power data is essentially the same as described in reference \cite{Miranda01}. However, this time the measurements were made at the IST/LATR
-(Laboratory of Accelerators and X-Ray Diffraction) in Lisbon.
+The procedure carried out to obtain the stopping power data in this work is essentially the same as the one detailed in Ref.
+\cite{Miranda01}.
+However, the present measurements were performed at the IST/LATR (Laboratory of Accelerators and X-Ray Diffraction) in Lisbon.
 This facility uses a 2.5 MV Van de Graaff accelerator to deliver $^1$H$^+$ primary ion beams through a series of electrostatic lenses and collimators onto a thin Au/SiO$_2$ sample which is used as a scattering center.
-This sample was placed in the center of a RBS/C scattering chamber where a high vacuum (pressure of
-$\sim$10$^{-6}$ Torr) was maintained during the measurements. The beam current on sample was kept at around 5.0 nA in order to attain sufficient statistics in each particle spectrum.
+This sample was placed in the center of an RBS/C scattering chamber, where a high vacuum (pressure of $\sim$10$^{-6}$ Torr) was maintained during the measurements.
+The beam current on the sample was kept at around 5.0 nA to attain sufficient statistics in each particle spectrum.
 By using a beam spot of about 1.0 mm in diameter, a solid angle of 11.4 msr was attained.
@@
 \subsection{Target}
-The stopping material under analysis was a hafnium foil with nominal thickness of 1.0 $\mu$m and 99.95\% purity which was supplied by Lebow Company \cite{Lebow}.
+The stopping material under analysis was a hafnium foil with a nominal thickness of 1.0 $\mu$m and 99.95\% purity, which was supplied by Lebow Company \cite{Lebow}.
 However, a more precise thickness value was achieved by measuring the energy loss of alpha particles coming from a calibrated ($^{239}$Pu, $^{241}$Am, $^{244}$Cm) source.
-From the alpha spectra with and without the Hf foil interposed, the characteristic energy shift $\delta$E was measured and then combined with the stopping power for 5.486 MeV alphas on hafnium (55.69 eV/$10^{15}$ at/cm$^2$) found in Ref.\cite{Ziegler01} in order to obtain an areal density of (4,13$\pm$0.21)$\times 10^{19}$ at/cm$^2$ which corresponds to a thickness of 0.920$\pm$0.046 $\mu$m.
+From the alpha spectra with and without the Hf foil interposed, the characteristic energy shift $\delta$E was measured and then combined with the stopping power for 5.486 MeV alphas on hafnium (55.69 eV/$10^{15}$ at/cm$^2$) found in Ref.\cite{Ziegler01} to obtain an areal density of ($4.13\pm0.21$)$\times 10^{19}$ at/cm$^2$, which corresponds to a thickness of $0.920\pm0.046 \mu$m.
+
 %---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
@@
 Once the beam impinges on the Au/SiO$_2$ sample, protons are backscattered towards a Si surface barrier detector located at 140$^{\circ}$ relative to the initial beam direction.
-Fig.\ref{F01} shows two particle spectra where ion energies $E_2$ and $E_1$ are associated to a placed and removed hafnium sample, respectively.
-Both energy distributions were fitted by gaussian functions to obtain the mean energy and width (FWHM) of the peaks \cite{Sun01}, and from the difference between these two peak positions in the spectrum, the total energy loss $\Delta E = (E_1 - E_2)$ in the foil was calculated.
-As established in previous studies \cite{Miranda01, Damache02}, the experimental stopping powers cross sections $\varepsilon (E) $ are determined at some mean energy $E_{avg}$ by measuring the ion energy losses $\Delta E$ within the investigated Hf foil which has a mean thickness denoted by $\Delta x$.
-In this way, only when the energy loss fraction $\Delta E/E_{avg}$ across the Hf foil is not exceeding 20\%
-it is possible to define the stopping cross section by \cite{Raisanen01,Schulz01}:
+Fig.\ref{F01} shows two-particle spectra, where ion energies $E_2$ and $E_1$ are associated to a placed and removed hafnium sample, respectively.
+Both energy distributions were fitted by Gaussian functions to obtain the mean energy and width (FWHM) of the peaks \cite{Sun01}, and from the difference between these two peak positions in the spectrum, the total energy loss $\Delta E = (E_1 - E_2)$ in the foil was calculated.
+As established in previous studies \cite{Miranda01, Damache02}, the experimental stopping power cross sections $\varepsilon (E) $ are determined at some mean energy $E_{\mathrm{avg}}$ by measuring the ion energy losses $\Delta E$ within the investigated Hf foil, which has a mean thickness denoted by $\Delta x$.
+In this way, only when the energy loss fraction $\Delta E/E_{\mathrm{avg}}$ across the Hf foil is not exceeding 20\% it is possible to define the stopping cross section by \cite{Raisanen01,Schulz01}
 \begin{equation}\label{eq:stcross}
- \varepsilon (E) = \frac{S(E)}{N} = -\frac{dE}{N dx} \approx -\frac{\Delta E}{N \Delta x},
+ \varepsilon (E) = \frac{S(E)}{N} = -\frac{dE}{N dx} \approx -\frac{\Delta E}{N \Delta x}\,,
 \end{equation}
 where $N$ denotes the atomic number density (atoms cm$^{-3}$) of the material under study.
-When this condition was not fulfilled, a small correction to the mean energies $E_{avg}$ was applied in order to account for the nonlinear dependence on ion energy of stopping powers \cite{Chilton,Rajatora}.
+When this condition was not fulfilled, a small correction to the mean energies $E_{\mathrm{avg}}$ was applied in order to account for the nonlinear dependence on ion energy of stopping powers \cite{Chilton,Rajatora}.
 
-\section{Theoretical Method} \label{theory}
+\section{Theoretical Method} 
+\label{theory}
 The energy loss of ions in metal targets responds to different physical mechanisms, depending on the impact ion velocity.
-At low velocities the binary collisions are responsible of the loss of energy by the ion.
-The main contribution is the ionization of electrons of the metal conduction band, which is well approximated by a free electron gas (FEG) of Fermi velocity $v_F$.
-Instead, above certain velocity (i.e.
-$v\geq 1.5 \ v_F$) , not only binary but also collective excitations (plasmons) are possible \cite{mon17}.
-At high energies, not only the FEG contributes to the stopping power but also the bound electrons.
+At low velocities, the binary collisions are responsible for the loss of energy by the ion.
+The main contribution is the ionization of electrons in the metal conduction band, which is well approximated by a free electron gas (FEG) of Fermi velocity $v_F$.
+Above a certain velocity (i.e.
+$v\geq 1.5 \ v_F$), not only binary but also collective excitations (plasmons) are possible \cite{mon17}.
+Moreover, at high energies, both the FEG and the bound electrons contribute to the stopping power.
 The method used in this work combines a description for the interaction with the valence (or conduction) electrons as a FEG, and a different one for the interaction with the bound electrons.
@@
 We used the SPCC model \cite{mon17} to describe the stopping power of low velocity charged particles in the FEG.
-It is a non-perturbative binary collisional model, thus valid at energies below that of plasmon excitations.
-The SPCC \cite{mon17} is based on a screened central potential with cusp condition of the electronic density close to the projectile.
-It proved to give a good description of the induced electron density even for negative projectiles \cite{mon17}, and reproduces the low velocity proton-antiproton differences in the stopping power (Barkas effect).
+The model relies on a non-perturbative binary collisional approximation.
+The SPCC \cite{mon17} is based on a screened central potential with a cusp condition of the electronic density close to the projectile.
+This model proved to give a good description of the induced electron density even for negative projectiles \cite{mon17} and reproduces the low-velocity proton-antiproton differences in the stopping power (Barkas effect).
 The SPCC formalism only depends on the Wigner-Seitz radio, $r_S$, which is a measure of the electronic density of the FEG.
-For metals of well-known value of $r_S$, the SPCC describes very well the low energy experimental stopping data \cite{mon17}, agreeing with the DFT results by Echenique and coworkers \cite{eche81,nagy89} at $v=0$.
+For metals of well-known $r_S$ values, the SPCC describes correctly the low energy experimental stopping data \cite{mon17}, agreeing with the DFT results by Echenique and coworkers \cite{eche81,nagy89} at $v=0$.
 
-%In the case of 
-Hafnium (Z=72, [Xe] $4f^{14} \ 6s^2 \ 5d^2$), belongs to the first groups of transition metals, with  4 electrons as FEG (theoretical $r_S=2.14$ a.u.) and $1s-4f$ electrons bound.
-We checked the $r_S$ value with the experimental one obtained from measured energy loss function by Lynch \textit{et al} \cite{lynch75}.
-The experimental plasmon energy of Hf is $\hbar\omega_P \approx 15.8$ eV, with a width at half maximum $\delta \approx 4.4$ eV, and $r_S \approx 2.07$ a.u.
+Hafnium ($Z=72$, [Xe] $4f^{14} \ 6s^2 \ 5d^2$) belongs to the first group of transition metals, with  4 electrons as FEG (theoretical $r_S=2.14$ a.u.) and $1s-4f$ electrons bound.
+We compared the computed $r_S$ with the experimental value obtained from the measured energy loss function by Lynch \textit{et al} \cite{lynch75}.
+The experimental plasmon energy of Hf is $\hbar\omega_P \approx 15.8$ eV, with a width at half maximum of $\delta \approx 4.4$ eV, and $r_S \approx 2.07$ a.u.
 \cite{lynch75}.
-Following \cite{mon17}, the difference of less than $5\%$ between theoretical and experimental $r_S$ assess Hf as a canonical target.
+The difference of less than $5\%$ between theoretical and experimental $r_S$ assess Hf as a canonical target\cite{mon17}.
 
-Above certain impact velocity the plasmon contribution is important (i.e. around and above the stopping maximum). An interesting value for our analysis is the minimum impact velocity to excite
+Above certain impact velocity, the plasmon contribution is important (i.e. around and above the stopping maximum). An interesting value for our analysis is the minimum impact velocity to excite
 plasmons, $v_P$.
-In the dielectric formalism this value can be obtained as $v_P\ \approx \ v_F[1+(3 \pi\ v_F)^{-1/2} ]$ \cite{suppression}.
-To describe the energy loss considering collective and binary excitation we resort to the ML dielectric formalism \cite{Mermin}, which is a linear response, perturbative approximation, so it depends on the square of the ion charge.
+In the dielectric formalism, this value can be obtained as $v_P\ \approx \ v_F[1+(3 \pi\ v_F)^{-1/2} ]$ \cite{suppression}.
+To describe the energy loss considering collective and binary excitation, we resort to the ML dielectric formalism \cite{Mermin}, which is a linear response, perturbative approximation, so it depends on the square of the ion charge.
 In this formalism, the response of target electrons to the ion passage is described through the quantum dielectric function, depending on the characteristic $r_S$ and $\delta$ parameters of the FEG.
@@
 For the stopping power due to bound electrons, the SLPA \cite{mon17,mon13} is employed.
-It is worth to mention that the only inputs for the SLPA are the space-dependent densities of each shell in the ground state, and their binding energies.
+It is worth mentioning that the only inputs for the SLPA are the space-dependent densities of each shell in the ground state, and their binding energies.
 Collective processes and screening among electrons are included.
-Hafnium is a relativistic target, therefore the wave functions and binding energies must be obtained as solutions of Dirac equation  \cite{mendez2019}.
+Hafnium is a relativistic target, therefore the wave functions and binding energies must be obtained as solutions of the Dirac equation\cite{mendez2019}.
 
-Figure \ref{Binding_E} (a) shows non relativistic and our relativistic binding energies $E_{nl\pm}$, where $\pm=j\pm1/2$.
+To asses the importance of the relativistic description of the bound electrons, Figure \ref{Binding_E} (a) shows non relativistic and relativistic binding energies $E_{nl\pm}$, where $\pm=j\pm1/2$.
 We compare our results with experimental measurements on solid Hf \cite{williams1995} from $1s$ to $4f_{\pm}$.
-We notice that not only the deep shells require relativistic calculations, but also the outer $5p$ and $4f$ shells.
-Figure \ref{Binding_E} (b) displays the relative error of non relativistic and relativistic energies with respect to the experimental data in logarithmic scale.
+We notice that not only the most inner shells require relativistic calculations, but also the outer $5p$ and $4f$ shells.
+Figure \ref{Binding_E} (b) presents the relative error of non relativistic and relativistic energies with respect to the experimental data in logarithmic scale.
 This figure shows very clearly the disability of non-relativistic calculations to describe the experimental data, which surprisingly worsens from the inner to the outer shells.

### 2020-02-21 hafnio_ale.tex
include theoretical and conclusion corrections

@@
 The theoretical developments involved fully relativistic atomic structure calculations for Hf, which required the solution of the Dirac equation. 
-The shell-wise local plasma approximation (SLPA) was used to describe the energy transferred to the bound 1s-4f electrons, and the outer four electrons were considered as a free electron gas (FEG). 
- We found the relativistic description of the 4f-shell, and the screening between $4f$ and $5p$ electrons are decisive around the stopping maximum. 
+The shell-wise local plasma approximation (SLPA) was used to describe the energy transferred to the bound $1s$-$4f$ electrons, and the outer four electrons were considered as a free electron gas (FEG).
+ We found the relativistic description of the $4f$-shell, and the screening between $4f$ and $5p$ electrons are decisive around the stopping maximum. 
 The present theoretical and experimental results have very good agreement in the energy region of the new measurements. 
@@
 As already noted \cite{Roth17}, the free electron gas (FEG) shows unexpected behavior in these elements, which casts doubts on its proper description.
-With Hf, we found the contribution of the 4f-shell decisive even at impact energies around the stopping maximum, as shown later.
+With Hf, we found the contribution of the $4f$-shell decisive even at impact energies around the stopping maximum, as shown later.
 
-The theoretical approach implemented in this work uses the shell-wise local plasma approximation (SLPA) \cite{mon13} to describe the energy transferred to the bounds 1s-4f electrons, and two different models for the FEG: the screened potential with cusp condition model (SPCC) \cite{mon17}, which is non-linear binary formalism; and the Mermin-Lindhard dielectric formalism (ML) \cite{Mermin}, for energies around the stopping maximum and above.
+The theoretical approach implemented in this work uses the shell-wise local plasma approximation (SLPA) \cite{mon13} to describe the energy transferred to the bounds $1s$-$4f$ electrons, and two different models for the FEG: the screened potential with cusp condition model (SPCC) \cite{mon17}, which is non-linear binary formalism; and the Mermin-Lindhard dielectric formalism (ML) \cite{Mermin}, for energies around the stopping maximum and above.
 Our model requires the relativistic description of the wave functions and binding energies of Hf \cite{mendez2019}, where we consider 4 electrons per atom in the FEG. 
-The 14 electrons occupying the 4f subshell of Hf makes up the main contribution to the stopping below the FEG, which causes the cross sections to be very sensitive to a good description of this shell.
-The screening between the 4f and 5p electrons has been considered, and it has been found to play a major role within the SLPA predictions.
+The 14 electrons occupying the $4f$ subshell of Hf makes up the main contribution to the stopping below the FEG, which causes the cross sections to be very sensitive to a good description of this shell.
+The screening between the $4f$ and $5p$ electrons has been considered, and it has been found to play a major role within the SLPA predictions.
 
@@
 
-Hafnium ($Z=72$, [Xe] $4f^{14} \ 6s^2 \ 5d^2$) belongs to the first group of transition metals, with  4 electrons as FEG (theoretical $r_S=2.14$ a.u.) and $1s-4f$ electrons bound.
+Hafnium ($Z=72$, [Xe] $4f^{14} \ 6s^2 \ 5d^2$) belongs to the first group of transition metals, with  4 electrons as FEG (theoretical $r_S=2.14$ a.u.) and $1s$-$4f$ electrons bound.
 We compared the computed $r_S$ with the experimental value obtained from the measured energy loss function by Lynch \textit{et al} \cite{lynch75}.
@@
 Collective processes and screening among electrons are included.
-Hafnium is a relativistic target, therefore the wave functions and binding energies must be obtained as solutions of the Dirac equation\cite{mendez2019}.
+Hafnium is a relativistic target, therefore the wave functions and binding energies must be obtained as solutions of the Dirac equation.
 
-To asses the importance of the relativistic description of the bound electrons, Figure \ref{Binding_E} (a) shows non relativistic and relativistic binding energies $E_{nl\pm}$, where $\pm=j\pm1/2$.
-We compare our results with experimental measurements on solid Hf \cite{williams1995} from $1s$ to $4f_{\pm}$.
+To assess the importance of the description of the bound electrons, Fig.
+\ref{Binding_E} shows the (a) non relativistic and relativistic binding energies\cite{mendez2019}  $E_{nl\pm}$, where $\pm=j\pm1/2$, and (b) their relative errors with respect to experimental measurements on solid Hf \cite{williams1995}, from $1s$ to $4f_{\pm}$.
 We notice that not only the most inner shells require relativistic calculations, but also the outer $5p$ and $4f$ shells. 
-Figure \ref{Binding_E} (b) presents the relative error of non relativistic and relativistic energies with respect to the experimental data in logarithmic scale.
-This figure shows very clearly the disability of non-relativistic calculations to describe the experimental data, which surprisingly worsens from the inner to the outer shells.
-%It is clear in this figure the disability of non-relativistic binding energies to describe the experimental data, which surprisingly worsens from the inner to the outer-shells.
-%We used here the theoretical binding energies and wave functions of Hf calculated in \cite{mendez2019}. 
-%The present theoretical description of Hf considers a FEG of $r_S=2.14$ and 1s-4f electrons as bound.
+%Figure \ref{Binding_E} (b) presents the relative error of non relativistic and relativistic energies with respect to the experimental data. 
+Furthermore, this figure shows very clearly the disability of non-relativistic calculations to describe the experimental data, which surprisingly worsens from the inner to the outer shells.
 
-%The relativistic results for binding energies and wave functions in \cite{mendez2019} 
 Our relativistic binding energies present spin-orbit split. 
-However, in total stopping power where the initial state of the excited electron is not measured, the quantum uncertainty in energy $\Delta E$ melts this split.
+However, the quantum uncertainty in energy $\Delta E$ melts this split in total stopping power, where the initial state of the excited electron is not measured. 
 The criteria $\Delta E\Delta t \geq \hbar /2$ merges the energies  $E_{nl+} - E_{nl-}$ for sufficiently small values of $\Delta t$ (the collisional mean-time). 
-In fact, at sufficiently high impact velocity we can expect all target electrons to respond together to the ion passage \cite{lindhard53,chu72}.
-As in \cite{mon09}, the collisional time was estimated as follows $\Delta t \approx \langle r_i\rangle/v$, with $\langle r_i\rangle$ and $v$ being the orbital mean radio and impact velocity, respectively.
+In fact, at sufficiently high impact velocity, we can expect all target electrons to respond together to the ion passage \cite{lindhard53,chu72}. 
+Following previous work\cite{mon09}, the collisional time is estimated as $\Delta t \approx \langle r_i\rangle/v$, with $\langle r_i\rangle$ and $v$ being the orbital mean radio and impact velocity, respectively.
 
@@
 \includegraphics[width=13.cm]{Fig02.eps}
-\caption{Theoretical total stopping cross sections of protons in afnium adding FEG and bound $1s$-$4f$ contributions, as mentioned in the inset.
+\caption{Theoretical total stopping cross sections of protons in hafnium adding FEG and bound $1s$-$4f$ contributions, as mentioned in the inset.
 The vertical grey dashed-line indicates the energy above which plasmon excitation is possible.
 %The FEG calculations employ the SPCC model at low energies (no plasmon excitation) and the ML formalism (includes plasmon excitation).
-The SLPA calculations for $1s$-$4f$ bound electrons with and without $5p-4f$ screening give different total stopping values, displayed with solid and dotted lines, respectively.}
+The SLPA total stopping values for $1s$-$4f$ bound electrons with and without $5p$-$4f$ screening are given with solid and dotted lines, respectively.}
 \label{slpa4f}
@@
 We found that for every sub-shell of Hf, at the impact energy at which this sub-shell began to contribute, the spin-orbit split was unresolved.
-Then the $nl$-electrons should be considered together, responding to the ion passage as a single gas of electrons with density $\delta_{nl}(r)$ and a mean binding energy $E_{nl}$.
-This is important within the SLPA calculations because of the screening among electrons of the \textit{same} binding energy.
-For example, the $4f_{-}$ and $4f_{+}$ of Hf can only be resolved for impact energies $E<0.05$ keV, but the contribution of $4f$ to the total stopping is negligible if $E<40$ keV.
-Moreover, the $5p$ and $4f$ electrons of Hf are very close in energy ( $\Delta E_{5p-4f} \approx 1 a.u.$ \cite{mendez2019}) %merges them
-and react together for impact energies $E>40$ keV (inter-shell screening).
+Then, the $nl$-electrons should be considered together, responding to the ion passage as a single gas of electrons with density $\delta_{nl}(r)$ and a mean binding energy $E_{nl}$. 
+This feature is important within the SLPA calculations because it accounts for the screening among electrons of the \textit{same} binding energy.
+For example, the $4f_{-}$ and $4f_{+}$ of Hf can only be resolved for impact energies $E<0.05$ keV, but the contribution of $4f$ to the total stopping is negligible for $E<40$ keV.
+Moreover, the $5p$ and $4f$ electrons of Hf are very close in energy ( $\Delta E_{5p-4f} \approx 1 $ a.u.\cite{mendez2019})
+and they react together for impact energies $E>40$ keV (inter-shell screening).
 
@@
 
-In figure \ref{slpa4f} we display the present theoretical stopping cross section of Hf for protons with and without the $5p-4f$ screening.
-We added the FEG and the bound electron contributions as explained above.
-The minimum energy for plasmon excitation was estimated as $37$ keV.
+In Fig.
+\ref{slpa4f}, we display the present theoretical stopping cross section of Hf for protons with and without the $5p$-$4f$ screening.
+We added the FEG and the bound electron contributions, as explained above.
+The minimum energy for plasmon excitation was estimated to $37$ keV.
 We used the non-perturbative SPCC model for impact energies $E \leq 37$ keV, and the perturbative ML calculation above this energy.
@@
 %The importance of the 4f-5p screening is signed with arrows in this figure. 
-Clearly, considering $5p-4f$ electrons as a single group of 20 electrons with screening among them gives lower stopping values than the addition of the separate $5p$ and $4f$ contributions.
-This is a shell correction that can only be considered within a many electron model such as the SLPA.
+Clearly, considering $5p$-$4f$ electrons as a single group of 20 electrons with screening among them gives lower stopping values than the addition of the separate $5p$ and $4f$ contributions.
+Notice that this shell correction can only be considered within a many electron model such as the SLPA.
 
@@
 
-The present data and theoretical results are displayed in table \ref{table01}, and figure \ref{F03}.
-As can be seen in Table~\ref{table01}, an overall relative uncertainty of around 5$\%$ was achieved for the experimental stopping power values, mainly due to the uncertainty in the hafnium foil thickness.
+The present data and theoretical results are displayed in Table \ref{table01} and Fig.
+\ref{F03}.
+As can be seen in Table~\ref{table01}, an overall relative uncertainty of around 5$\%$ was achieved for the experimental stopping power values, which are mainly due to the uncertainty in the hafnium foil thickness.
 %A good agreement in the whole energy range is found between our new data and the previous measurements made by Sirotinin \textit{et al.} \cite{Sirotinin}, as depicted in Fig.~\ref{F03}.
@@
 
-There is good agreement between the present theoretical results, the new data in Table~\ref{table01}, and also with previous data by Sirotinin \cite{Sirotinin}, except for the lowest energy measurement at 80 keV.
-It is interesting that our full theoretical curve differs from SRIM-2013 one for impact energies below 100 keV.
+In Fig.
+\ref{F03}, we have good agreement between the present theoretical results and the measurements given in Table~\ref{table01}.
+Our theoretical approach also agrees with previous data by Sirotinin \cite{Sirotinin}, except for the lowest energy measurement at 80 keV.
+It is interesting that our full theoretical curve differs from SRIM-2013 for impact energies below 100 keV.
 We obtain a stopping maximum around 40 $10^{-15}$ eV cm$^2$/atom  at 65 keV.
@@
 at impact energy of 115 kev.
-It is worth to mention that similar theoretical results using the experimental $r_S=2.07$ a.u.
-instead of the theoretical $r_S=2.14$ a.u.
-give the stopping maximum at the same impact energy but $4\%$ higher.
-Future experiments for
-impact energies below 100 keV would be important to complete this study.
+It is worth mentioning that similar theoretical results, using the experimental $r_S=2.07$ a.u.
+instead of the theoretical $r_S=2.14$ a.u., give the stopping maximum at the same impact energy but $4\%$ higher.
+Future experiments for impact energies below 100 keV would be important to complete this study.
 
@@
 Additionally, we calculated values extracted from the theoretical framework that involved the relativistic wave functions and binding energies of Hf, and considered 4 electrons per atom in the free electron gas.
-The shell-wise local plasma approximation was employed to describe the energy transferred to the bound 1s-4f electrons, and two different models for the FEG: the screened potential with cusp condition (SPCC model) for energies below that of the plasmon excitation, and the Mermin-Lindhard dielectric formalism, for energies around the stopping maximum and above.
+The shell-wise local plasma approximation was employed to describe the energy transferred to the bound $1s$-$4f$ electrons, and two different models for the FEG: the screened potential with cusp condition (SPCC model) for energies below that of the plasmon excitation, and the Mermin-Lindhard dielectric formalism for energies around the stopping maximum and above.
 Present theoretical stopping cross sections cover an extensive energy range from 1 keV/amu-10 meV/amu.
 
-%Our results have been confronted to both experimental values found in the literature and semi-empirical results. 
-At high impact energies, the new stopping  measurements are in good agreement with our theoretical results, and also with already published experimental data and semi-empirical values calculated by means of SRIM-2013 and ICRU-49.
+At high impact energies, the new stopping measurements are in good agreement with our theoretical results.
+We also agree with previously published experimental data and semi-empirical values calculated by means of SRIM-2013 and ICRU-49.
 However, we call the attention that around the stopping maximum and at lower impact energies the difference between our full-theoretical results and SRIM is substantial.
-To our knowledge, these are the first theoretical calculations of stopping in Hf taking into account the atomic relativistic effect.
-%for energies around and below the stopping maximum, the present full-theoretical results clearly differ from semiempirical SRIM values, which are based on previous data. 
-Future experiments for impact energies below 100 keV would be important to complete this study.
+To the best of our knowledge, these are the first theoretical calculations of stopping in Hf taking into account the atomic relativistic effect.
+Future experiments for impact energies below 100 keV would be required to validate the predictions of our theoretical model.
 %=======================================================================================================================================
 \begin{acknowledgments}
-This work was funded by VRAC Grant Number L1-17 of Universidad Tecnol\'ogica Metropolitana, Chile; also by the Consejo Nacional de Investigaciones Científicas y Técnicas (CONICET), the Agencia Nacional de Promoción Científica y Tecnológica (ANPCyT), and Universidad de Buenos Aires (UBA), of Argentina.
+This work was funded by VRAC Grant Number L1-17 of Universidad Tecnol\'ogica Metropolitana, Chile; also by the Consejo Nacional de Investigaciones Cient\'ificas y T\'ecnicas (CONICET), the Agencia Nacional de Promoci\'on Cient\'ifica y Tecnol\'ogica (ANPCyT), and Universidad de Buenos Aires (UBA), from Argentina.
 

### 2020-03-11 Paper_SPHfP_feb2020_prb.tex
final corrections

@@
 %
-\documentclass[aps,prb,reprint,groupedaddress]{revtex4-1}
+\documentclass[aps,pra,reprint,groupedaddress]{revtex4-1}
 
@@
 %Title of paper
-\title{The role of $4f$ electrons in the stopping power of hafnium}
+\title{The stopping power of hafnium and the importance of relativistic $4f$ electrons}
 %Authors and affiliations
@@
  The stopping power of protons through Hf foil has been studied both experimentally and theoretically.
-The measurements were performed at the Laboratory of Accelerators and X-Ray Diffraction in Lisbon, by using the transmission method on self-supporting stopping material, for (0.6-2.5) MeV protons.
+The measurements were performed at the Laboratory of Accelerators and X-Ray Diffraction in Lisbon, by using the transmission method on self-supporting stopping material.
+%for (0.6-2.5) MeV protons.
+ The overall uncertainty of around $5\%$ was established over the protons energy range  (0.6-2.5) MeV.
 The theoretical developments involved fully relativistic atomic structure calculations for %this Z=72 target
  Hf, which required the solution of the Dirac equation.
-The shell-wise local plasma approximation (SLPA) was employed to describe the energy transferred to the bound 1s-4f electrons, and the outer four electrons were considered as a free electron gas (FEG).
+The shell-wise local plasma approximation (SLPA) was used to describe the energy transferred to the bound $1s$-$4f$ electrons, and the outer four electrons were considered as a free electron gas (FEG).
 %We contemplated two different models for the FEG: the screened potential with cusp condition (SPCC) for energies below that of the plasmon excitation, and the dielectric formalism, for energies around the stopping maximum and above.
- We found the relativistic description of the 4f-shell and the screening between $4f$ and $5p$ electrons are decisive around the stopping maximum. %for energies above 40 keV.
- Present theoretical and experimental results are in very good agreement in the energy region of the new measurements. However, our theoretical stopping cross sections show %interesting 
- substantial differences with the semi-empirical values by SRIM2013 and ICRU-49 at intermediate to low energies.
+ We found the relativistic description of the $4f$-shell and the screening between $4f$ and $5p$ electrons are decisive around the stopping maximum. %for energies above 40 keV.
+ Present theoretical and experimental results are in very good agreement in the energy region of the new measurements. However, the present theoretical stopping cross sections show %interesting 
+ substantial differences with the most used semi-empirical models (SRIM2013 and ICRU-49) at intermediate to low energies.
 Our calculations suggest the stopping maximum %around 40 eV$/(10^{15}$ atom/cm$^2)$ at 70 keV, which is
- higher and shifted to lower energies than previous predictions. Future measurements below 100 keV would be decisive for a better knowledge of the stopping maximum and below.
+ higher and shifted to lower energies than these previous predictions. Future measurements below 100 keV would be decisive for a better knowledge  around the maximum and below.
 \end{abstract}
@@
 % insert suggested PACS numbers in braces on next line
-\pacs{34.50.Bw} \keywords{Stopping power, Hafnium, Relativistic atomic structure, Transmission method}
+\pacs{34.50.Bw} \keywords{Stopping power, hafnium, Relativistic atomic structure, Transmission method}
 
@@
 The information given by the energy loss process is essential not only to have a better knowledge of the physics behind the fundamental interactions, but also because it plays an important role in many applied fields such as materials science, nuclear physics, ionic implantation and radiotherapy \cite{Sigmund,Schardt}.
-Moreover, experimental values on ion mean energy loss per unit path $S(E)$ are of crucial relevance in order to check the reliability of semi-empirical models and to determine some key parameters
+Experimental data on ion mean energy loss per unit path $S(E)$ is of crucial relevance to check the reliability of semi-empirical models and to determine some key parameters
 %included in theoretical approaches of stopping power %\cite{Diwan,Damache01,Damache02,Damache03,Damache04}. 
-\cite{Diwan,Damache04,Damache02}.
-However, very often, the experimental data available is rather scarce which is particularly troublesome when the material under study corresponds to an element of low occurrence on the Earth's upper crust, such as Hf.
+\cite{Diwan,Damache04,Damache02}. The experimental data available is often rather scarce, which is %particularly 
+troublesome when the material under study corresponds to an element of low occurrence on the Earth's upper crust, such as hafnium.
 
 So far, only one experimental work has been published regarding the stopping power cross section of pure hafnium for protons \cite{Sirotinin}, while more attention has been recently given to studies involving hafnium oxide due to its practical use \cite{Abril,Behar,Primetzhofer}.
-It is well known that significant attention has been paid in recent years to transition metaloxides such as HfO$_2$ for their potential as alternative gate dielectrics to replace SiO$_2$ for the future generation of nano-electronics with less than 45 nm gate length \cite{Choi,Robertson}.
-In this regard, some important physical properties of the above mentioned metal-oxide films depend on its thickness, which is often measured by using Rutherford Backscattering Spectrometry \cite{Alfassi01, Tesmer01}, a methodology that relies heavily on the determination of the stopping power of ion beams in the material of interest.
+It is well known that significant attention has been paid in recent years to transition metal-oxides such as HfO$_2$ because of their potential as alternative gate dielectrics to replace SiO$_2$ for the future generation of nano-electronics with less than 45 nm gate length \cite{Choi,Robertson}.
+Some important physical properties of the above mentioned metal-oxide films depend on its thickness, which is often measured by using Rutherford Backscattering Spectrometry \cite{Alfassi01, Tesmer01}, a method that relies heavily on the determination of the stopping power of ion beams in the material of interest.
 
-In this study we report stopping power cross section experimental values over the incident energy range (0.6-2.5) MeV for protons crossing self-supported Hf thin-film by using the transmission method.
-Our aim is not only to upgrade stopping power data compilations \cite{HPaul03,mondim17} but also to provide useful information about the processes governing the slowing down of protons in multielectronic targets.
-In the rare earth metals, the 4f electrons play a relevant role in the stopping power, being the first shell of bound electrons below the conduction band.
-As already noted \cite{Roth17}, the FEG shows unexpected %a peculiar
-behaviour in these elements that states doubts on the proper description.
-In the case of Hf we found the contribution of the 4f-shell decisive even at impact energies around the stopping maximum, as we will show later in this work.
+In this study, we report experimental stopping power cross sections over the incident energy range (0.6-2.5) MeV for protons crossing self-supported Hf thin-film by using the transmission method.
+We aim not only to upgrade stopping power data compilations \cite{HPaul03,mondim17} but also to provide useful information about the processes governing the slowing down of protons in multi-electronic targets.
+In the rare earth metals, the $4f$ electrons play a relevant role in the stopping power, being the first shell of bound electrons below the conduction band.
+As already noted \cite{Roth17}, the free electron gas (FEG) shows unexpected %a peculiar
+behavior in these elements, which casts %that states 
+doubts on its proper description. In the case of Hf, we found the contribution of the $4f$-shell decisive even at impact energies around the stopping maximum, as shown later.
 
-The theoretical approach used in this work requires 
-the relativistic wave functions and binding energies of Hf, and considering 4 electrons per atom in the free electron gas (FEG) \cite{mendez2019}.
-The shell-wise local plasma approximation (SLPA) \cite{mon13} was employed to describe the energy transferred to the bound 1s-4f electrons, and two different models for the FEG: the screened potential with cusp condition model (SPCC) \cite{mon17}, which is non-linear binary formalism%,  for energies below that of the plasmon excitation
-; and the Mermin-Lindhard dielectric formalism (ML) \cite{Mermin}, for energies around the stopping maximum and above.
-On the other hand, Hf has the extra interest of the filled $4f$-subshell (with 14 electrons) as the main contribution below the FEG, causing the stopping cross sections to be very sensitive to a good description of this shell.
+The theoretical approach implemented in this work uses the shell-wise local plasma approximation (SLPA) \cite{mon13} to describe the energy transferred to the bound $1s$-$4f$ electrons, and two different models for the FEG: the screened potential with cusp condition model (SPCC) \cite{mon17}, which is non-linear binary formalism%,  for energies below that of the plasmon excitation
+; and the Mermin-Lindhard dielectric formalism (ML) \cite{Mermin}, for energies around the stopping maximum and above. Our model requires 
+the relativistic wave functions and binding energies of Hf, and considers 4 electrons per atom in the  FEG \cite{mendez2019}. 
+%On the other hand,
+Hf has the extra interest of the filled $4f$-subshell (with 14 electrons) as the main contribution below the FEG, causing the stopping cross sections to be very sensitive to a good description of this shell.
 The screening among the $4f$ and $5p$ electrons has been considered and found to play a major role within the SLPA calculations.
 
-The experimental details and data are given in section \ref{experiment}, while the theoretical method is explained in section \ref{theory}.
+The experimental details and data are given in Section \ref{experiment}, while the theoretical method is explained in Section \ref{theory}.
 Present theoretical and experimental values are finally compared to the available experimental values \cite{Sirotinin},  the ICRU-49 calculations \cite{ICRU49} and the SRIM-2013 package \cite{Ziegler01}.
-Conclusions and discusions are given in section \ref{conclusion}.
+Conclusions and discussions are given in Section \ref{conclusion}.
+All the present data can be found at the Zenodo platform \cite{zenodo}.
 
@@
 \includegraphics[width=9cm]{Fig01.eps}
-\caption{RBS spectrum for $E_{avg} = 921,1$ keV protons on hafnium sample which is subsequently used to determine the energy loss in the foil.}
+\caption{RBS spectrum for $E_{\mathrm{avg}} = 921.1$ keV protons on hafnium sample which is subsequently used to determine the energy loss in the foil.}
 \label{F01}
@@
 
-\section{Experimental arrangements}
+\section{EXPERIMENTAL ARRANGEMENTS}
 \label{experiment}
 \subsection{Accelerator and scattering Chamber}
-The procedure used in this work to obtain stopping power data is essentially the same as described in reference \cite{Miranda01}. However, this time the measurements were made at the IST/LATR
+The procedure used in this work to obtain stopping power data is essentially the same as described in Ref. \cite{Miranda01}. The present measurements were made at the IST/LATR
 (Laboratory of Accelerators and X-Ray Diffraction) in Lisbon.
 This facility uses a 2.5 MV Van de Graaff accelerator to deliver $^1$H$^+$ primary ion beams through a series of electrostatic lenses and collimators onto a thin Au/SiO$_2$ sample which is used as a scattering center.
-This sample was placed in the center of a RBS/C scattering chamber where a high vacuum (pressure of
-$\sim$10$^{-6}$ Torr) was maintained during the measurements. The beam current on sample was kept at around 5.0 nA in order to attain sufficient statistics in each particle spectrum.
+This sample was placed in the center of a RBS/C scattering chamber, where a high vacuum (pressure of
+$\sim$10$^{-6}$ Torr) was maintained during the measurements. The beam current on the sample was kept at around 5.0 nA to attain sufficient statistics in each particle spectrum.
 By using a beam spot of about 1.0 mm in diameter, a solid angle of 11.4 msr was attained.
@@
 However, a more precise thickness value was achieved by measuring the energy loss of alpha particles coming from a calibrated ($^{239}$Pu, $^{241}$Am, $^{244}$Cm) source.
-From the alpha spectra with and without the Hf foil interposed, the characteristic energy shift $\delta$E was measured and then combined with the stopping power for 5.486 MeV alphas on hafnium (55.69 eV/$10^{15}$ at/cm$^2$) found in Ref.\cite{Ziegler01} in order to obtain an areal density of (4,13$\pm$0.21)$\times 10^{19}$ at/cm$^2$ which corresponds to a thickness of 0.920$\pm$0.046 $\mu$m.
+From the alpha spectra with and without the Hf foil interposed, the characteristic energy shift $\delta$E was measured and then combined with the stopping power for 5.486 MeV alphas on hafnium (55.69 eV/$10^{15}$ at/cm$^2$) found in Ref.\cite{Ziegler01} to obtain an areal density of ($4.13 \pm 0.21$)$\times 10^{19}$ at/cm$^2$ which corresponds to a thickness of $0.920\pm0.046 \mu$m.
 %---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
 \subsection{Energy loss measurement}
-Once the beam impinges on the Au/SiO$_2$ sample, protons are backscattered towards a Si surface barrier detector located at 140$^{\circ}$ relative to the initial beam direction.
+Once the beam impinges on the Au/SiO$_2$ sample, protons are back-scattered towards a Si surface barrier detector located at 140$^{\circ}$ relative to the initial beam direction.
 Fig.\ref{F01} shows two particle spectra where ion energies $E_2$ and $E_1$ are associated to a placed and removed hafnium sample, respectively.
-Both energy distributions were fitted by gaussian functions to obtain the mean energy and width (FWHM) of the peaks \cite{Sun01}, and from the difference between these two peak positions in the spectrum, the total energy loss $\Delta E = (E_1 - E_2)$ in the foil was calculated.
-As established in previous studies \cite{Miranda01, Damache02}, the experimental stopping powers cross sections $\varepsilon (E) $ are determined at some mean energy $E_{avg}$ by measuring the ion energy losses $\Delta E$ within the investigated Hf foil which has a mean thickness denoted by $\Delta x$.
-In this way, only when the energy loss fraction $\Delta E/E_{avg}$ across the Hf foil is not exceeding 20\%
+Both energy distributions were fitted by Gaussian functions to obtain the mean energy and width (FWHM) of the peaks \cite{Sun01}, and from the difference between these two peak positions in the spectrum, the total energy loss $\Delta E = (E_1 - E_2)$ in the foil was calculated.
+As established in previous studies \cite{Miranda01, Damache02}, the experimental stopping power cross sections $\varepsilon (E) $ are determined at some mean energy $E_{\mathrm{avg}}$ by measuring the ion energy losses $\Delta E$ within the investigated Hf foil, which has a mean thickness denoted by $\Delta x$.
+In this way, only when the energy loss fraction $\Delta E/E_{\mathrm{avg}}$ across the Hf foil is not exceeding 20\%
 it is possible to define the stopping cross section by \cite{Raisanen01,Schulz01}:
@@
 where $N$ denotes the atomic number density (atoms cm$^{-3}$) of the material under study.
-When this condition was not fulfilled, a small correction to the mean energies $E_{avg}$ was applied in order to account for the nonlinear dependence on ion energy of stopping powers \cite{Chilton,Rajatora}.
+When this condition was not fulfilled, a small correction to the mean energies $E_{\mathrm{avg}}$ was applied in order to account for the nonlinear dependence on ion energy of stopping powers \cite{Chilton,Rajatora}.
 
@@
 The energy loss of ions in metal targets responds to different physical mechanisms, depending on the impact ion velocity.
-At low velocities the binary collisions are responsible of the loss of energy by the ion.
+At low velocities, the binary collisions are responsible for the loss of energy by the ion.
 The main contribution is the ionization of electrons of the metal conduction band, which is well approximated by a free electron gas (FEG) of Fermi velocity $v_F$.
-Instead, above certain velocity (i.e.
+Above certain velocity (i.e.
 $v\geq 1.5 \ v_F$) , not only binary but also collective excitations (plasmons) are possible \cite{mon17}.
-At high energies, not only the FEG contributes to the stopping power but also the bound electrons.
+Moreover, at high energies, also the bound electrons contribute to the stopping power.
 The method used in this work combines a description for the interaction with the valence (or conduction) electrons as a FEG, and a different one for the interaction with the bound electrons.
@@
 We used the SPCC model \cite{mon17} to describe the stopping power of low velocity charged particles in the FEG.
-It is a non-perturbative binary collisional model, thus valid at energies below that of plasmon excitations.
+It is a non-perturbative binary collisional approximation, thus valid at energies below that of plasmon excitations.
 The SPCC \cite{mon17} is based on a screened central potential with cusp condition of the electronic density close to the projectile.
-It proved to give a good description of the induced electron density even for negative projectiles \cite{mon17}, and reproduces the low velocity proton-antiproton differences in the stopping power (Barkas effect).
+This model proved to give a good description of the induced electron density even for negative projectiles \cite{mon17}, and reproduces the low velocity proton-antiproton differences in the stopping power (Barkas effect).
 The SPCC formalism only depends on the Wigner-Seitz radio, $r_S$, which is a measure of the electronic density of the FEG.
-For metals of well-known value of $r_S$, the SPCC describes very well the low energy experimental stopping data \cite{mon17}, agreeing with the DFT results by Echenique and coworkers \cite{eche81,nagy89} at $v=0$.
+For metals of well-known $r_S$, the SPCC describes correctly the low energy experimental stopping data \cite{mon17}, agreeing with the DFT results by Echenique and coworkers \cite{eche81,nagy89} at $v=0$.
 
 %In the case of 
-Hafnium (Z=72, [Xe] $4f^{14} \ 6s^2 \ 5d^2$), belongs to the first groups of transition metals, with  4 electrons as FEG (theoretical $r_S=2.14$ a.u.) and $1s-4f$ electrons bound.
-We checked the $r_S$ value with the experimental one obtained from measured energy loss function by Lynch \textit{et al} \cite{lynch75}.
+Hafnium ($Z=72$, [Xe] $4f^{14} \ 6s^2 \ 5d_{3/2}^1 \ 5d_{5/2}^1$), belongs to the first groups of transition metals, with  4 electrons as FEG (theoretical $r_S=2.14$ a.u.) and $1s$-$4f$ electrons bound.
+We compared the computed $r_S$ with the experimental value obtained from measured energy loss function by Lynch \textit{et al} \cite{lynch75}.
 The experimental plasmon energy of Hf is $\hbar\omega_P \approx 15.8$ eV, with a width at half maximum $\delta \approx 4.4$ eV, and $r_S \approx 2.07$ a.u.
 \cite{lynch75}.
-Following \cite{mon17}, the difference of less than $5\%$ between theoretical and experimental $r_S$ assess Hf as a canonical target.
+The difference of less than $5\%$ between theoretical and experimental $r_S$ assess Hf as a canonical target \cite{mon17}.
 
-Above certain impact velocity the plasmon contribution is important (i.e. around and above the stopping maximum). An interesting value for our analysis is the minimum impact velocity to excite
+Above certain impact velocity, the plasmon contribution is important (i.e. around and above the stopping maximum). An interesting value for our analysis is the minimum impact velocity to excite
 plasmons, $v_P$.
-In the dielectric formalism this value can be obtained as $v_P\ \approx \ v_F[1+(3 \pi\ v_F)^{-1/2} ]$ \cite{suppression}.
-To describe the energy loss considering collective and binary excitation we resort to the ML dielectric formalism \cite{Mermin}, which is a linear response, perturbative approximation, so it depends on the square of the ion charge.
+In the dielectric formalism, this value can be obtained as $v_P\ \approx \ v_F[1+(3 \pi\ v_F)^{-1/2} ]$ \cite{suppression}.
+To describe the energy loss considering collective and binary excitation, we resort to the ML dielectric formalism \cite{Mermin}, which is a linear response, perturbative approximation, so it depends on the square of the ion charge.
 In this formalism, the response of target electrons to the ion passage is described through the quantum dielectric function, depending on the characteristic $r_S$ and $\delta$ parameters of the FEG.
@@
 \centering
-\includegraphics[width=11.cm]{Hf_bindener_bar.eps}
+\includegraphics[width=11.cm]{bindener.eps}
 % \includegraphics[width=13.cm]{Hf_bindener_ratio.eps}
 \caption{(a) Binding energies of Hf.
-Non relativistic and relativistic calculations are given with filled symbols.
+Non-relativistic and relativistic calculations are given with filled symbols.
 Experimental measurements for solids\cite{williams1995} are depicted with hollow circles.
@@
 Collective processes and screening among electrons are included.
-Hafnium is a relativistic target, therefore the wave functions and binding energies must be obtained as solutions of Dirac equation  \cite{mendez2019}.
+Hafnium is a relativistic target, therefore the wave functions and binding energies must be obtained by solving the many–electron Dirac Hamiltonian.
+Details of these calculations have been published in \cite{mendez2019}, and the Slater-type orbital expansions in \cite{Hf_arxiv}.
 
-Figure \ref{Binding_E} (a) shows non relativistic and our relativistic binding energies $E_{nl\pm}$, where $\pm=j\pm1/2$.
-We compare our results with experimental measurements on solid Hf \cite{williams1995} from $1s$ to $4f_{\pm}$.
-We notice that not only the deep shells require relativistic calculations, but also the outer $5p$ and $4f$ shells.
-Figure \ref{Binding_E} (b) displays the relative error of non relativistic and relativistic energies with respect to the experimental data in logarithmic scale.
-This figure shows very clearly the disability of non-relativistic calculations to describe the experimental data, which surprisingly worsens from the inner to the outer shells.
+Too assess the importance of the description of bound electrons, Fig.
+\ref{Binding_E} shows (a) non-relativistic \cite{badnell97} and our relativistic binding energies, $E_{nl\pm}$, with $\pm=j\pm1/2$; and (b) the relative errors with respect to the experimental data on solid Hf \cite{williams1995} from $1s$ to $4f_{\pm}$.
+We notice that not only the most inner shells require relativistic calculations, but also the outer $5p$ and $4f$ shells.
+%Figure \ref{Binding_E} (b) displays the relative error of non relativistic and relativistic energies with respect to the experimental data in logarithmic scale. 
+Furthermore, this figure shows very clearly the disability of non-relativistic calculations to describe the experimental data, which surprisingly worsens from the inner to the outer shells. 
 %It is clear in this figure the disability of non-relativistic binding energies to describe the experimental data, which surprisingly worsens from the inner to the outer-shells.
@@
 The criteria $\Delta E\Delta t \geq \hbar /2$ merges the energies  $E_{nl+} - E_{nl-}$ for sufficiently small values of $\Delta t$ (the collisional mean-time).
-In fact, at sufficiently high impact velocity we can expect all target electrons to respond together to the ion passage \cite{lindhard53,chu72}.
-As in \cite{mon09}, the collisional time was estimated as follows $\Delta t \approx \langle r_i\rangle/v$, with $\langle r_i\rangle$ and $v$ being the orbital mean radio and impact velocity, respectively.
+In fact, at sufficiently high impact velocity, we can expect all target electrons to respond together to the ion passage \cite{lindhard53,chu72}.
+Following previous work \cite{mon09}, the collisional time is estimated as $\Delta t \approx \langle r_i\rangle/v$, with $\langle r_i\rangle$ and $v$ being the orbital mean radio and impact velocity, respectively.
 
@@
 \includegraphics[width=13.cm]{Fig02.eps}
-\caption{Theoretical total stopping cross sections of protons in afnium adding FEG and bound $1s$-$4f$ contributions, as mentioned in the inset.
+\caption{Theoretical total stopping cross sections of protons in hafnium adding FEG and bound $1s$-$4f$ contributions, as mentioned in the inset.
 The vertical grey dashed-line indicates the energy above which plasmon excitation is possible.
 %The FEG calculations employ the SPCC model at low energies (no plasmon excitation) and the ML formalism (includes plasmon excitation).
-The SLPA calculations for $1s$-$4f$ bound electrons with and without $5p-4f$ screening give different total stopping values, displayed with solid and dotted lines, respectively.}
+The SLPA total stopping values for $1s$-$4f$ bound electrons with and without $5p-4f$ screening are displayed with solid and dotted lines, respectively.}
 \label{slpa4f}
@@
 We found that for every sub-shell of Hf, at the impact energy at which this sub-shell began to contribute, the spin-orbit split was unresolved.
-Then the $nl$-electrons should be considered together, responding to the ion passage as a single gas of electrons with density $\delta_{nl}(r)$ and a mean binding energy $E_{nl}$.
-This is important within the SLPA calculations because of the screening among electrons of the \textit{same} binding energy.
-For example, the $4f_{-}$ and $4f_{+}$ of Hf can only be resolved for impact energies $E<0.05$ keV, but the contribution of $4f$ to the total stopping is negligible if $E<40$ keV.
-Moreover, the $5p$ and $4f$ electrons of Hf are very close in energy ( $\Delta E_{5p-4f} \approx 1 a.u.$ \cite{mendez2019}) %merges them
-and react together for impact energies $E>40$ keV (inter-shell screening).
+Then, the $nl$-electrons should be considered together, responding to the ion passage as a single gas of electrons with density $\delta_{nl}(r)$ and a mean binding energy $E_{nl}$.
+This feature is important within the SLPA calculations because it accounts for the screening among electrons of the \textit{same} binding energy.
+For example, the $4f_{-}$ and $4f_{+}$ of Hf can only be resolved for impact energies $E<0.05$ keV, but the contribution of $4f$ to the total stopping is negligible for $E<40$ keV.
+Moreover, the $5p$ and $4f$ electrons of Hf are very close in energy ( $\Delta E_{5p-4f} \approx 1$ a.u.
+\cite{mendez2019}) %merges them
+and they react together for impact energies $E>40$ keV (inter-shell screening).
 
@@
 
-In figure \ref{slpa4f} we display the present theoretical stopping cross section of Hf for protons with and without the $5p-4f$ screening.
+In Fig.
+\ref{slpa4f}, we display the present theoretical stopping cross section of Hf for protons with and without the $5p$-$4f$ screening.
 We added the FEG and the bound electron contributions as explained above.
-The minimum energy for plasmon excitation was estimated as $37$ keV.
+The minimum energy for plasmon excitation was estimated to $37$ keV.
 We used the non-perturbative SPCC model for impact energies $E \leq 37$ keV, and the perturbative ML calculation above this energy.
@@
 
-The present data and theoretical results are displayed in table \ref{table01}, and figure \ref{F03}.
-As can be seen in Table~\ref{table01}, an overall relative uncertainty of around 5$\%$ was achieved for the experimental stopping power values, mainly due to the uncertainty in the hafnium foil thickness.
+The present data and theoretical results are displayed in Table \ref{table01}, and Fig.
+\ref{F03}.
+As can be seen in Table~\ref{table01}, an overall relative uncertainty of around 5$\%$ was achieved for the experimental stopping power values, which are mainly due to the uncertainty in the hafnium foil thickness.
 %A good agreement in the whole energy range is found between our new data and the previous measurements made by Sirotinin \textit{et al.} \cite{Sirotinin}, as depicted in Fig.~\ref{F03}.
@@
 
-There is good agreement between the present theoretical results, the new data in Table~\ref{table01}, and also with previous data by Sirotinin \cite{Sirotinin}, except for the lowest energy measurement at 80 keV.
-It is interesting that our full theoretical curve differs from SRIM-2013 one for impact energies below 100 keV.
-We obtain a stopping maximum around 40 $10^{-15}$ eV cm$^2$/atom  at 65 keV.
-Instead, following the up-to-now only set of data \cite{Sirotinin}, SRIM-2013 suggests a lower stopping maximum %of 37.4 $10^{-15}$ eV cm$^2$/atom
-at impact energy of 115 kev.
-It is worth to mention that similar theoretical results using the experimental $r_S=2.07$ a.u.
+In Fig.
+\ref{F03}, we have good agreement between the present theoretical results and the measurements displayed in Table~\ref{table01}.
+Our theoretical approach also agrees with previous data by Sirotinin \cite{Sirotinin}, except for the lowest energy measurement at 80 keV.
+It is interesting that our full theoretical curve differs from SRIM-2013 for impact energies below 100 keV.
+We obtain a stopping maximum around $40 \times 10^{-15}$ eV cm$^2$/atom  at 65 keV.
+Instead, following the up-to-now only set of data \cite{Sirotinin}, SRIM-2013 suggests a lower stopping maximum  at impact energy of 115 kev.
+It is worth mentioning that similar theoretical results using the experimental $r_S=2.07$ a.u.
 instead of the theoretical $r_S=2.14$ a.u.
@@
 Additionally, we calculated values extracted from the theoretical framework that involved the relativistic wave functions and binding energies of Hf, and considered 4 electrons per atom in the free electron gas.
-The shell-wise local plasma approximation was employed to describe the energy transferred to the bound 1s-4f electrons, and two different models for the FEG: the screened potential with cusp condition (SPCC model) for energies below that of the plasmon excitation, and the Mermin-Lindhard dielectric formalism, for energies around the stopping maximum and above.
+The shell-wise local plasma approximation was employed to describe the energy transferred to the bound $1s$-$4f$ electrons, and two different models for the FEG: the screened potential with cusp condition (SPCC model) for energies below that of the plasmon excitation, and the Mermin-Lindhard dielectric formalism, for energies around the stopping maximum and above.
 Present theoretical stopping cross sections cover an extensive energy range from 1 keV/amu-10 meV/amu.
 
-%Our results have been confronted to both experimental values found in the literature and semi-empirical results. 
-At high impact energies, the new stopping  measurements are in good agreement with our theoretical results, and also with already published experimental data and semi-empirical values calculated by means of SRIM-2013 and ICRU-49.
+At high impact energies, the new stopping  measurements are in good agreement with our theoretical results, with previous experimental data and semi-empirical values by SRIM-2013 and ICRU-49.
 However, we call the attention that around the stopping maximum and at lower impact energies the difference between our full-theoretical results and SRIM is substantial.
-To our knowledge, these are the first theoretical calculations of stopping in Hf taking into account the atomic relativistic effect.
-%for energies around and below the stopping maximum, the present full-theoretical results clearly differ from semiempirical SRIM values, which are based on previous data. 
+To the best of our knowledge, these are the first theoretical calculations of stopping in Hf taking into account the atomic relativistic effect.
 Future experiments for impact energies below 100 keV would be important to complete this study.
@@
 \begin{acknowledgments}
-This work was funded by VRAC Grant Number L1-17 of Universidad Tecnol\'ogica Metropolitana, Chile; also by the Consejo Nacional de Investigaciones Científicas y Técnicas (CONICET), the Agencia Nacional de Promoción Científica y Tecnológica (ANPCyT), and Universidad de Buenos Aires (UBA), of Argentina.
+This work was funded by VRAC Grant Number L1-17 of Universidad Tecnol\'ogica Metropolitana, Chile; also by the Consejo Nacional de Investigaciones Cient\'ificas y T\'ecnicas (CONICET), the Agencia Nacional de Promoci\'on Cient\'ifica y Tecnol\'ogica (ANPCyT), and Universidad de Buenos Aires (UBA), from Argentina.
 
@@
 \bibitem{Chu01} W.K. Chu, J.W. Mayer, M.A. Nicolet (Eds.) Backscattering Spectrometry, Academic Press, New York, 1978.
-\bibitem{Sigmund} P. Sigmund, ''Particle Penetration and Radiation Effects. General Aspects and Stopping of Swift Point Charges``.
-Springer Series in Solid-State Sciences, Vol. 151, Springer, Berlin (2006).
+\bibitem{Sigmund} P.
+Sigmund, ``Particle Penetration and Radiation Effects.
+General Aspects and Stopping of Swift Point Charges''.
+Springer Series in Solid-State Sciences, Vol.
+151, Springer, Berlin (2006).
 \bibitem{Schardt} D. Schardt, T. Els\"asser, D. Schulz-Ertner, Rev. Mod. Phys. \textbf{82},  383-425 (2010).
@@
 \bibitem{Sirotinin} E.I. Sirotinin, A.F. Tulinov, V.A. Khodyrev, V.N. Mizgulin, Nucl. Instr. and Meth. B \textbf{4}, 337-345 (1984).
-\bibitem{Abril} I. Abril, M. Behar, R. Garc\'a Molina, R.C. Fadanelli, L.C.C.M. Nagamine, P.L. Grande, L. Sch\"unemann, C.D. Denton, N.R. Arista, E.B. Saitovich,
+\bibitem{Abril} I. Abril, M. Behar, R. Garcia Molina, R.C. Fadanelli, L.C.C.M. Nagamine, P.L. Grande, L. Sch\"unemann, C.D. Denton, N.R. Arista, E.B. Saitovich,
 Eur. Phys. J. D \text{54}, 65-70 (2009).
-\bibitem{Behar} M. Behar, R.C. Fadanelli, I. Abril, R. García-Molina, C.D. Denton, L.C.C.M. Nagamine, N.R. Arista, Phys. Rev. A \textbf{80},  062901 (2009).
-\bibitem{Primetzhofer} D. Primetzhofer, Nucl. Instr. and Meth. B \textbf{320}, 100-103 (2014) .
+\bibitem{Behar} M. Behar, R.C. Fadanelli, I. Abril, R. Garcia-Molina, C.D. Denton, L.C.C.M. Nagamine, N.R. Arista, Phys. Rev. A \textbf{80},  062901 (2009).
+\bibitem{Primetzhofer} D. Primetzhofer, Nucl. Instr. and Meth. B \textbf{320}, 100-103 (2014).
 %============================================================================================================================
 % Why Hafnium is so important:
-\bibitem{Choi} J.H. Choi, Y. Mao, J.P. Chang, Mat. Sci. Eng. R \textbf{72}, 97-136 (2011) .
-\bibitem{Robertson} J. Robertson, R.M. Wallace, Mat. Sci. Eng. R \textbf{88}, 1-41 (2015) .
+\bibitem{Choi} J.H. Choi, Y. Mao, J.P. Chang, Mat. Sci. Eng. R \textbf{72}, 97-136 (2011).
+\bibitem{Robertson} J. Robertson, R.M. Wallace, Mat. Sci. Eng. R \textbf{88}, 1-41 (2015).
 %============================================================================================================================
@@
 Primetzhofer, and P.
-Bauer1, Phys.
+Bauer, Phys.
 Rev.
-Lett \textbf{118}, 103401 (2017)
+Lett.
+\textbf{118}, 103401 (2017).
 
@@
 %Semi-empirical and Theoretical approach
-
-\bibitem{mendez2019} A.M.P. Mendez, C.C. Montanari, D.M. Mitnik, %\textit{Relativistic atomic structure calculations of heavy targets for inelastic collisions}, 
-Nucl. Instrum. Meth. B \textbf{460}, 114-118 (2019) .
-\bibitem{williams1995}
-G. Williams in http://xdb.lbl.gov/Section1/Sec\_1-1.html
 \bibitem{mon13} C.C. Montanari and J.E. Miraglia, Adv. Quant. Chem. \textbf{65}, edited by Dz. Belkic (Elsevier, Amsterdam, 2013), Chap. 7, pp. 165-201.
@@
 \bibitem{Mermin} N.D. Mermin, Phys. Rev. B \textbf{1}, 2362 (1970).
-\bibitem{ICRU49} ICRU report 49, Stopping Powers and Ranges for Protons and Alpha Particles, International Commission on
-Radiation Units and Measurements, 1993.
+%\bibitem{badnell97} N.R. Badnell, Comput. Phys. Commun., 182(7), 1528-1535 (2011). %https://doi.org/10.1016/j.cpc.2011.03.023
+\bibitem{mendez2019} A.M.P. Mendez, C.C. Montanari, D.M. Mitnik, Nucl. Instrum. Meth. B \textbf{460}, 114-118 (2019).
+\bibitem{ICRU49} ICRU report 49, Stopping Powers and Ranges for Protons and Alpha Particles, International Commission on Radiation Units and Measurements, 1993.
 \bibitem{Ziegler01} J.F. Ziegler, J.P. Biersack, SRIM2013, Computer Program and Manual. Available from www.srim.org.
+\bibitem{zenodo} 10.5281/zenodo.3678785 %10.5281/zenodo.3678760 
 %============================================================================================================================
@@
 \bibitem{Sun01} G. Sun, M. D\"{o}belli, A.M. M\"{u}ller, M. Stocker, M. Suter and L. Wacker, Nucl. Instr. and Meth. B \textbf{256}, 586-590 (2007).
-\bibitem{Raisanen01} J. Raisanen, U. Watjen, A.J.M. Plompen, F. Munnik, Nucl. Instr. and Meth. \textbf{B} 118 (1996) 1-6.
+\bibitem{Raisanen01} J. Raisanen, U. Watjen, A.J.M. Plompen, F. Munnik, Nucl. Instr. and Meth. \textbf{B} 118, 1-6  (1996).
 \bibitem{Schulz01} F. Schulz and J. Shchuchinzky, Nucl. Instr. and Meth. B \textbf{12},  90-94 (1985).
@@
 \bibitem{eche81} P. M. Echenique, R. M. Nieminen and R. H. Ritchie, Sol. State Comm. \textbf{37}, 779-781 (1981).
-\bibitem{nagy89} I. Nagy, A. Arnau, P. M. Echenique, and E. Zaremba, Phys. Rev. B \textbf{40}, 11986 (1989).
+\bibitem{nagy89} I. Nagy, A. Arnau, P. M. Echenique, and E. Zaremba, Phys. Rev. B \textbf{40}, 11983 (1989).
 % \bibitem{Vs94} J. E. Valdes, J. C. Eckardt, G. H. Lantschner, and N. R. Arista, Phys. Rev. A \textbf{49}, 1083 (1994).
@@
 \bibitem{suppression} C. C. Montanari, J. E. Miraglia, and N. R. Arista, Phys. Rev. A \textbf{62}, 052902 (2000).
+\bibitem{Hf_arxiv} A.M.P.
+Mendez, C.C.
+Montanari, D.M.
+Mitnik, \textit{Slater-type orbital expansion of neutral hafnium, numerical solution of the relativistic Dirac equation}, available soon in arXiv.org.
+\bibitem{badnell97} N.R. Badnell, Comput. Phys. Commun., 182(7), 1528-1535 (2011). %https://doi.org/10.1016/j.cpc.2011.03.023
+\bibitem{williams1995} G. Williams in http://xdb.lbl.gov/Section1/Sec\_1-1.html
 \bibitem{lindhard53} J. Lindhard, M. Scharff,  Mat. Fys. Medd. Dan. Vid. Selsk  \textbf{27}, 1 (1953).
@@
 \bibitem{mon09} C.C. Montanari, D.M. Mitnik, C.D. Archubi, and J.E. Miraglia, Phys. Rev. A \textbf{80}, 012901 (2009).
-
 \end{thebibliography}

### 2020-04-11 response.tex
include changes upon referees' comments

\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}
\setlength{\parindent}{0pt}
\usepackage{soul}
\usepackage{xcolor}

\setstcolor{red}

% Title Page
\title{Response to referee reports \\ AP11756/Montanari }
\author{C.C. Montanari, A. M. P. Mendez, D. M. Mitnik, J. E. Miraglia, \\
P.A. Miranda, R. Correa, J. Wachter, M. Aguilera, \\
E. Alves, N. Catarino and R.C. da Silva}
\date{\today}

\begin{document}
\maketitle

% ----------------------------------------------------------------------
\section{Reply to report of the First Referee}
% ----------------------------------------------------------------------

\textsl{1. Since only protons are considered as projectiles, the title 
needs to be modified accordingly.}

\vspace{0.1cm}
The title has been modified following the referee's recommendation to
``Stopping power of hafnium by proton impact and the importance of 
relativistic $4f$ electrons''.

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{2. It appears that calculations assume bare protons. According 
to CasP, neutrals play a considerable role in charge equilibrium around 
and below the Bragg peak. Some explanation is needed here.}

\vspace{0.1cm}
{\color{red}Reply here.}

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{3. Thickness measurements by Rutherford backscattering rely 
‘heavily’ on the scattering cross section, while energy loss is a minor 
correction depending on foil thickness (p.1, 2nd paragraph).}

\vspace{0.1cm}
{\color{red}Reply here.}

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{4. In order to get a more precise thickness value of the Hf 
foil, the authors use the energy loss of 5.486 MeV alphas as a standard, 
with SRIM as a reference. SRIM is based on empirical data, and for alpha 
particles in Hf, the IAEA database lists one single data point for He 
in Hf. Considering the spread in published stopping cross sections for 
e.g. He in Au, listed in the IAEA database, some explanation is needed 
to justify an error estimate of only 5\% (1st paragraph in section 
IIB).}

\vspace{0.1cm}
{\color{red}Reply here.}

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{5. In connection with the discussion of relativity and 
outer-shell electrons (p.3) it might be instructive to mention that 
this is the reason for the existence of relativistic quantum chemistry 
programs. Since the authors find it surprising that the effect increases 
in magnitude from inner to outer shells, the reader needs an 
explanation. I should think the effect of deviations in the screening 
add up shell by shell, but I may be wrong. Moreover, the sign of the 
discrepancy in figure 2a deserves attention, which seems to change from 
positive up to 4p- to negative from 5p+.}

\vspace{0.1cm}
The referee is correct. We included the following modifications in the text
(lines 200-214):

\vspace{0.1cm}
{\small ``One would expect the relativistic contributions to only affect 
significantly the inner shells. The relativistic binding energies of the 
electrons belonging to these shells are 2.5\% from the experimental values,
while the non-relativistic ones are within $\sim$10\%. Moreover, we notice 
the relativistic corrections greatly influences the outer $5p$ and $4f$ 
shells. This effect may be a consequence of the orthogonality with the 
relativistic inner shells. Also, the sign of the binding energy deviations 
is inverted near the FEG limit, i.e. the electrons go from being more 
bounded --than the experimental values-- to being less bounded. 
This change of sign can be attributed to the fact that the experimental 
values correspond to hafnium in solid-state, while our theoretical 
calculations correspond to the element in the gas phase.''}

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{6. The importance of 4f-5p screening is documented clearly 
(p.3), but the reader will ask why this does not affect other shells. 
I assume that the effect is important when two or more subshells have 
similar energies and, hence, overlap in real and velocity space. If so, 
I suggest to mention this. If not, another explanation is needed.}

\vspace{0.1cm}
{\color{red}Reply here.}

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{7. In view of the large contribution of the outermost shell, I 
suggest to include the core contribution in figure 3.}

\vspace{0.1cm}
{\color{red}Reply here.}

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{8. In addition to comparisons with experiments and empirical 
models, it would be instructive to see comparisons with theoretical 
models like CasP and DPASS in figure 4.}

\vspace{0.1cm}
{\color{red}Reply here.}

\newpage
% ----------------------------------------------------------------------
\section{Reply to report of the Second Referee}
% ----------------------------------------------------------------------

\textsl{1. Can the authors discuss how possible impurities in the sample and -especially- on both sides of the target would affect the final
stopping data? The authors must discuss it in the manuscript.}

\vspace{0.1cm}
{\color{red}Reply here.}

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{2. In the same context, the authors have to clarify whether potential
surface roughness and/or target non-uniformity might affect the final
results obtained in transmission method.}

\vspace{0.1cm}
{\color{red}Reply here.}

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{3. What is accuracy of the primary beam energy of the accelerator?
Moreover: when the target is placed and removed from the front of the
detector, does the detector itself need to unbiased and the chamber
opened? If so, might there be any offset ($\sim$keV) in the detection
system that would affect the energy distribution peaks, hence the
Gaussian position from the fits (Fig. 1)?}

\vspace{0.1cm}
{\color{red}Reply here.}

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{4. The experimental data above $\sim$1.5 MeV is systematically 
below the data from Ref. [7], hence SRIM semi-empirical approach, and 
also below the authors’ theory output. At these energies (and higher), 
the agreement between experiment and theory is expected to be better than
a few percent (that is why you used stopping power for 5.486 MeV alpha
on Hf to obtain the target thickness). Can the authors explain why
there is such systematic difference for the present measurements using
protons?}

\vspace{0.1cm}
{\color{red}Reply here.}

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{5. The authors claim an overall uncertainty of 5\%, but it is not clear
how they obtained this value, or whether only statistical
contributions were taken into account. Nevertheless, uncertainties of
measurements have to be treated following the Guide to the Expression
of Uncertainty in Measurement, International Organization for
Standardization, Geneva Switzerland, 1995. Especially for stopping
power measurements in transmission approach, there are other sources
of uncertainties that might potentially affect the final results
systematically (see comments 1, 2 and 3), and they must be discussed
in more details.}

\vspace{0.1cm}
{\color{red}Reply here.}

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{6. Can the present theory in the manuscript be compared to other
models (e.g., CasP and DPASS), especially towards low energies? Aiming
thus a constructive discussion on the overall-agreement among
different models to predict SCS of “challenge” materials, as
transition metals (f-subshell)?}

\vspace{0.1cm}
{\color{red}Reply here.}

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{7. In the last sentence of Sec. IV, the authors point out for a need
of more experimental data below 100 keV H+ but, in my opinion, more
data are needed even at higher energies, e.g., nearby (and slightly
above) the Bragg peak... Please enhance discussions about it.}

\vspace{0.1cm}
{\color{red}Reply here.}

\newpage
% ----------------------------------------------------------------------
\section{Reply to report of the Third Referee}
% ----------------------------------------------------------------------

\textsl{
- Regarding the theoretical model that describes the contribution to
the stopping power of the bound electrons, the authors emphasize the
importance of including relativistic effects in the description of the
binding energies and densities of the Hf 1s-4f states. However, it is
not clear if the calculations performed by the authors are done for
gas phase Hf atoms or for the solid Hf as in the experiments of Ref.
[38]. In case of being for the former, the slightly worse agreement
found for the 4f states as compared to the experimental values could
be related to this point.}

\vspace{0.1cm}
Our calculations are performed for gas phase Hf atoms. We have clarify
this in the manuscript (lines 211-214).

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{
1. The results from the three different theoretical calculations shown
in Fig. 3 are not completely clear. The authors emphasize the
importance of considering both screening effects and relativistic
effects for the 4f and 5p electrons, but it is not clear if later in
the text the authors use indistinctly the terminology ``relativistic''
and ``screening'' to refer to the same issue. For instance, are the ML
results shown in Fig. 3 calculated with the relativistic binding
energies and densities, but with and without including screening
corrections, or do the two ML curves differ in whether the
relativistic corrections are or not included? It is neither clear what
description of the bound electrons is used in the curve labeled as
``SPCC(FEG)+SLPA(bound)''.}

\vspace{0.1cm}
{\color{red}Reply here.}

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{
2. If all the results of Fig. 3 were calculated for the relativistic
description, it would be also interesting to show the results obtained
with the non-relativistic description in order to demonstrate the
importance of this correction.}

\vspace{0.1cm}
{\color{red}Te mando adjunto las funciones de onda y energías de ligaduras
del caso no relativista (NR). Habría que ver como queda.}

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{
3. Furthermore, it would also be meaningful to show the results from
the ML(FEG) calculation alone in order to identify better the
contribution of the bound electrons to the SP.}

\vspace{0.1cm}
{\color{red}Reply here.}

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{
4. When comparing the new and existing experimental data, clarify what
experimental technique was used in Ref. 7 to measure the stopping
power.}

\vspace{0.1cm}
{\color{red}Reply here.}

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{
5. The data by Sirotonin et al. plotted in Fig. 4 seem to suggest that
the stopping power maximum would be around 100-200 keV, while the
maximum predicted in the present work by the theoretical calculations
is located at slightly smaller energies. Please comment on this point
and discuss possible reasons for this discrepancy: What are the main
limitations in the theoretical model that can be causing this energy
shift?}

\vspace{0.1cm}
{\color{red}Reply here.}

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{
6. When reviewing on the experimental work performed on Hf oxides in
the introduction, note that there is a quite recent work by Roth and
co-workers published in Phys. Rev. Lett. 119, 163401 (2017), in which
the SP of protons is precisely measured in this target.}

\vspace{0.1cm}
We thank the referee for suggesting the missing reference. We have now
include it in the revised manuscript.

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{Finally, there are few misprints in the generated pdf file:}

\textsl{- Y-axis labels in Figs. 3 an 4 are unreadable }
{\color{red} ¿Cómo quedó el preprint? El pdf que genera mi latex tiene 
bien los labels, así que debe un problema del generador de latex del PRA...}

\st{\textsl{- Page 3,2nd column, 2nd full paragraph: "Too assess"}}

\st{\textsl{- Caption of Fig.2: "hollow circles" -$>$ "open circles" or
"unfilled circles"}}

\end{document}

### 2020-04-11 referees.tex
include changes upon referees' comments

----------------------------------------------------------------------
Report of the First Referee -- AP11756/Montanari
----------------------------------------------------------------------

The stopping power of hafnium and the importance of relativistic 4f electrons
Montanari et al, Phys. Rev. A

This manuscript continues previous theoretical work by members of this 
group on stopping of protons in heavy metals, including measurements in 
hafnium over an energy range ∼ 0.5 − 2MeV. This work is of interest not 
only in view of a notorious lack of experimental data in that part of 
the periodic table, but also because of an interesting shell structure, 
the significance of which is a central part of this study. While 
publication in PRA is indicated, I have a number of requests, questions 
and suggestions, which need consideration.

1. Since only protons are considered as projectiles, the title needs to 
be modified accordingly.

2. It appears that calculations assume bare protons. According to CasP, 
neutrals play a considerable role in charge equilibrium around and 
below the Bragg peak. Some explanation is needed here.

3. Thickness measurements by Rutherford backscattering rely ‘heavily’ 
on the scattering cross section, while energy loss is a minor correction 
depending on foil thickness (p.1, 2nd paragraph).

4. In order to get a more precise thickness value of the Hf foil, the 
authors use the energy loss of 5.486 MeV alphas as a standard, with 
SRIM as a reference. SRIM is based on empirical data, and for alpha 
particles in Hf, the IAEA database lists one single data point for He 
in Hf. Considering the spread in published stopping cross sections for 
e.g. He in Au, listed in the IAEA database, some explanation is needed 
to justify an error estimate of only 5\% (1st paragraph in section IIB).

5. In connection with the discussion of relativity and outer-shell 
electrons (p.3) it might be instructive to mention that this is the 
reason for the existence of relativistic quantum chemistry programs. 
Since the authors find it surprising that the effect increases in 
magnitude from inner to outer shells, the reader needs an explanation. 
I should think the effect of deviations in the screening add up shell 
by shell, but I may be wrong. Moreover, the sign of the discrepancy in 
figure 2a deserves attention, which seems to change from positive up to 
4p- to negative from 5p+.

6. The importance of 4f-5p screening is documented clearly (p.3), but 
the reader will ask why this does not affect other shells. I assume 
that the effect is important when two or more subshells have similar 
energies and, hence, overlap in real and velocity space. If so, I 
suggest to mention this. If not, another explanation is needed.

7. In view of the large contribution of the outermost shell, I suggest 
to include the core contribution in figure 3.

8. In addition to comparisons with experiments and empirical models, 
it would be instructive to see comparisons with theoretical models like 
CasP and DPASS in figure 4.

----------------------------------------------------------------------
Report of the Second Referee -- AP11756/Montanari
----------------------------------------------------------------------

C. C. Montanari and co-authors report experimental and theoretical
stopping cross section data of protons in hafnium. Experimental data
are obtained from transmission experiments in the energy interval of
[600 – 2500] keV, whereas theoretical results are based on a
combination of different approaches: shell-wise local plasma
approximation model (SLPA - from the same author) for higher impact
ion velocities (around and above Bragg peak) and two free electron gas
models (FEG) for low impact velocities (below the Bragg peak). The
study is thorough and the manuscript well organized, emphasizing on
the importance of the material under analysis. The experimental and
theoretical approaches are scientifically robust. However, for the
experimental data, the values are obtained only for higher energies (>
1 MeV/u), and no significant new physics is expected, as in this
energy regime most of the well-known semi-empirical models and ab
initio calculations work satisfactorily well. This lack of
experimental evidence (at low energies) to verify the reliability of
the present model is indeed the major drawback of the present work.

Nevertheless, the core of the present work (i.e. a theoretical
approach capable to predict stopping power from low to high energies,
over several orders of magnitude in ion energy describing the Bragg
peak) is of broad enough interest for the readership of Physical
Review A. In this context, I recommend the manuscript to be accepted    
after minor revisions, especially to address more attention (and more
comments) to the aforementioned weakness of the manuscript.

I also request the authors to address attention to some comments to
further enhance the quality of the manuscript:

1. Can the authors discuss how possible impurities in the sample and –
especially - on both sides of the target would affect the final
stopping data? The authors must discuss it in the manuscript.

2. In the same context, the authors have to clarify whether potential
surface roughness and/or target non-uniformity might affect the final
results obtained in transmission method.

3. What is accuracy of the primary beam energy of the accelerator?
Moreover: when the target is placed and removed from the front of the
detector, does the detector itself need to unbiased and the chamber
opened? If so, might there be any offset (~keV) in the detection
system that would affect the energy distribution peaks, hence the
Gaussian position from the fits (Fig. 1)?

4. The experimental data above ~1.5 MeV is systematically below the
data from Ref. [7], hence SRIM semi-empirical approach, and also below
the authors’ theory output. At these energies (and higher), the
agreement between experiment and theory is expected to be better than
a few percent (that is why you used stopping power for 5.486 MeV alpha
on Hf to obtain the target thickness). Can the authors explain why
there is such systematic difference for the present measurements using
protons?

5. The authors claim an overall uncertainty of 5%, but it is not clear
how they obtained this value, or whether only statistical
contributions were taken into account. Nevertheless, uncertainties of
measurements have to be treated following the Guide to the Expression
of Uncertainty in Measurement, International Organization for
Standardization, Geneva Switzerland, 1995. Especially for stopping
power measurements in transmission approach, there are other sources
of uncertainties that might potentially affect the final results
systematically (see comments 1, 2 and 3), and they must be discussed
in more details.

6. Can the present theory in the manuscript be compared to other
models (e.g., CasP and DPASS), especially towards low energies? Aiming
thus a constructive discussion on the overall-agreement among
different models to predict SCS of “challenge” materials, as
transition metals (f-subshell)?

7. In the last sentence of Sec. IV, the authors point out for a need
of more experimental data below 100 keV H+ but, in my opinion, more
data are needed even at higher energies, e.g., nearby (and slightly
above) the Bragg peak... Please enhance discussions about it.

----------------------------------------------------------------------
Report of the Third Referee -- AP11756/Montanari
----------------------------------------------------------------------

Montarani and co-workers perform a joined experimental and theoretical
study of the stopping power (SP) of protons in Hafnium thin films. In
my opinion, the contribution of the present study is two-fold. Despite
its fundamental interest, experimental data on the SP of particles on
metals with large atomic number are scarce. Hence, the new
measurements reported in this work for 0.6-2.5 MeV protons in Hf,
which is characterized by occupied 4f and 5p shells, are valuable per
se. Moreover, the authors also provide new theoretical calculations of
the SP, in which they combine different models in order to cover an
ample range of proton energies. In particular, the theoretical study
mainly focuses on understanding the role of the excitation of 4f-5p
electrons in the SP. Overall, I consider this work is valuable to
extend our knowledge on the interaction of particles (protons in this
case) with heavy metals and, as such, it will be of interest for the
ion-solid interaction community.

In general the manuscript is well written, but there are some points
regarding the employed theoretical approximations that are a bit
confusing. Therefore, my recommendation is publishing in Physical
Review A once the authors clarify the following points:

- Regarding the theoretical model that describes the contribution to
the stopping power of the bound electrons, the authors emphasize the
importance of including relativistic effects in the description of the
binding energies and densities of the Hf 1s-4f states. However, it is
not clear if the calculations performed by the authors are done for
gas phase Hf atoms or for the solid Hf as in the experiments of Ref.
[38]. In case of being for the former, the slightly worse agreement
found for the 4f states as compared to the experimental values could
be related to this point.

- The results from the three different theoretical calculations shown
in Fig. 3 are not completely clear. The authors emphasize the
importance of considering both screening effects and relativistic
effects for the 4f and 5p electrons, but it is not clear if later in
the text the authors use indistinctly the terminology 'relativistic'
and 'screening' to refer to the same issue. For instance, are the ML
results shown in Fig.3 calculated with the relativistic binding
energies and densities, but with and without including screening
corrections, or do the two ML curves differ in whether the
relativistic corrections are or not included? It is neither clear what
description of the bound electrons is used in the curve labeled as
'SPCC(FEG)+SLPA(bound)'.

- If all the results of Fig. 3 were calculated for the relativistic
description, it would be also interesting to show the results obtained
with the non-relativistic description in order to demonstrate the
importance of this correction.

- Furthermore, it would also be meaningful to show the results from
the ML(FEG) calculation alone in order to identify better the
contribution of the bound electrons to the SP.

- When comparing the new and existing experimental data, clarify what
experimental technique was used in Ref.7 to measure the stopping
power.

- The data by Sirotonin et al. plotted in Fig.4 seem to suggest that
the stopping power maximum would be around 100-200 keV, while the
maximum predicted in the present work by the theoretical calculations
is located at slightly smaller energies. Please comment on this point
and discuss possible reasons for this discrepancy: What are the main
limitations in the theoretical model that can be causing this energy
shift?

- When reviewing on the experimental work performed on Hf oxides in
the introduction, note that there is a quite recent work by Roth and
co-workers published in Phys. Rev. Lett. 119, 163401 (2017), in which
the SP of protons is precisely measured in this target.

Finally, there are few misprints in the generated pdf file:

- Y-axis labels in Figs. 3 an 4 are unreadable

- Page 3,2nd column, 2nd full paragraph: "Too assess"

- Caption of Fig.2: "hollow circles" -> "open circles" or "unfilled
circles"

### 2020-04-11 sp_hafnium.tex
include changes upon referees' comments

--- 2020-03-11 Paper_SPHfP_feb2020_prb.tex
@@
-%%   This file is part of the APS files in the REVTeX 4 distribution.
-%%   Version 4.1r of REVTeX, August 2010
-%%
-%%
-%%   Copyright (c) 2001, 2009, 2010 The American Physical Society.
-%%
-%
-\documentclass[aps,pra,reprint,groupedaddress]{revtex4-1}
-
-%\documentclass[aps,pra,twocolumn,groupedaddress]{revtex4-1}
-%\documentclass[aps,prb,reprint,superscriptaddress]{revtex4-1}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass[aps,pra,reprint,groupedaddress,showpacs,showkeys]{revtex4-1}
 \usepackage{amssymb}
@@
 \usepackage{graphicx}
+\usepackage{soul}
+\usepackage{xcolor}
+\setstcolor{red}
 \usepackage[mathlines]{lineno}% Enable numbering of text and display math
@@
 %Title of paper
-\title{The stopping power of hafnium and the importance of relativistic $4f$ electrons}
+\title{Stopping power of hafnium by proton impact and \\
+the importance of relativistic $4f$ electrons}
 %Authors and affiliations
@@
 \email{mclaudia@iafe.uba.ar}
-\affiliation{Instituto de
-Astronom\'{\i}a y F\'{\i}sica del Espacio, Consejo Nacional de Investigaciones Cient\'{\i}ficas y
-T\'{e}cnicas - Universidad de Buenos Aires, Pabell\'on IAFE, 1428 Buenos Aires, Argentina}%
-%\affiliation{Universidad de Tres de Febrero, Sáenz Peña, Buenos Aires, Argentina.}
+\affiliation{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio, 
+Consejo Nacional de Investigaciones Cient\'{\i}ficas y T\'{e}cnicas - 
+Universidad de Buenos Aires, Pabell\'on IAFE, 1428 Buenos Aires, 
+Argentina}
 %------------------------------
 \author{A. M. P. Mendez}
-\affiliation{Instituto de
-Astronom\'{\i}a y F\'{\i}sica del Espacio, Consejo Nacional de Investigaciones Cient\'{\i}ficas y
-T\'{e}cnicas - Universidad de Buenos Aires, Pabell\'on IAFE, 1428 Buenos Aires, Argentina}%
+\affiliation{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio, 
+Consejo Nacional de Investigaciones Cient\'{\i}ficas y T\'{e}cnicas - 
+Universidad de Buenos Aires, Pabell\'on IAFE, 1428 Buenos Aires, 
+Argentina}
 %-------------------------------
 \author{D. M. Mitnik}
-\affiliation{Instituto de
-Astronom\'{\i}a y F\'{\i}sica del Espacio, Consejo Nacional de Investigaciones Cient\'{\i}ficas y
-T\'{e}cnicas - Universidad de Buenos Aires, Pabell\'on IAFE, 1428 Buenos Aires, Argentina}%
-%\affiliation{Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales, Departamento de F\'{\i}sica, Ciudad Universitaria, 1428 Buenos Aires, Argentina.}
+\affiliation{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio, 
+Consejo Nacional de Investigaciones Cient\'{\i}ficas y T\'{e}cnicas - 
+Universidad de Buenos Aires, Pabell\'on IAFE, 1428 Buenos Aires, 
+Argentina}
 %----------------------------------------------
 \author{J. E. Miraglia}
-\affiliation{Instituto de
-Astronom\'{\i}a y F\'{\i}sica del Espacio, Consejo Nacional de Investigaciones Cient\'{\i}ficas y
-T\'{e}cnicas - Universidad de Buenos Aires, Pabell\'on IAFE, 1428 Buenos Aires, Argentina}%
+\affiliation{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio, 
+Consejo Nacional de Investigaciones Cient\'{\i}ficas y T\'{e}cnicas - 
+Universidad de Buenos Aires, Pabell\'on IAFE, 1428 Buenos Aires, 
+Argentina}
 %---------------------------Chile
@@
 \author{P.A. Miranda}
-\affiliation{Departmento de F\'isica, Facultad de Ciencias Naturales, Matem\'atica y del Medio Ambiente. Universidad Tecnol\'ogica Metropolitana. 7800002, Chile}
+\affiliation{Departamento de F\'isica, Facultad de Ciencias Naturales, 
+Matem\'atica y del Medio Ambiente. Universidad Tecnol\'ogica Metropolitana. 
+7800002, Chile}
 %----------------------------
 \author{R. Correa}
-\affiliation{Departmento de F\'isica, Facultad de Ciencias Naturales, Matem\'atica y del Medio Ambiente. Universidad Tecnol\'ogica Metropolitana. 7800002, Chile}
+\affiliation{Departamento de F\'isica, Facultad de Ciencias Naturales, 
+Matem\'atica y del Medio Ambiente. Universidad Tecnol\'ogica Metropolitana. 
+7800002, Chile}
+%----------------------------
 \author{J. Wachter}
-\affiliation{Departmento de F\'isica, Facultad de Ciencias Naturales, Matem\'atica y del Medio Ambiente. Universidad Tecnol\'ogica Metropolitana. 7800002, Chile}
+\affiliation{Departamento de F\'isica, Facultad de Ciencias Naturales, 
+Matem\'atica y del Medio Ambiente. Universidad Tecnol\'ogica Metropolitana. 
+7800002, Chile}
+%----------------------------
 \author{M. Aguilera}
-\affiliation{Departmento de F\'isica, Facultad de Ciencias Naturales, Matem\'atica y del Medio Ambiente. Universidad Tecnol\'ogica Metropolitana. 7800002, Chile}
+\affiliation{Departamento de F\'isica, Facultad de Ciencias Naturales, 
+Matem\'atica y del Medio Ambiente. Universidad Tecnol\'ogica Metropolitana. 
+7800002, Chile}
 %-------------------------Lisboa
 \author{E. Alves}
-\affiliation{Centro de Ci\^{e}ncias e Tecnologias Nucleares, Instituto Superior T\'ecnico, Universidade de Lisboa,
-2696-953 Sacav\'{e}m, Portugal}
-\affiliation{Instituto de Plasma e Fus\~{a}o Nuclear, Instituto Superior T\'ecnico, Universidade de Lisboa, 2696-953 Sacav\'{e}m, Portugal}
+\affiliation{Centro de Ci\^{e}ncias e Tecnologias Nucleares, Instituto 
+Superior T\'ecnico, Universidade de Lisboa, 2696-953 Sacav\'{e}m, 
+Portugal}
+\affiliation{Instituto de Plasma e Fus\~{a}o Nuclear, Instituto Superior 
+T\'ecnico, Universidade de Lisboa, 2696-953 Sacav\'{e}m, Portugal}
 %---------------------------
 \author{N. Catarino}
-\affiliation{Centro de Ci\^{e}ncias e Tecnologias Nucleares, Instituto Superior T\'ecnico, Universidade de Lisboa,
-2696-953 Sacav\'{e}m, Portugal}
+\affiliation{Centro de Ci\^{e}ncias e Tecnologias Nucleares, Instituto 
+Superior T\'ecnico, Universidade de Lisboa, 2696-953 Sacav\'{e}m, 
+Portugal}
 %---------------------------
 \author{R.C. da Silva}
-\affiliation{Centro de Ci\^{e}ncias e Tecnologias Nucleares, Instituto Superior T\'ecnico, Universidade de Lisboa,
-2696-953 Sacav\'{e}m, Portugal}
+\affiliation{Centro de Ci\^{e}ncias e Tecnologias Nucleares, Instituto 
+Superior T\'ecnico, Universidade de Lisboa, 2696-953 Sacav\'{e}m, 
+Portugal}
 
@@
 
-%=============================================================================================================================================================================================
-%=============================================================================================================================================================================================
+%=======================================================================
+%=======================================================================
 \begin{abstract}
- The stopping power of protons through Hf foil has been studied both experimentally and theoretically.
-The measurements were performed at the Laboratory of Accelerators and X-Ray Diffraction in Lisbon, by using the transmission method on self-supporting stopping material.
-%for (0.6-2.5) MeV protons.
- The overall uncertainty of around $5\%$ was established over the protons energy range  (0.6-2.5) MeV.
-The theoretical developments involved fully relativistic atomic structure calculations for %this Z=72 target
- Hf, which required the solution of the Dirac equation.
-The shell-wise local plasma approximation (SLPA) was used to describe the energy transferred to the bound $1s$-$4f$ electrons, and the outer four electrons were considered as a free electron gas (FEG).
-%We contemplated two different models for the FEG: the screened potential with cusp condition (SPCC) for energies below that of the plasmon excitation, and the dielectric formalism, for energies around the stopping maximum and above.
- We found the relativistic description of the $4f$-shell and the screening between $4f$ and $5p$ electrons are decisive around the stopping maximum. %for energies above 40 keV.
- Present theoretical and experimental results are in very good agreement in the energy region of the new measurements. However, the present theoretical stopping cross sections show %interesting 
- substantial differences with the most used semi-empirical models (SRIM2013 and ICRU-49) at intermediate to low energies.
-Our calculations suggest the stopping maximum %around 40 eV$/(10^{15}$ atom/cm$^2)$ at 70 keV, which is
- higher and shifted to lower energies than these previous predictions. Future measurements below 100 keV would be decisive for a better knowledge  around the maximum and below.
+ The stopping power of protons through Hf foil has been studied both 
+experimentally and theoretically. The measurements were performed at 
+the Laboratory of Accelerators and X-Ray Diffraction in Lisbon, by 
+using the transmission method on self-supporting stopping material. 
+ The overall uncertainty of around $5\%$ was established over the 
+protons energy range  (0.6-2.5) MeV. The theoretical developments 
+involved fully relativistic atomic structure calculations for Hf, which 
+required the solution of the Dirac equation. The shell-wise local plasma 
+approximation (SLPA) was used to describe the energy transferred to the 
+bound $1s$-$4f$ electrons, and the outer four electrons were considered 
+as a free electron gas (FEG). We found the relativistic description of 
+the $4f$-shell and the screening between $4f$ and $5p$ electrons are 
+decisive around the stopping maximum. Present theoretical and 
+experimental results are in very good agreement in the energy region 
+of the new measurements. However, the present theoretical stopping 
+cross sections show substantial differences with the most used 
+semi-empirical models (SRIM2013 and ICRU-49) at intermediate to low 
+energies. Our calculations suggest the stopping maximum higher and 
+shifted to lower energies than these previous predictions. Future 
+measurements below 100 keV would be decisive for a better knowledge 
+around the maximum and below.
 \end{abstract}
@@
 % insert suggested PACS numbers in braces on next line
-\pacs{34.50.Bw} \keywords{Stopping power, hafnium, Relativistic atomic structure, Transmission method}
+\pacs{34.50.Bw} \keywords{Stopping power, hafnium, Relativistic atomic 
+structure, Transmission method}
 
@@
 
-%\linenumbers
-%=============================================================================================================================================================================================
-%=============================================================================================================================================================================================
+\linenumbers
+%=======================================================================
+%=======================================================================
 \section{Introduction}
@@
 
-For impact energies above a few keV/amu, mono-energetic charged particles penetrating a foil of any material lose their energy through a series of consecutive inelastic collisions, mainly with target electrons \cite{Chu01,Sigmund}.
-The information given by the energy loss process is essential not only to have a better knowledge of the physics behind the fundamental interactions, but also because it plays an important role in many applied fields such as materials science, nuclear physics, ionic implantation and radiotherapy \cite{Sigmund,Schardt}.
-Experimental data on ion mean energy loss per unit path $S(E)$ is of crucial relevance to check the reliability of semi-empirical models and to determine some key parameters
-%included in theoretical approaches of stopping power %\cite{Diwan,Damache01,Damache02,Damache03,Damache04}. 
-\cite{Diwan,Damache04,Damache02}. The experimental data available is often rather scarce, which is %particularly 
-troublesome when the material under study corresponds to an element of low occurrence on the Earth's upper crust, such as hafnium.
+For impact energies above a few keV/amu, mono-energetic charged 
+particles penetrating a foil of any material lose their energy through 
+a series of consecutive inelastic collisions, mainly with target 
+electrons~\cite{Chu01,Sigmund}. The information given by the energy 
+loss process is essential not only to have a better knowledge of the 
+physics behind the fundamental interactions, but also because it plays 
+an important role in many applied fields such as materials science, 
+nuclear physics, ionic implantation and radiotherapy~\cite{Sigmund,Schardt}. 
+Experimental data on ion mean energy loss per unit path $S(E)$ is of 
+crucial relevance to check the reliability of semi-empirical models and 
+to determine some key parameters~\cite{Diwan,Damache04,Damache02}. The 
+experimental data available is often rather scarce, which is troublesome 
+when the material under study corresponds to an element of low 
+occurrence on the Earth's upper crust, such as hafnium.
 
-So far, only one experimental work has been published regarding the stopping power cross section of pure hafnium for protons \cite{Sirotinin}, while more attention has been recently given to studies involving hafnium oxide due to its practical use \cite{Abril,Behar,Primetzhofer}.
-It is well known that significant attention has been paid in recent years to transition metal-oxides such as HfO$_2$ because of their potential as alternative gate dielectrics to replace SiO$_2$ for the future generation of nano-electronics with less than 45 nm gate length \cite{Choi,Robertson}.
-Some important physical properties of the above mentioned metal-oxide films depend on its thickness, which is often measured by using Rutherford Backscattering Spectrometry \cite{Alfassi01, Tesmer01}, a method that relies heavily on the determination of the stopping power of ion beams in the material of interest.
+So far, only one experimental work has been published regarding the 
+stopping power cross section of pure hafnium for protons~\cite{Sirotinin}, 
+while more attention has been recently given to studies involving 
+hafnium oxide due to its practical use~\cite{Abril,Behar,Primetzhofer,Roth}. 
+It is well known that significant attention has been paid in recent 
+years to transition metal-oxides such as HfO$_2$ because of their 
+potential as alternative gate dielectrics to replace SiO$_2$ for the 
+future generation of nano-electronics with less than 45 nm gate 
+length~\cite{Choi,Robertson}. Some important physical properties of the 
+above mentioned metal-oxide films depend on its thickness, which is 
+often measured by using Rutherford Backscattering Spectrometry~\cite{Alfassi01,Tesmer01}, 
+a method that relies heavily on the determination of the stopping power 
+of ion beams in the material of interest.
 
-In this study, we report experimental stopping power cross sections over the incident energy range (0.6-2.5) MeV for protons crossing self-supported Hf thin-film by using the transmission method.
-We aim not only to upgrade stopping power data compilations \cite{HPaul03,mondim17} but also to provide useful information about the processes governing the slowing down of protons in multi-electronic targets.
-In the rare earth metals, the $4f$ electrons play a relevant role in the stopping power, being the first shell of bound electrons below the conduction band.
-As already noted \cite{Roth17}, the free electron gas (FEG) shows unexpected %a peculiar
-behavior in these elements, which casts %that states 
-doubts on its proper description. In the case of Hf, we found the contribution of the $4f$-shell decisive even at impact energies around the stopping maximum, as shown later.
+In this study, we report experimental stopping power cross sections 
+over the incident energy range (0.6-2.5) MeV for protons crossing 
+self-supported Hf thin-film by using the transmission method. We aim 
+not only to upgrade stopping power data compilations~\cite{HPaul03,mondim17} 
+but also to provide useful information about the processes governing 
+the slowing down of protons in multi-electronic targets. In the rare 
+earth metals, the $4f$ electrons play a relevant role in the stopping 
+power, being the first shell of bound electrons below the conduction 
+band. As already noted \cite{Roth17}, the free electron gas (FEG) shows 
+unexpected behavior in these elements, which casts doubts on its proper 
+description. In the case of Hf, we found the contribution of the 
+$4f$-shell decisive even at impact energies around the stopping maximum, 
+as shown later.
 
-The theoretical approach implemented in this work uses the shell-wise local plasma approximation (SLPA) \cite{mon13} to describe the energy transferred to the bound $1s$-$4f$ electrons, and two different models for the FEG: the screened potential with cusp condition model (SPCC) \cite{mon17}, which is non-linear binary formalism%,  for energies below that of the plasmon excitation
-; and the Mermin-Lindhard dielectric formalism (ML) \cite{Mermin}, for energies around the stopping maximum and above. Our model requires 
-the relativistic wave functions and binding energies of Hf, and considers 4 electrons per atom in the  FEG \cite{mendez2019}. 
-%On the other hand,
-Hf has the extra interest of the filled $4f$-subshell (with 14 electrons) as the main contribution below the FEG, causing the stopping cross sections to be very sensitive to a good description of this shell.
-The screening among the $4f$ and $5p$ electrons has been considered and found to play a major role within the SLPA calculations.
+The theoretical approach implemented in this work uses the shell-wise 
+local plasma approximation (SLPA)~\cite{mon13} to describe the energy 
+transferred to the bound $1s$-$4f$ electrons, and two different models 
+for the FEG: the screened potential with cusp condition model (SPCC)~\cite{mon17}, 
+which is non-linear binary formalism; and the Mermin-Lindhard dielectric 
+formalism (ML)~\cite{Mermin}, for energies around the stopping maximum 
+and above. Our model requires the relativistic wave functions and 
+binding energies of Hf, and considers 4 electrons per atom in the 
+FEG~\cite{mendez2019}. Hf has the extra interest of the filled 
+$4f$-subshell (with 14 electrons) as the main contribution below the 
+FEG, causing the stopping cross sections to be very sensitive to a good 
+description of this shell. The screening among the $4f$ and $5p$ 
+electrons has been considered and found to play a major role within the 
+SLPA calculations.
 
-The experimental details and data are given in Section \ref{experiment}, while the theoretical method is explained in Section \ref{theory}.
-Present theoretical and experimental values are finally compared to the available experimental values \cite{Sirotinin},  the ICRU-49 calculations \cite{ICRU49} and the SRIM-2013 package \cite{Ziegler01}.
-Conclusions and discussions are given in Section \ref{conclusion}.
-All the present data can be found at the Zenodo platform \cite{zenodo}.
+The experimental details and data are given in Section~\ref{experiment}, 
+while the theoretical method is explained in Section~\ref{theory}. 
+Present theoretical and experimental values are finally compared to the 
+available experimental values~\cite{Sirotinin}, the ICRU-49 
+calculations~\cite{ICRU49} and the SRIM-2013 package~\cite{Ziegler01}. 
+Conclusions and discussions are given in Section~\ref{conclusion}. All 
+the present data can be found at the Zenodo platform \cite{zenodo}.
 
-%=============================================================================================================================================================================================
-
+%-----------------------------------------------------------------------
 \begin{figure}[!t]
@@
 \includegraphics[width=9cm]{Fig01.eps}
-\caption{RBS spectrum for $E_{\mathrm{avg}} = 921.1$ keV protons on hafnium sample which is subsequently used to determine the energy loss in the foil.}
+\caption{RBS spectrum for $E_{\mathrm{avg}}=921.1$ keV protons on 
+hafnium sample which is subsequently used to determine the energy loss 
+in the foil.}
 \label{F01}
@@
 
-%=============================================================================================================================================================================================
-
+%=======================================================================
 \section{EXPERIMENTAL ARRANGEMENTS}
 \label{experiment}
+
+%-----------------------------------------------------------------------
 \subsection{Accelerator and scattering Chamber}
-The procedure used in this work to obtain stopping power data is essentially the same as described in Ref. \cite{Miranda01}. The present measurements were made at the IST/LATR
-(Laboratory of Accelerators and X-Ray Diffraction) in Lisbon.
-This facility uses a 2.5 MV Van de Graaff accelerator to deliver $^1$H$^+$ primary ion beams through a series of electrostatic lenses and collimators onto a thin Au/SiO$_2$ sample which is used as a scattering center.
-This sample was placed in the center of a RBS/C scattering chamber, where a high vacuum (pressure of
-$\sim$10$^{-6}$ Torr) was maintained during the measurements. The beam current on the sample was kept at around 5.0 nA to attain sufficient statistics in each particle spectrum.
-By using a beam spot of about 1.0 mm in diameter, a solid angle of 11.4 msr was attained.
-The overall energy resolution (FWHM) of the detection system was about 15 keV relative to 5.486 MeV alpha particles from a $^{241}$Am source.
+The procedure used in this work to obtain stopping power data is 
+essentially the same as described in Ref.~\cite{Miranda01}. The present 
+measurements were made at the IST/LATR (Laboratory of Accelerators and 
+X-Ray Diffraction) in Lisbon. This facility uses a 2.5 MV Van de Graaff 
+accelerator to deliver $^1$H$^+$ primary ion beams through a series of 
+electrostatic lenses and collimators onto a thin Au/SiO$_2$ sample which 
+is used as a scattering center. This sample was placed in the center of 
+a RBS/C scattering chamber, where a high vacuum (pressure of 
+$\sim$10$^{-6}$ Torr) was maintained during the measurements. The beam 
+current on the sample was kept at around 5.0 nA to attain sufficient 
+statistics in each particle spectrum. By using a beam spot of about 
+1.0 mm in diameter, a solid angle of 11.4 msr was attained. The overall 
+energy resolution (FWHM) of the detection system was about 15 keV 
+relative to 5.486 MeV alpha particles from a $^{241}$Am source.
 
-%---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
+%-----------------------------------------------------------------------
 \subsection{Target}
-The stopping material under analysis was a hafnium foil with nominal thickness of 1.0 $\mu$m and 99.95\% purity which was supplied by Lebow Company \cite{Lebow}.
-However, a more precise thickness value was achieved by measuring the energy loss of alpha particles coming from a calibrated ($^{239}$Pu, $^{241}$Am, $^{244}$Cm) source.
-From the alpha spectra with and without the Hf foil interposed, the characteristic energy shift $\delta$E was measured and then combined with the stopping power for 5.486 MeV alphas on hafnium (55.69 eV/$10^{15}$ at/cm$^2$) found in Ref.\cite{Ziegler01} to obtain an areal density of ($4.13 \pm 0.21$)$\times 10^{19}$ at/cm$^2$ which corresponds to a thickness of $0.920\pm0.046 \mu$m.
-%---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
+The stopping material under analysis was a hafnium foil with nominal 
+thickness of 1.0 $\mu$m and 99.95\% purity which was supplied by Lebow 
+Company~\cite{Lebow}. However, a more precise thickness value was 
+achieved by measuring the energy loss of alpha particles coming from a 
+calibrated ($^{239}$Pu, $^{241}$Am, $^{244}$Cm) source. From the alpha 
+spectra with and without the Hf foil interposed, the characteristic 
+energy shift $\delta$E was measured and then combined with the stopping 
+power for 5.486 MeV alphas on hafnium (55.69 eV/$10^{15}$ at/cm$^2$) 
+found in Ref.~\cite{Ziegler01} to obtain an areal density of 
+($4.13 \pm 0.21$)$\times 10^{19}$ at/cm$^2$ which corresponds to a 
+thickness of $0.920\pm0.046 \mu$m.
+
+%-----------------------------------------------------------------------
 \subsection{Energy loss measurement}
-Once the beam impinges on the Au/SiO$_2$ sample, protons are back-scattered towards a Si surface barrier detector located at 140$^{\circ}$ relative to the initial beam direction.
-Fig.\ref{F01} shows two particle spectra where ion energies $E_2$ and $E_1$ are associated to a placed and removed hafnium sample, respectively.
-Both energy distributions were fitted by Gaussian functions to obtain the mean energy and width (FWHM) of the peaks \cite{Sun01}, and from the difference between these two peak positions in the spectrum, the total energy loss $\Delta E = (E_1 - E_2)$ in the foil was calculated.
-As established in previous studies \cite{Miranda01, Damache02}, the experimental stopping power cross sections $\varepsilon (E) $ are determined at some mean energy $E_{\mathrm{avg}}$ by measuring the ion energy losses $\Delta E$ within the investigated Hf foil, which has a mean thickness denoted by $\Delta x$.
-In this way, only when the energy loss fraction $\Delta E/E_{\mathrm{avg}}$ across the Hf foil is not exceeding 20\%
-it is possible to define the stopping cross section by \cite{Raisanen01,Schulz01}:
+Once the beam impinges on the Au/SiO$_2$ sample, protons are 
+back-scattered towards a Si surface barrier detector located at 
+140$^{\circ}$ relative to the initial beam direction. Fig.~\ref{F01} 
+shows two particle spectra where ion energies $E_2$ and $E_1$ are 
+associated to a placed and removed hafnium sample, respectively. Both 
+energy distributions were fitted by Gaussian functions to obtain the 
+mean energy and width (FWHM) of the peaks \cite{Sun01}, and from the 
+difference between these two peak positions in the spectrum, the total 
+energy loss $\Delta E = (E_1 - E_2)$ in the foil was calculated. As 
+established in previous studies \cite{Miranda01,Damache02}, the 
+experimental stopping power cross sections $\varepsilon (E) $ are 
+determined at some mean energy $E_{\mathrm{avg}}$ by measuring the ion 
+energy losses $\Delta E$ within the investigated Hf foil, which has a 
+mean thickness denoted by $\Delta x$. In this way, only when the energy 
+loss fraction $\Delta E/E_{\mathrm{avg}}$ across the Hf foil is not 
+exceeding 20\% it is possible to define the stopping cross section by 
+\cite{Raisanen01,Schulz01}:
 \begin{equation}\label{eq:stcross}
- \varepsilon (E) = \frac{S(E)}{N} = -\frac{dE}{N dx} \approx -\frac{\Delta E}{N \Delta x},
+ \varepsilon(E)=\frac{S(E)}{N}=-\frac{dE}{N\,dx}\approx-\frac{\Delta E}{N\Delta x},
 \end{equation}
-where $N$ denotes the atomic number density (atoms cm$^{-3}$) of the material under study.
-When this condition was not fulfilled, a small correction to the mean energies $E_{\mathrm{avg}}$ was applied in order to account for the nonlinear dependence on ion energy of stopping powers \cite{Chilton,Rajatora}.
+where $N$ denotes the atomic number density (atoms cm$^{-3}$) of the 
+material under study. When this condition was not fulfilled, a small 
+correction to the mean energies $E_{\mathrm{avg}}$ was applied in order 
+to account for the nonlinear dependence on ion energy of stopping 
+powers \cite{Chilton,Rajatora}.
 
-\section{Theoretical Method} \label{theory}
-The energy loss of ions in metal targets responds to different physical mechanisms, depending on the impact ion velocity.
-At low velocities, the binary collisions are responsible for the loss of energy by the ion.
-The main contribution is the ionization of electrons of the metal conduction band, which is well approximated by a free electron gas (FEG) of Fermi velocity $v_F$.
-Above certain velocity (i.e.
-$v\geq 1.5 \ v_F$) , not only binary but also collective excitations (plasmons) are possible \cite{mon17}.
-Moreover, at high energies, also the bound electrons contribute to the stopping power.
-The method used in this work combines a description for the interaction with the valence (or conduction) electrons as a FEG, and a different one for the interaction with the bound electrons.
+%=======================================================================
+\section{Theoretical Method} 
+\label{theory}
+The energy loss of ions in metal targets responds to different physical 
+mechanisms, depending on the impact ion velocity. At low velocities, the 
+binary collisions are responsible for the loss of energy by the ion. 
+The main contribution is the ionization of electrons of the metal 
+conduction band, which is well approximated by a free electron gas (FEG) 
+of Fermi velocity $v_F$. Above certain velocity (i.e. $v\geq 1.5\,v_F$), 
+not only binary but also collective excitations (plasmons) 
+occur~\cite{mon17}. Moreover, at high energies, also the bound electrons 
+contribute to the stopping power. The method used in this work combines 
+a description for the interaction with the valence (or conduction) 
+electrons as a FEG, and a different one for the interaction with the 
+bound electrons.
 
-We used the SPCC model \cite{mon17} to describe the stopping power of low velocity charged particles in the FEG.
-It is a non-perturbative binary collisional approximation, thus valid at energies below that of plasmon excitations.
-The SPCC \cite{mon17} is based on a screened central potential with cusp condition of the electronic density close to the projectile.
-This model proved to give a good description of the induced electron density even for negative projectiles \cite{mon17}, and reproduces the low velocity proton-antiproton differences in the stopping power (Barkas effect).
-The SPCC formalism only depends on the Wigner-Seitz radio, $r_S$, which is a measure of the electronic density of the FEG.
-For metals of well-known $r_S$, the SPCC describes correctly the low energy experimental stopping data \cite{mon17}, agreeing with the DFT results by Echenique and coworkers \cite{eche81,nagy89} at $v=0$.
+We used the SPCC model~\cite{mon17} to describe the stopping power of 
+low velocity charged particles in the FEG. It is a non-perturbative 
+binary collisional approximation, thus valid at energies below that of 
+plasmon excitations.  The SPCC~\cite{mon17} is based on a screened 
+central potential with cusp condition of the electronic density close 
+to the projectile. This model proved to give a good description of the 
+induced electron density even for negative projectiles~\cite{mon17}, 
+and reproduces the low velocity proton-antiproton differences in the 
+stopping power (Barkas effect). The SPCC formalism only depends on the 
+Wigner-Seitz radio, $r_S$, which is a measure of the electronic density 
+of the FEG. For metals of well-known $r_S$, the SPCC describes correctly 
+the low energy experimental stopping data~\cite{mon17}, agreeing with 
+the DFT results by Echenique and coworkers~\cite{eche81,nagy89} at $v=0$.
 
-%In the case of 
-Hafnium ($Z=72$, [Xe] $4f^{14} \ 6s^2 \ 5d_{3/2}^1 \ 5d_{5/2}^1$), belongs to the first groups of transition metals, with  4 electrons as FEG (theoretical $r_S=2.14$ a.u.) and $1s$-$4f$ electrons bound.
-We compared the computed $r_S$ with the experimental value obtained from measured energy loss function by Lynch \textit{et al} \cite{lynch75}.
-The experimental plasmon energy of Hf is $\hbar\omega_P \approx 15.8$ eV, with a width at half maximum $\delta \approx 4.4$ eV, and $r_S \approx 2.07$ a.u.
-\cite{lynch75}.
-The difference of less than $5\%$ between theoretical and experimental $r_S$ assess Hf as a canonical target \cite{mon17}.
+Hafnium ($Z=72$, [Xe] $4f^{14}\,6s^2\,5d_{3/2}^1\,5d_{5/2}^1$), belongs 
+to the first groups of transition metals, with four electrons as FEG 
+(theoretical $r_S=2.14$ a.u.) and $1s$-$4f$ electrons bound. We compared 
+the computed $r_S$ with the experimental value obtained from measured 
+energy loss function by Lynch \textit{et al} \cite{lynch75}. The 
+experimental plasmon energy of Hf is $\hbar\omega_P \approx 15.8$ eV, 
+with a width at half maximum $\delta \approx 4.4$ eV, and 
+$r_S \approx 2.07$ a.u. \cite{lynch75}. The difference of less than 
+$5\%$ between theoretical and experimental $r_S$ assess Hf as a 
+canonical target~\cite{mon17}.
 
-Above certain impact velocity, the plasmon contribution is important (i.e. around and above the stopping maximum). An interesting value for our analysis is the minimum impact velocity to excite
-plasmons, $v_P$.
-In the dielectric formalism, this value can be obtained as $v_P\ \approx \ v_F[1+(3 \pi\ v_F)^{-1/2} ]$ \cite{suppression}.
-To describe the energy loss considering collective and binary excitation, we resort to the ML dielectric formalism \cite{Mermin}, which is a linear response, perturbative approximation, so it depends on the square of the ion charge.
-In this formalism, the response of target electrons to the ion passage is described through the quantum dielectric function, depending on the characteristic $r_S$ and $\delta$ parameters of the FEG.
+Above certain impact velocity, the plasmon contribution is important 
+(i.e. around and above the stopping maximum). An interesting value for 
+our analysis is the minimum impact velocity to excite plasmons, $v_P$. 
+In the dielectric formalism, this value can be obtained as 
+$v_P\,\approx\,v_F[1+(3\pi\,v_F)^{-1/2}]$~\cite{suppression}. To 
+describe the energy loss considering collective and binary excitation, 
+we resort to the ML dielectric formalism \cite{Mermin}, which is a 
+linear response, perturbative approximation, so it depends on the 
+square of the ion charge. In this formalism, the response of target 
+electrons to the ion passage is described through the quantum dielectric 
+function, depending on the characteristic $r_S$ and $\delta$ parameters 
+of the FEG.
 
@@
 \includegraphics[width=11.cm]{bindener.eps}
-% \includegraphics[width=13.cm]{Hf_bindener_ratio.eps}
-\caption{(a) Binding energies of Hf.
-Non-relativistic and relativistic calculations are given with filled symbols.
-Experimental measurements for solids\cite{williams1995} are depicted with hollow circles.
+\caption{(a) Binding energies of Hf. Non-relativistic and relativistic 
+calculations are given with filled symbols. Experimental measurements 
+for solids~\cite{williams1995} are depicted with open circles. 
 (b) Corresponding relative errors respect to experimental data.}
@@
 
-For the stopping power due to bound electrons, the SLPA \cite{mon17,mon13} is employed.
-It is worth to mention that the only inputs for the SLPA are the space-dependent densities of each shell in the ground state, and their binding energies.
-Collective processes and screening among electrons are included.
-Hafnium is a relativistic target, therefore the wave functions and binding energies must be obtained by solving the many–electron Dirac Hamiltonian.
-Details of these calculations have been published in \cite{mendez2019}, and the Slater-type orbital expansions in \cite{Hf_arxiv}.
+For the stopping power due to bound electrons, the SLPA~\cite{mon17,mon13} 
+is employed. It is worth mentioning that the only inputs for the SLPA 
+are the space-dependent densities of each shell in the ground state, 
+and their binding energies. Collective processes and screening among 
+electrons are included. Since hafnium is a relativistic target, the 
+wave functions and binding energies must be obtained by solving the 
+many-electron Dirac Hamiltonian. Details of these calculations have been 
+published in \cite{mendez2019}, while Slater-type orbital expansions are 
+given in \cite{Hf_arxiv}.
 
-Too assess the importance of the description of bound electrons, Fig.
-\ref{Binding_E} shows (a) non-relativistic \cite{badnell97} and our relativistic binding energies, $E_{nl\pm}$, with $\pm=j\pm1/2$; and (b) the relative errors with respect to the experimental data on solid Hf \cite{williams1995} from $1s$ to $4f_{\pm}$.
-We notice that not only the most inner shells require relativistic calculations, but also the outer $5p$ and $4f$ shells.
-%Figure \ref{Binding_E} (b) displays the relative error of non relativistic and relativistic energies with respect to the experimental data in logarithmic scale. 
-Furthermore, this figure shows very clearly the disability of non-relativistic calculations to describe the experimental data, which surprisingly worsens from the inner to the outer shells. 
-%It is clear in this figure the disability of non-relativistic binding energies to describe the experimental data, which surprisingly worsens from the inner to the outer-shells.
-%We used here the theoretical binding energies and wave functions of Hf calculated in \cite{mendez2019}. 
-%The present theoretical description of Hf considers a FEG of $r_S=2.14$ and 1s-4f electrons as bound.
+To assess the importance of the description of bound electrons, 
+Fig.~\ref{Binding_E} shows (a) non-relativistic~\cite{badnell97} and 
+our relativistic binding energies $E_{nl\pm}$, with $\pm=j\pm1/2$; and 
+(b) the relative errors with respect to the experimental data on solid 
+Hf~\cite{williams1995} from $1s$ to $4f_{\pm}$. 
+%\st{We notice that not only the most inner shells require relativistic calculations, but also the outer $5p$ and $4f$ shells.
+Furthermore, this figure shows very clearly the disability of non-relativistic calculations to describe the experimental data, which surprisingly worsens from the inner to the outer shells.}
+One would expect the relativistic contributions to only affect 
+significantly the inner shells. The relativistic binding energies of the 
+electrons belonging to these shells are 2.5\% from the experimental values,
+while the non-relativistic ones are within $\sim$10\%. Moreover, we notice 
+the relativistic corrections greatly influences the outer $5p$ and $4f$ 
+shells. This effect may be a consequence of the orthogonality with the 
+relativistic inner shells. Also, the sign of the binding energy deviations 
+is inverted near the FEG limit, i.e. the electrons go from being more 
+bounded --than the experimental values-- to being less bounded. 
+This change of sign can be attributed to the fact that the experimental 
+values correspond to hafnium in solid-state, while our theoretical 
+calculations correspond to the element in the gas phase.
 
-%The relativistic results for binding energies and wave functions in \cite{mendez2019} 
-Our relativistic binding energies present spin-orbit split.
-However, in total stopping power where the initial state of the excited electron is not measured, the quantum uncertainty in energy $\Delta E$ melts this split.
-The criteria $\Delta E\Delta t \geq \hbar /2$ merges the energies  $E_{nl+} - E_{nl-}$ for sufficiently small values of $\Delta t$ (the collisional mean-time).
-In fact, at sufficiently high impact velocity, we can expect all target electrons to respond together to the ion passage \cite{lindhard53,chu72}.
-Following previous work \cite{mon09}, the collisional time is estimated as $\Delta t \approx \langle r_i\rangle/v$, with $\langle r_i\rangle$ and $v$ being the orbital mean radio and impact velocity, respectively.
+Our relativistic binding energies present spin-orbit split. However, in 
+total stopping power, where the initial state of the excited electron is 
+not measured, the quantum uncertainty in energy $\Delta E$ melts this 
+split. The criteria $\Delta E\Delta t\geq\hbar/2$ merges the energies 
+$E_{nl+}-E_{nl-}$ for sufficiently small values of $\Delta t$ (the 
+collisional mean-time). In fact, at sufficiently high impact velocity, 
+we can expect all target electrons to respond together to the ion 
+passage~\cite{lindhard53,chu72}. Following previous work~\cite{mon09}, 
+the collisional time is estimated as $\Delta t\approx\langle r_i\rangle/v$, 
+with $\langle r_i\rangle$ and $v$ being the orbital mean radio and 
+impact velocity, respectively.
 
-%--------------------------------------------------------------------------------
+%-----------------------------------------------------------------------
 \begin{figure*}[!t]
@@
 \includegraphics[width=13.cm]{Fig02.eps}
-\caption{Theoretical total stopping cross sections of protons in hafnium adding FEG and bound $1s$-$4f$ contributions, as mentioned in the inset.
-The vertical grey dashed-line indicates the energy above which plasmon excitation is possible.
-%The FEG calculations employ the SPCC model at low energies (no plasmon excitation) and the ML formalism (includes plasmon excitation).
-The SLPA total stopping values for $1s$-$4f$ bound electrons with and without $5p-4f$ screening are displayed with solid and dotted lines, respectively.}
+\caption{Theoretical total stopping cross sections of protons in hafnium 
+adding FEG and bound $1s$-$4f$ contributions, as mentioned in the inset. 
+The vertical grey dashed-line indicates the energy above which plasmon 
+excitation is possible. The SLPA total stopping values for $1s$-$4f$ 
+bound electrons with and without $5p$-$4f$ screening are displayed with 
+solid and dotted lines, respectively.}
 \label{slpa4f}
 \end{figure*}
-%========================================================================================
+%-----------------------------------------------------------------------
 
-The SLPA calculates the contribution of each subshell of bound electrons to the total stopping cross sections.
-We found that for every sub-shell of Hf, at the impact energy at which this sub-shell began to contribute, the spin-orbit split was unresolved.
-Then, the $nl$-electrons should be considered together, responding to the ion passage as a single gas of electrons with density $\delta_{nl}(r)$ and a mean binding energy $E_{nl}$.
-This feature is important within the SLPA calculations because it accounts for the screening among electrons of the \textit{same} binding energy.
-For example, the $4f_{-}$ and $4f_{+}$ of Hf can only be resolved for impact energies $E<0.05$ keV, but the contribution of $4f$ to the total stopping is negligible for $E<40$ keV.
-Moreover, the $5p$ and $4f$ electrons of Hf are very close in energy ( $\Delta E_{5p-4f} \approx 1$ a.u.
-\cite{mendez2019}) %merges them
-and they react together for impact energies $E>40$ keV (inter-shell screening).
+The SLPA calculates the contribution of each subshell of bound electrons 
+to the total stopping cross sections. We found that for every sub-shell 
+of Hf, at the impact energy at which this sub-shell began to contribute, 
+the spin-orbit split was unresolved. Then, the $nl$-electrons should be 
+considered together, responding to the ion passage as a single gas of 
+electrons with density $\delta_{nl}(r)$ and a mean binding energy 
+$E_{nl}$. This feature is important within the SLPA calculations because 
+it accounts for the screening among electrons of the \textit{same} 
+binding energy. For example, the $4f_{-}$ and $4f_{+}$ of Hf can only 
+be resolved for impact energies $E<0.05$ keV, but the contribution of 
+$4f$ to the total stopping is negligible for $E<40$ keV. Moreover, the 
+$5p$ and $4f$ electrons of Hf are very close in energy 
+($\Delta E_{5p-4f} \approx 1$ a.u.~\cite{mendez2019}) and they react 
+together for impact energies $E>40$ keV (inter-shell screening).
 
-%================================================================================
+%-----------------------------------------------------------------------
 \begin{figure*}[!t]
@@
 \includegraphics[width=13.0cm]{Fig03.eps}
-\caption{Stopping power cross section of hafnium for protons.
-Symbols: solid circles, present values; open circles, previous data \cite{Sirotinin}.
-Curves: Black solid-line, present full theoretical results with $4f$-$5p$ screening; violet dashed-line, values from ICRU-49 \cite{ICRU49}, and green dotted line, SRIM-2013 \cite{Ziegler01}.}
+\caption{Stopping power cross section of hafnium for protons. Symbols: 
+solid circles, present values; open circles, previous data~\cite{Sirotinin}. 
+Curves: Black solid-line, present full theoretical results with 
+$4f$-$5p$ screening; violet dashed-line, values from ICRU-49~\cite{ICRU49}, 
+and green dotted line, SRIM-2013~\cite{Ziegler01}.}
 \label{F03}
 \end{figure*}
-%=============================================================================================================================================================================================
+%-----------------------------------------------------------------------
 
-In Fig.
-\ref{slpa4f}, we display the present theoretical stopping cross section of Hf for protons with and without the $5p$-$4f$ screening.
-We added the FEG and the bound electron contributions as explained above.
-The minimum energy for plasmon excitation was estimated to $37$ keV.
-We used the non-perturbative SPCC model for impact energies $E \leq 37$ keV, and the perturbative ML calculation above this energy.
-%The range of validity of the perturbative model below 100 keV is arguable, however the asymetric $Z_{ion}/Z_{target}$ relation is clear.
-%The importance of the 4f-5p screening is signed with arrows in this figure. 
-Clearly, considering $5p$-$4f$ electrons as a single group of 20 electrons with screening among them gives lower stopping values than the addition of the separate $5p$ and $4f$ contributions.
-Notice that this shell correction can only be considered within a many electron model such as the SLPA.
+In Fig.~\ref{slpa4f}, we display the present theoretical stopping cross 
+section of Hf for protons with and without the $5p$-$4f$ screening. We 
+added the FEG and the bound electron contributions as explained above. 
+The minimum energy for plasmon excitation was estimated to $37$ keV. We 
+used the non-perturbative SPCC model for impact energies $E \leq 37$~keV, 
+and the perturbative ML calculation above this energy. Clearly, 
+considering $5p$-$4f$ electrons as a single group of 20 electrons with 
+screening among them gives lower stopping values than the addition of 
+the separate $5p$ and $4f$ contributions. Notice that this shell 
+correction can only be considered within a many electron model such as 
+the SLPA.
 
-%=============================================================================================================================================================================================
+%=======================================================================
 \section{Analysis of the results and discussion}
@@
 
-The present data and theoretical results are displayed in Table \ref{table01}, and Fig.
-\ref{F03}.
-As can be seen in Table~\ref{table01}, an overall relative uncertainty of around 5$\%$ was achieved for the experimental stopping power values, which are mainly due to the uncertainty in the hafnium foil thickness.
-%A good agreement in the whole energy range is found between our new data and the previous measurements made by Sirotinin \textit{et al.} \cite{Sirotinin}, as depicted in Fig.~\ref{F03}.
-Also shown in the same figure are our theoretical stopping cross sections together with the semi-empirical SRIM-2013 curve \cite{Ziegler01} and the suggested values by the ICRU-49 report \cite{ICRU49}.
+The present data and theoretical results are displayed in 
+Table~\ref{table01}, and Fig.~\ref{F03}. As can be seen in 
+Table~\ref{table01}, an overall relative uncertainty of around 5\% was 
+achieved for the experimental stopping power values, which are mainly 
+due to the uncertainty in the hafnium foil thickness.
 
-%-----------------TABLE--------------------------------------------------------------------------
+%-----------------------------------------------------------------------
 \begin{table*}[!t]
 \centering
-\caption{Stopping power values S$_{exp}$ of hafnium for protons measured in this work. $\Delta$E/E values are also shown.
+\caption{Stopping power values S$_{\mathrm{exp}}$ of hafnium for protons measured 
+in this work. $\Delta$E/E values are also shown.
 \label{table01}}
@@
 \begin{tabular}{ccc|ccc|ccc} %\hline\hline
-E$_{avg}$	&S$_{exp}$			&$\Delta$E/E	&E$_{avg}$	&S$_{exp}$			&$\Delta$E/E	&E$_{avg}$	&S$_{exp}$			&$\Delta$E/E\\
-keV		&eV/(10$^{15}$ at/cm$^2$)	&$\%$		&keV		&eV/(10$^{15}$ at/cm$^2$)	&$\%$		&keV		&eV/(10$^{15}$ at/cm$^2$)	&$\%$\\ \hline
-516,6	&	25,8	$\pm$	1,3	&	20,5	&	1170,3	&	18,25	$\pm$	0,91	&	6,4	&	1813,4	&	15,10	$\pm$	0,76	&	3,4	\\
-567,8	&	24,8	$\pm$	1,2	&	17,9	&	1220,0	&	18,08	$\pm$	0,90	&	6,1	&	1862,7	&	14,79	$\pm$	0,74	&	3,3	\\
-618,8	&	23,9	$\pm$	1,2	&	15,8	&	1269,6	&	17,57	$\pm$	0,88	&	5,7	&	1912,0	&	14,21	$\pm$	0,71	&	3,0	\\
-669,6	&	23,2	$\pm$	1,2	&	14,2	&	1319,2	&	17,32	$\pm$	0,87	&	5,4	&	1961,2	&	14,46	$\pm$	0,72	&	3,0	\\
-720,1	&	22,5	$\pm$	1,1	&	12,8	&	1368,8	&	17,15	$\pm$	0,86	&	5,1	&	2010,4	&	14,34	$\pm$	0,72	&	2,9	\\
-770,5	&	21,8	$\pm$	1,1	&	11,6	&	1418,3	&	16,69	$\pm$	0,83	&	4,8	&	2059,6	&	13,76	$\pm$	0,69	&	2,7	\\
-820,8	&	21,3	$\pm$	1,1	&	10,7	&	1467,8	&	16,43	$\pm$	0,82	&	4,6	&	2108,8	&	13,78	$\pm$	0,69	&	2,7	\\
-871,0	&	20,8	$\pm$	1,0	&	9,8	&	1517,2	&	16,13	$\pm$	0,81	&	4,4	&	2158,0	&	13,70	$\pm$	0,69	&	2,6	\\
-921,1	&	20,3	$\pm$	1,0	&	9,1	&	1566,7	&	16,04	$\pm$	0,80	&	4,2	&	2206,5	&	13,33	$\pm$	0,67	&	2,5	\\
-971,1	&	19,9	$\pm$	1,0	&	8,4	&	1616,0	&	15,77	$\pm$	0,79	&	4,0	&	2256,4	&	13,27	$\pm$	0,66	&	2,4	\\
-1021,0	&	19,33	$\pm$	0,97	&	7,8	&	1665,4	&	15,51	$\pm$	0,78	&	3,8	&	2305,5	&	13,07	$\pm$	0,65	&	2,3	\\
-1070,8	&	19,03	$\pm$	0,95	&	7,3	&	1714,8	&	15,46	$\pm$	0,77	&	3,7	&	2354,7	&	12,91	$\pm$	0,65	&	2,2	\\
-1120,6	&	18,73	$\pm$	0,94	&	6,9	&	1764,1	&	14,93	$\pm$	0,75	&	3,5	&	2403,8	&	12,61	$\pm$	0,63	&	2,2	\\ \\ %\hline
+E$_{\mathrm{avg}}$ & S$_{\mathrm{exp}}$       & $\Delta$E/E & E$_{\mathrm{avg}}$ & S$_{\mathrm{exp}}$       & $\Delta$E/E & E$_{\mathrm{avg}}$ & S$_{\mathrm{exp}}$       & $\Delta$E/E \\
+keV                & eV/(10$^{15}$ at/cm$^2$) & \%          & keV                & eV/(10$^{15}$ at/cm$^2$)	& \%          & keV                & eV/(10$^{15}$ at/cm$^2$) & \% \\ \hline
+516.6	 & 25.8	$\pm$	1.3	&	20.5	&	1170.3	&	18.25	$\pm$	0.91	&	6.4	&	1813.4	&	15.10	$\pm$	0.76	&	3.4	\\
+567.8	 & 24.8	$\pm$	1.2	&	17.9	&	1220.0	&	18.08	$\pm$	0.90	&	6.1	&	1862.7	&	14.79	$\pm$	0.74	&	3.3	\\
+618.8	 & 23.9	$\pm$	1.2	&	15.8	&	1269.6	&	17.57	$\pm$	0.88	&	5.7	&	1912.0	&	14.21	$\pm$	0.71	&	3.0	\\
+669.6	 & 23.2	$\pm$	1.2	&	14.2	&	1319.2	&	17.32	$\pm$	0.87	&	5.4	&	1961.2	&	14.46	$\pm$	0.72	&	3.0	\\
+720.1	 & 22.5	$\pm$	1.1	&	12.8	&	1368.8	&	17.15	$\pm$	0.86	&	5.1	&	2010.4	&	14.34	$\pm$	0.72	&	2.9	\\
+770.5	 & 21.8	$\pm$	1.1	&	11.6	&	1418.3	&	16.69	$\pm$	0.83	&	4.8	&	2059.6	&	13.76	$\pm$	0.69	&	2.7	\\
+820.8	 & 21.3	$\pm$	1.1	&	10.7	&	1467.8	&	16.43	$\pm$	0.82	&	4.6	&	2108.8	&	13.78	$\pm$	0.69	&	2.7	\\
+871.0	 & 20.8	$\pm$	1.0	&	9.8	&	1517.2	&	16.13	$\pm$	0.81	&	4.4	&	2158.0	&	13.70	$\pm$	0.69	&	2.6	\\
+921.1	 & 20.3	$\pm$	1.0	&	9.1	&	1566.7	&	16.04	$\pm$	0.80	&	4.2	&	2206.5	&	13.33	$\pm$	0.67	&	2.5	\\
+971.1	 & 19.9	$\pm$	1.0	&	8.4	&	1616.0	&	15.77	$\pm$	0.79	&	4.0	&	2256.4	&	13.27	$\pm$	0.66	&	2.4	\\
+1021.0 & 19.33	$\pm$	0.97	&	7.8	&	1665.4	&	15.51	$\pm$	0.78	&	3.8	&	2305.5	&	13.07	$\pm$	0.65	&	2.3	\\
+1070.8 & 19.03	$\pm$	0.95	&	7.3	&	1714.8	&	15.46	$\pm$	0.77	&	3.7	&	2354.7	&	12.91	$\pm$	0.65	&	2.2	\\
+1120.6 & 18.73	$\pm$	0.94	&	6.9	&	1764.1	&	14.93	$\pm$	0.75	&	3.5	&	2403.8	&	12.61	$\pm$	0.63	&	2.2	\\ \\ 
 \end{tabular}
@@
 
-In Fig.
-\ref{F03}, we have good agreement between the present theoretical results and the measurements displayed in Table~\ref{table01}.
-Our theoretical approach also agrees with previous data by Sirotinin \cite{Sirotinin}, except for the lowest energy measurement at 80 keV.
-It is interesting that our full theoretical curve differs from SRIM-2013 for impact energies below 100 keV.
-We obtain a stopping maximum around $40 \times 10^{-15}$ eV cm$^2$/atom  at 65 keV.
-Instead, following the up-to-now only set of data \cite{Sirotinin}, SRIM-2013 suggests a lower stopping maximum  at impact energy of 115 kev.
-It is worth mentioning that similar theoretical results using the experimental $r_S=2.07$ a.u.
-instead of the theoretical $r_S=2.14$ a.u.
-give the stopping maximum at the same impact energy but $4\%$ higher.
-Future experiments for
-impact energies below 100 keV would be important to complete this study.
+In Fig.~\ref{F03}, we have good agreement between the present 
+theoretical results and the measurements displayed in Table~\ref{table01}. 
+Our theoretical approach also agrees with previous data by 
+Sirotinin~\cite{Sirotinin}, except for the lowest energy measurement at 
+80 keV. It is interesting that our full theoretical curve differs from 
+SRIM-2013 for impact energies below 100 keV. We obtain a stopping 
+maximum around $40 \times 10^{-15}$ eV cm$^2$/atom  at 65 keV. Instead, 
+following the up-to-now only set of data \cite{Sirotinin}, SRIM-2013 
+suggests a lower stopping maximum  at impact energy of 115 kev. It is 
+worth mentioning that similar theoretical results using the experimental 
+$r_S=2.07$ a.u. instead of the theoretical $r_S=2.14$ a.u. give the 
+stopping maximum at the same impact energy but $4\%$ higher. Future 
+experiments for impact energies below 100 keV would be important to 
+complete this study.
 
-%=============================================================================================================================================================================================
-%=============================================================================================================================================================================================
+%=======================================================================
 \section{Conclusion}
 \label{conclusion}
-In this work, we have used the transmission method to experimentally determine stopping power cross section values for (0.6-2.5) MeV protons incident on self-supporting Hf foils with an overall uncertainty of around 5\%.
-Additionally, we calculated values extracted from the theoretical framework that involved the relativistic wave functions and binding energies of Hf, and considered 4 electrons per atom in the free electron gas.
-The shell-wise local plasma approximation was employed to describe the energy transferred to the bound $1s$-$4f$ electrons, and two different models for the FEG: the screened potential with cusp condition (SPCC model) for energies below that of the plasmon excitation, and the Mermin-Lindhard dielectric formalism, for energies around the stopping maximum and above.
-Present theoretical stopping cross sections cover an extensive energy range from 1 keV/amu-10 meV/amu.
 
-At high impact energies, the new stopping  measurements are in good agreement with our theoretical results, with previous experimental data and semi-empirical values by SRIM-2013 and ICRU-49.
-However, we call the attention that around the stopping maximum and at lower impact energies the difference between our full-theoretical results and SRIM is substantial.
-To the best of our knowledge, these are the first theoretical calculations of stopping in Hf taking into account the atomic relativistic effect.
-Future experiments for impact energies below 100 keV would be important to complete this study.
-%=======================================================================================================================================
+In this work, we have used the transmission method to experimentally 
+determine stopping power cross section values for (0.6-2.5) MeV protons 
+incident on self-supporting Hf foils with an overall uncertainty of 
+around 5\%. Additionally, we calculated values extracted from the 
+theoretical framework that involved the relativistic wave functions and 
+binding energies of Hf, and considered four electrons per atom in the free 
+electron gas. The shell-wise local plasma approximation was employed to 
+describe the energy transferred to the bound $1s$-$4f$ electrons, and 
+two different models for the FEG: the screened potential with cusp 
+condition (SPCC model) for energies below that of the plasmon 
+excitation, and the Mermin-Lindhard dielectric formalism, for energies 
+around the stopping maximum and above. Present theoretical stopping 
+cross sections cover an extensive energy range from 1 keV/amu-10 meV/amu.
+
+At high impact energies, the new stopping  measurements are in good 
+agreement with our theoretical results, with previous experimental data 
+and semi-empirical values by SRIM-2013 and ICRU-49.  However, we call 
+the attention that around the stopping maximum and at lower impact 
+energies the difference between our full-theoretical results and SRIM 
+is substantial. To the best of our knowledge, these are the first 
+theoretical calculations of stopping in Hf taking into account the 
+atomic relativistic effect. Future experiments for impact energies 
+below 100 keV would be important to complete this study.
+
+%=======================================================================
 \begin{acknowledgments}
-This work was funded by VRAC Grant Number L1-17 of Universidad Tecnol\'ogica Metropolitana, Chile; also by the Consejo Nacional de Investigaciones Cient\'ificas y T\'ecnicas (CONICET), the Agencia Nacional de Promoci\'on Cient\'ifica y Tecnol\'ogica (ANPCyT), and Universidad de Buenos Aires (UBA), from Argentina.
+This work was funded by VRAC Grant Number L1-17 of Universidad 
+Tecnol\'ogica Metropolitana, Chile; also by the Consejo Nacional de 
+Investigaciones Cient\'ificas y T\'ecnicas (CONICET), the Agencia 
+Nacional de Promoci\'on Cient\'ifica y Tecnol\'ogica (ANPCyT), and 
+Universidad de Buenos Aires (UBA), from Argentina.
 
-The authors gratefully acknowledge the invaluable contribution of F. Baptista from ITN/IST-UTL, Sacav\'{e}m, Portugal for his constant availability. 
+The authors gratefully acknowledge the invaluable contribution of F. 
+Baptista from ITN/IST-UTL, Sacav\'{e}m, Portugal for his constant 
+availability. 
 \end{acknowledgments}
-
-%=======================================================================================================================================
+%=======================================================================
 %  references
 \begin{thebibliography}{00}
-%============================================================================================================================
+%=======================================================================
 %Fundamentals of Stopping Power and Energy Loss processes
-\bibitem{Chu01} W.K. Chu, J.W. Mayer, M.A. Nicolet (Eds.) Backscattering Spectrometry, Academic Press, New York, 1978.
-\bibitem{Sigmund} P.
-Sigmund, ``Particle Penetration and Radiation Effects.
-General Aspects and Stopping of Swift Point Charges''.
-Springer Series in Solid-State Sciences, Vol.
-151, Springer, Berlin (2006).
-\bibitem{Schardt} D. Schardt, T. Els\"asser, D. Schulz-Ertner, Rev. Mod. Phys. \textbf{82},  383-425 (2010).
-\bibitem{Diwan} P.K. Diwan, S. Kumar, Nucl. Instr. and Meth. B \textbf{359}, 78-84 (2015).
-%\bibitem{Damache01} D. Moussa, S. Damache and S. Ouichaoui, Nucl. Instr. and Meth. \textbf{B} 343,  44-47 (2015).
-\bibitem{Damache02} D. Moussa, S. Damache and S. Ouichaoui, Nucl. Instr. and Meth. B \textbf{268}, 1754-1758 (2010); \textbf{343},  44-47 (2015).
-%\bibitem{Damache03} S. Damache, S. Ouichaoui, D. Moussa and A. Dib, Nucl. Instr. and Meth. \textbf{B} 249,  22-25 (2006).
-\bibitem{Damache04} S. Damache, S. Ouichaoui, A. Belhout, A. Medouni and I. Toumert, Nucl. Instr. and Meth. B \textbf{225}, 449-463 (2004).
-%============================================================================================================================
+\bibitem{Chu01} 
+W.K. Chu, J.W. Mayer, and M.A. Nicolet,
+\textit{Backscattering Spectrometry}
+(Academic Press, New York, 1978).
+
+\bibitem{Sigmund} 
+P. Sigmund, 
+\textit{Particle Penetration and Radiation Effects. General Aspects and 
+Stopping of Swift Point Charges}.
+(Springer Series in Solid-State Sciences, Springer, Berlin, 2006), Vol. 151.
+
+\bibitem{Schardt} 
+D. Schardt, T. Els\"asser, D. Schulz-Ertner, 
+Rev. Mod. Phys. \textbf{82},  383-425 (2010).
+
+\bibitem{Diwan} 
+P.K. Diwan, S. Kumar, 
+Nucl. Instr. and Meth. B \textbf{359}, 78-84 (2015).
+
+\bibitem{Damache02} 
+D. Moussa, S. Damache and S. Ouichaoui, 
+Nucl. Instr. and Meth. B \textbf{268}, 1754-1758 (2010); 
+\textbf{343},  44-47 (2015).
+
+\bibitem{Damache04} 
+S. Damache, S. Ouichaoui, A. Belhout, A. Medouni and I. Toumert, 
+Nucl. Instr. and Meth. B \textbf{225}, 449-463 (2004).
+
+%=======================================================================
 %% Use of Hafnium and Stopping Power of Hafnium
-\bibitem{Sirotinin} E.I. Sirotinin, A.F. Tulinov, V.A. Khodyrev, V.N. Mizgulin, Nucl. Instr. and Meth. B \textbf{4}, 337-345 (1984).
-\bibitem{Abril} I. Abril, M. Behar, R. Garcia Molina, R.C. Fadanelli, L.C.C.M. Nagamine, P.L. Grande, L. Sch\"unemann, C.D. Denton, N.R. Arista, E.B. Saitovich,
+\bibitem{Sirotinin} 
+E.I. Sirotinin, A.F. Tulinov, V.A. Khodyrev, V.N. Mizgulin, 
+Nucl. Instr. and Meth. B \textbf{4}, 337-345 (1984).
+
+\bibitem{Abril} 
+I. Abril, M. Behar, R. Garcia Molina, R.C. Fadanelli, L.C.C.M. Nagamine, 
+P.L. Grande, L. Sch\"unemann, C.D. Denton, N.R. Arista, E.B. Saitovich,
 Eur. Phys. J. D \text{54}, 65-70 (2009).
-\bibitem{Behar} M. Behar, R.C. Fadanelli, I. Abril, R. Garcia-Molina, C.D. Denton, L.C.C.M. Nagamine, N.R. Arista, Phys. Rev. A \textbf{80},  062901 (2009).
-\bibitem{Primetzhofer} D. Primetzhofer, Nucl. Instr. and Meth. B \textbf{320}, 100-103 (2014).
-%============================================================================================================================
+
+\bibitem{Behar} 
+M. Behar, R.C. Fadanelli, I. Abril, R. Garcia-Molina, C.D. Denton, 
+L.C.C.M. Nagamine, N.R. Arista, 
+Phys. Rev. A \textbf{80},  062901 (2009).
+
+\bibitem{Primetzhofer} 
+D. Primetzhofer, 
+Nucl. Instr. and Meth. B \textbf{320}, 100-103 (2014).
+
+\bibitem{Roth}
+D. Roth, B. Bruckner, G. Undeutsch, V. Paneta, A. I. Mardare, 
+C. L. McGahan, M. Dosmailov, J. I. Juaristi, M. Alducin, 
+J. D. Pedarnig, R. F. Haglund, Jr., D. Primetzhofer, and P. Bauer
+Phys. Rev. Lett. \textbf{119}, 163401 (2017).
+
+%=======================================================================
 % Why Hafnium is so important:
-\bibitem{Choi} J.H. Choi, Y. Mao, J.P. Chang, Mat. Sci. Eng. R \textbf{72}, 97-136 (2011).
-\bibitem{Robertson} J. Robertson, R.M. Wallace, Mat. Sci. Eng. R \textbf{88}, 1-41 (2015).
-%============================================================================================================================
+\bibitem{Choi} 
+J.H. Choi, Y. Mao, J.P. Chang, 
+Mat. Sci. Eng. R \textbf{72}, 97-136 (2011).
+
+\bibitem{Robertson} 
+J. Robertson, R.M. Wallace, 
+Mat. Sci. Eng. R \textbf{88}, 1-41 (2015).
+
+%=======================================================================
 %Importance of Stopping Power for Ion Beam Analysis
-\bibitem{Alfassi01} Z.B. Alfassi, "Non-destructive elemental analysis", Blackwell Science Ltd. (2001).
-\bibitem{Tesmer01} J.R. Tesmer, M. Nastasi, J.C. Barbour, C.J. Maggiore and J.W. Mayer, ''Handbook of Modern Ion Beam Material Analysis``, Materials Research Society (1995).
-\bibitem{HPaul03} https://www-nds.iaea.org/stopping/.
-\bibitem{mondim17} C.C. Montanari, and P. Dimitriou, Nucl. Instr. and Meth. B \textbf{408},  50-55 (2017).
-\bibitem{Roth17} D.
-Roth, B.
-Bruckner, M.
-V.
-Moro, S.
-Gruber, D.
-Goebl, J.
-I.
-Juaristi, M.
-Alducin,R.
-Steinberger, J.
-Duchoslav, D.
-Primetzhofer, and P.
-Bauer, Phys.
-Rev.
-Lett.
-\textbf{118}, 103401 (2017).
+\bibitem{Alfassi01} 
+Z.B. Alfassi,
+\textit{Non-destructive elemental analysis}
+(Blackwell Publishing, Oxford, 2001).
 
-%============================================================================================================================
+\bibitem{Tesmer01} 
+J.R. Tesmer, M. Nastasi, J.C. Barbour, C.J. Maggiore and J.W. Mayer,
+\textit{Handbook of Modern Ion Beam Material Analysis}.
+(Materials Research Society, Pittsburgh, 1995).
+
+\bibitem{HPaul03} 
+https://www-nds.iaea.org/stopping/.
+
+\bibitem{mondim17} 
+C.C. Montanari, and P. Dimitriou, 
+Nucl. Instr. and Meth. B \textbf{408},  50-55 (2017).
+
+\bibitem{Roth17} 
+D. Roth, B. Bruckner, M. V. Moro, S. Gruber, D. Goebl, J. I. Juaristi, 
+M. Alducin,R. Steinberger, J. Duchoslav, D. Primetzhofer, and P. Bauer, 
+Phys. Rev. Lett. \textbf{118}, 103401 (2017).
+
+%=======================================================================
 %Semi-empirical and Theoretical approach
-\bibitem{mon13} C.C. Montanari and J.E. Miraglia, Adv. Quant. Chem. \textbf{65}, edited by Dz. Belkic (Elsevier, Amsterdam, 2013), Chap. 7, pp. 165-201.
-\bibitem{mon17} C.C. Montanari, and J.E. Miraglia, Phys. Rev. A \textbf{96}, 012707 (2017).
-\bibitem{Mermin} N.D. Mermin, Phys. Rev. B \textbf{1}, 2362 (1970).
-%\bibitem{badnell97} N.R. Badnell, Comput. Phys. Commun., 182(7), 1528-1535 (2011). %https://doi.org/10.1016/j.cpc.2011.03.023
-\bibitem{mendez2019} A.M.P. Mendez, C.C. Montanari, D.M. Mitnik, Nucl. Instrum. Meth. B \textbf{460}, 114-118 (2019).
-\bibitem{ICRU49} ICRU report 49, Stopping Powers and Ranges for Protons and Alpha Particles, International Commission on Radiation Units and Measurements, 1993.
-\bibitem{Ziegler01} J.F. Ziegler, J.P. Biersack, SRIM2013, Computer Program and Manual. Available from www.srim.org.
-\bibitem{zenodo} 10.5281/zenodo.3678785 %10.5281/zenodo.3678760 
-%============================================================================================================================
+\bibitem{mon13} 
+C.C. Montanari and J.E. Miraglia in 
+\textit{Advances in Quantum Chemistry}, 
+edited by D. Belki\'c, (Elsevier, Amsterdam, 2013), Vol 65, Chap. 7, pp. 165-201.
+
+\bibitem{mon17} 
+C.C. Montanari, and J.E. Miraglia, 
+Phys. Rev. A \textbf{96}, 012707 (2017).
+
+\bibitem{Mermin} 
+N.D. Mermin, 
+Phys. Rev. B \textbf{1}, 2362 (1970).
+
+\bibitem{mendez2019} 
+A.M.P. Mendez, C.C. Montanari, D.M. Mitnik, 
+Nucl. Instrum. Meth. B \textbf{460}, 114-118 (2019).
+
+\bibitem{ICRU49} 
+ICRU report 49, 
+Stopping Powers and Ranges for Protons and Alpha Particles, 
+International Commission on Radiation Units and Measurements, 1993.
+
+\bibitem{Ziegler01} 
+J.F. Ziegler, J.P. Biersack, SRIM2013, 
+Computer Program and Manual. Available from www.srim.org.
+
+\bibitem{zenodo} 
+C.C. Montanari {\it et al.} DOI: 10.5281/zenodo.3678785
+
+%=======================================================================
 % Experimental
-\bibitem{Miranda01} P.A. Miranda, A. Sep\'ulveda, J.R. Morales, T. Rodriguez, E. Burgos, H. Fern\'andez, Nucl. Instr. and Meth. B \textbf{318}, 292-296  (2014).
-\bibitem{Lebow} Lebow Company. 5960 Mandarin Ave. Goleta CA, 93117, USA.
+\bibitem{Miranda01} P.A. Miranda, A. Sep\'ulveda, J.R. Morales, 
+T. Rodriguez, E. Burgos, H. Fern\'andez,
+Nucl. Instr. and Meth. B \textbf{318}, 292-296  (2014).
 
-%============================================================================================================================
+\bibitem{Lebow} 
+Lebow Company. 5960 Mandarin Ave. Goleta CA, 93117, USA.
+
+%=======================================================================
 % Transmission Method
-\bibitem{Sun01} G. Sun, M. D\"{o}belli, A.M. M\"{u}ller, M. Stocker, M. Suter and L. Wacker, Nucl. Instr. and Meth. B \textbf{256}, 586-590 (2007).
-\bibitem{Raisanen01} J. Raisanen, U. Watjen, A.J.M. Plompen, F. Munnik, Nucl. Instr. and Meth. \textbf{B} 118, 1-6  (1996).
-\bibitem{Schulz01} F. Schulz and J. Shchuchinzky, Nucl. Instr. and Meth. B \textbf{12},  90-94 (1985).
-\bibitem{Chilton} A.B. Chilton, J.N. Cooper and J.C. Harris, Phys. Rev. \textbf{93}, 413-418  (1954).
-\bibitem{Rajatora} M. Rajatora, K. Vakevainen, T. Ahlgre, E. Rauhala, J. Raisanen, K. Rakennus, Nucl. Instr. and Meth. B \textbf{119}, 457-462 (1996).
-%============================================================================================================================
+\bibitem{Sun01} 
+G. Sun, M. D\"{o}belli, A.M. M\"{u}ller, M. Stocker, M. Suter and 
+L. Wacker, 
+Nucl. Instr. and Meth. B \textbf{256}, 586-590 (2007).
+
+\bibitem{Raisanen01} 
+J. Raisanen, U. Watjen, A.J.M. Plompen, F. Munnik, 
+Nucl. Instr. and Meth. \textbf{B} 118, 1-6  (1996).
+
+\bibitem{Schulz01} 
+F. Schulz and J. Shchuchinzky, 
+Nucl. Instr. and Meth. B \textbf{12},  90-94 (1985).
+
+\bibitem{Chilton} 
+A.B. Chilton, J.N. Cooper and J.C. Harris, 
+Phys. Rev. \textbf{93}, 413-418  (1954).
+
+\bibitem{Rajatora} 
+M. Rajatora, K. Vakevainen, T. Ahlgre, E. Rauhala, J. Raisanen, 
+K. Rakennus, 
+Nucl. Instr. and Meth. B \textbf{119}, 457-462 (1996).
+
+%=======================================================================
 %X Theoretical Approach (Montanari)
-\bibitem{eche81} P. M. Echenique, R. M. Nieminen and R. H. Ritchie, Sol. State Comm. \textbf{37}, 779-781 (1981).
-\bibitem{nagy89} I. Nagy, A. Arnau, P. M. Echenique, and E. Zaremba, Phys. Rev. B \textbf{40}, 11983 (1989).
-% \bibitem{Vs94} J. E. Valdes, J. C. Eckardt, G. H. Lantschner, and N. R. Arista, Phys. Rev. A \textbf{49}, 1083 (1994).
-\bibitem{lynch75} D. W. Lynch, C. G. Olson, and J. H. Weaver, Phys. Rev. B \textbf{11}, 3617-3624 (1975).
-\bibitem{suppression} C. C. Montanari, J. E. Miraglia, and N. R. Arista, Phys. Rev. A \textbf{62}, 052902 (2000).
-\bibitem{Hf_arxiv} A.M.P.
-Mendez, C.C.
-Montanari, D.M.
-Mitnik, \textit{Slater-type orbital expansion of neutral hafnium, numerical solution of the relativistic Dirac equation}, available soon in arXiv.org.
-\bibitem{badnell97} N.R. Badnell, Comput. Phys. Commun., 182(7), 1528-1535 (2011). %https://doi.org/10.1016/j.cpc.2011.03.023
-\bibitem{williams1995} G. Williams in http://xdb.lbl.gov/Section1/Sec\_1-1.html
-\bibitem{lindhard53} J. Lindhard, M. Scharff,  Mat. Fys. Medd. Dan. Vid. Selsk  \textbf{27}, 1 (1953).
-\bibitem{chu72} W. K. Chu, D. Powers, Rev. Lett. A \textbf{40}, 23 (1972).
-\bibitem{mon09} C.C. Montanari, D.M. Mitnik, C.D. Archubi, and J.E. Miraglia, Phys. Rev. A \textbf{80}, 012901 (2009).
+\bibitem{eche81} 
+P. M. Echenique, R. M. Nieminen and R. H. Ritchie, 
+Sol. State Comm. \textbf{37}, 779-781 (1981).
+
+\bibitem{nagy89} 
+I. Nagy, A. Arnau, P. M. Echenique, and E. Zaremba, 
+Phys. Rev. B \textbf{40}, 11983 (1989).
+
+\bibitem{lynch75} 
+D. W. Lynch, C. G. Olson, and J. H. Weaver, 
+Phys. Rev. B \textbf{11}, 3617-3624 (1975).
+
+\bibitem{suppression} 
+C. C. Montanari, J. E. Miraglia, and N. R. Arista, 
+Phys. Rev. A \textbf{62}, 052902 (2000).
+
+\bibitem{Hf_arxiv} 
+A.M.P. Mendez, C.C. Montanari, D.M. Mitnik, 
+\textit{Slater-type orbital expansion of neutral hafnium, numerical 
+solution of the relativistic Dirac equation}, 
+available soon in arXiv.org.
+
+\bibitem{badnell97} 
+N.R. Badnell, 
+Comput. Phys. Commun., \textbf{182}, 1528-1535 (2011).
+
+\bibitem{williams1995} 
+G. Williams in http://xdb.lbl.gov/Section1/Sec\_1-1.html
+
+\bibitem{lindhard53} 
+J. Lindhard, M. Scharff,  
+Mat. Fys. Medd. Dan. Vid. Selsk  \textbf{27}, 1 (1953).
+
+\bibitem{chu72} 
+W. K. Chu, D. Powers, 
+Rev. Lett. A \textbf{40}, 23 (1972).
+
+\bibitem{mon09} 
+C.C. Montanari, D.M. Mitnik, C.D. Archubi, and J.E. Miraglia, 
+Phys. Rev. A \textbf{80}, 012901 (2009).
+
 \end{thebibliography}

### 2020-04-20 response.tex
checked with grammarly

@@
 \usepackage{xcolor}
+\usepackage[margin=0.8in]{geometry}
+\usepackage{graphicx}
 
@@
 \title{Response to referee reports \\ AP11756/Montanari }
-\author{C.C. Montanari, A. M. P. Mendez, D. M. Mitnik, J. E. Miraglia, \\
-P.A. Miranda, R. Correa, J. Wachter, M. Aguilera, \\
-E. Alves, N. Catarino and R.C. da Silva}
-\date{\today}
+%\author{C. C. Montanari, A. M. P. Mendez, D. M. Mitnik, J. E. Miraglia, \\ P.A. Miranda, R. Correa, J. Wachter, M. Aguilera, \\ E. Alves, N. Catarino and R.C. da Silva}
+%\date{\today}
 
@@
 \vspace{0.1cm}
-The title has been modified following the referee's recommendation to
-``Stopping power of hafnium by proton impact and the importance of 
-relativistic $4f$ electrons''.
+{\color{red} The title has been modified following the referee's recommendation:}
+{\small ''The stopping power of hydrgen in hafnium and the importance of relativistic $4f$ electrons''}.
 
 % ----------------------------------------------------------------------
-\vspace{0.25cm}
+\vspace{0.35cm}
 \textsl{2. It appears that calculations assume bare protons. According 
@@
 \vspace{0.1cm}
-{\color{red}Reply here.}
+{\color{red} It is correct, our calculations assume bare protons.
+We consider that there is no bound electron or very small probability because free electrons of metals screens the hydrogen and this screening is enough at low energies as to have no bound electron.
+CasP includes a very good fitting of experimental charge states and a powerful scaling rule about them.
+However, it is based on measurements of charge states outside the metal.
+For multi-charged ions almost no differences are expected and CasP charge state works very well.
+For H and He ions, differences are important.
+We add the following explanation about why considering just proton impact, and the reference to CasP and the paper of charge states by Schwietz and Grande.}
 
+\vspace{0.1cm}
+{\small ''Finally, in all our calculations we assumed the projectile to be proton and not neutral hydrogen.
+When an ion moves inside a metal, the FEG screens the nucleus, so the binding energies will be smaller than outside the metal, and this effect is more important a low impact velocities $v$.
+In the case of hydrogen the difference is drastic, i.e.
+for H inside Hf ($rs=2.07$), the $1s$-bound state is almost null at $v<2$.
+It is worth to mention that this assumption agrees with Ziegler SRIM code, but differs from CasP code that predicts neutral hydrogen at very low velocities.''}
 % ----------------------------------------------------------------------
-\vspace{0.25cm}
+
+\vspace{0.35cm}
 \textsl{3. Thickness measurements by Rutherford backscattering rely 
@@
 \vspace{0.1cm}
-{\color{red}Reply here.}
+{\color{red} Pedro Miranda ..........................................................}
+\vspace{0.2cm}
 
@@
 \vspace{0.1cm}
-{\color{red}Reply here.}
+{\color{red} The referee is correct, we have used the stopping of 5.486 MeV alphas and SRIM as reference to get the thickness.
+Although SRIM fits the experimental data available, it is not just a fitting, it is a mix of model and fitting (semiempirical).
+At high impact energies where the stopping is well known srim is quite reliable.
+Figure \ref{HeHf} here can clarify these assertions.
+a) In this figure we plotted together H and He in Hf, dividing by $Z^2$.
+Above 1 MeV/amu both srim curves and our theoretical results agree within $2\%$.
+The vertical pink line indicates the impact energy we use as reference.
+It is in the region where the srim predictions are reliable.
+b) srim is semiempirical, the comparison of H and He in the same target shows clearly how srim for H ions curves down and up to fit Sirotonin data.
+It is also clear how srim for He ion overvalue in order to fit the single measurement by Chu, Ziegler and others.
+Why do we say "overvalue"?
+Because the stopping of He ions divided by 4 can be equal or below the stopping of H ions.
+}
+
+%-----------------------------------------------------------------------
+\begin{figure}[!t]
+\centering
+\includegraphics[width=12.cm]{Fig_HeHf.eps}
+\caption{Comparison of stopping of Hf for H and He at high impact energies}
+\label{HeHf}
+\end{figure}
+%-----------------------------------------------------------------------
 
 % ----------------------------------------------------------------------
-\vspace{0.25cm}
-\textsl{5. In connection with the discussion of relativity and 
-outer-shell electrons (p.3) it might be instructive to mention that 
-this is the reason for the existence of relativistic quantum chemistry 
-programs. Since the authors find it surprising that the effect increases 
-in magnitude from inner to outer shells, the reader needs an 
-explanation. I should think the effect of deviations in the screening 
-add up shell by shell, but I may be wrong. Moreover, the sign of the 
-discrepancy in figure 2a deserves attention, which seems to change from 
-positive up to 4p- to negative from 5p+.}
+\vspace{0.35cm}
+\textsl{5. In connection with the discussion of relativity and outer-shell electrons (p.3) it might be instructive to mention that 
+this is the reason for the existence of relativistic quantum chemistry programs.
+Since the authors find it surprising that the effect increases in magnitude from inner to outer shells, the reader needs an explanation.
+I should think the effect of deviations in the screening add up shell by shell, but I may be wrong.
+Moreover, the sign of the discrepancy in figure 2a deserves attention, which seems to change from positive up to 4p- to negative from 5p+.}
 
 \vspace{0.1cm}
-The referee is correct. We included the following modifications in the text
-(lines 200-214):
+{\color{red} The referee is correct. We included the following modifications in the text
+(page 3, lines 210-228):}
 
 \vspace{0.1cm}
-{\small ``One would expect the relativistic contributions to only affect 
-significantly the inner shells. The relativistic binding energies of the 
-electrons belonging to these shells are 2.5\% from the experimental values,
-while the non-relativistic ones are within $\sim$10\%. Moreover, we notice 
-the relativistic corrections greatly influences the outer $5p$ and $4f$ 
-shells. This effect may be a consequence of the orthogonality with the 
-relativistic inner shells. Also, the sign of the binding energy deviations 
-is inverted near the FEG limit, i.e. the electrons go from being more 
-bounded --than the experimental values-- to being less bounded. 
-This change of sign can be attributed to the fact that the experimental 
-values correspond to hafnium in solid-state, while our theoretical 
-calculations correspond to the element in the gas phase.''}
+{\small ``In the comparison with the experimental values in Fig.
+2 (a) it can be noted that the sign of the binding energy deviations is inverted for the outer $5s$-$4f\pm$ electrons, with the experimental binding energies being less bounded than our theoretical ones.
+Small differences for the outer shells are expected since  the experimental
+values correspond to hafnium in solid-state, while our theoretical  calculations correspond to the element in the gas phase.
+For more detail, Fig. 2 (b) shows  the relative errors
+with respect to the experimental  values.
+This figure shows clearly that the relativistic corrections are very important to describe
+the atomic structure of hafnium, even for the outer shells.
+It turns out that the errors committed in the non-relativistic calculations of the inner shell orbitals 
+propagate, through the Hartree-Fock approximation, to the outer shells.The importance of full-relativistic calculations for the outer shells has already been noted before for Au, Pb and Bi and W.''}
+
+%{\small ``One would expect the relativistic contributions to only affect  significantly the inner shells.
+The relativistic binding energies of the  electrons belonging to these shells are 2.5\% from the experimental values, while the non-relativistic ones are within $\sim$10\%.
+Moreover, we notice the relativistic corrections greatly influences the outer $4f$ shell.
+This effect may be a consequence of the orthogonality with the  relativistic inner shells.
+The relativistic $1s$-$4d$ electrons are little but more bounded than the experimental values.
+This sign is inverted for the outer electrons, being less bounded experimentally than the relativistic calculations.
+This change of sign can be attributed to the fact that the experimental values correspond to hafnium in solid-state, while our theoretical calculations correspond to the element in the gas phase.''}
 
 % ----------------------------------------------------------------------
-\vspace{0.25cm}
-\textsl{6. The importance of 4f-5p screening is documented clearly 
-(p.3), but the reader will ask why this does not affect other shells. 
-I assume that the effect is important when two or more subshells have 
-similar energies and, hence, overlap in real and velocity space. If so, 
-I suggest to mention this. If not, another explanation is needed.}
+\vspace{0.35cm}
+\textsl{6.
+The importance of 4f-5p screening is documented clearly (p.3), but the reader will ask why this does not affect other shells.
+I assume that the effect is important when two or more subshells have similar energies and, hence, overlap in real and velocity space.
+If so, I suggest to mention this.
+If not, another explanation is needed.}
 
 \vspace{0.1cm}
-{\color{red}Reply here.}
+{\color{red} The mixing criteria is explained on page 3: we consider the quantum uncertainty in energy so bound electrons with similar energies and mean radio $<r>$ response as a single density of electrons including screening among them.
+We recognize that a more detailed study is needed, we are developing such work as a systematic for different targets involving 4f electrons.
+What we have done here for Hf is to check the importance of the contribution to the total stopping.
+Only the 5p-4f mixing have influence in the total stopping because they are the second most important contribution after the FEG.
+Some other shells may be mixed, and we calculate them, but the total stopping changes in very small $\%$.
+We added lines 262-264 to page 4.}
+
+{\small '' As already mentioned, at higher energies inter-shell screening is possible for other subshells (i.e.
+4p-4d for impact energies above 0.9 MeV), but its weight in the total stopping is minor for deeper shells.'' }
 
 % ----------------------------------------------------------------------
-\vspace{0.25cm}
+\vspace{0.35cm}
 \textsl{7. In view of the large contribution of the outermost shell, I 
@@
 \vspace{0.1cm}
-{\color{red}Reply here.}
+{\color{red} We thank the referee for this suggestion, actually we had this figure, but we doubted on including all the curves.
+Now Fig.
+3 includes separate FEG and bound electron contribution, and the total curves too.}
 
 % ----------------------------------------------------------------------
-\vspace{0.25cm}
+\vspace{0.35cm}
 \textsl{8. In addition to comparisons with experiments and empirical 
@@
 \vspace{0.1cm}
-{\color{red}Reply here.}
+{\color{red} We thank this suggestion.
+We had already compared with Casp5.2 and DPASS for ourselves before submitting.
+In this version we added CasP5.2 and DPASS to figure 4, we added a mention in the introduction(last paragraph, lines 68-73), and the following comment in page 6, lines 315-322:}
+
+{\small ''The stopping maximum is a very sensitive region for any full theoretical description, with no parameters at all, and this is quite visible in a linear-scale plot like Fig.
+[4].
+However, the impact energy for the maximum seem to agree between our curve and DPASS, though $10 \%$ below.
+Instead, CasP predicts a stopping maximum $10 \%$ above and a completely different shape at lower energies.''}
 
@@
 \vspace{0.1cm}
-{\color{red}Reply here.}
+{\color{red} Pedro Miranda, ..........................................................}
+\vspace{0.2cm}
 
@@
 
-\vspace{0.1cm}
-{\color{red}Reply here.}
+{\color{red} There is a small systematic difference between our experiments and theory for impact energies above 1.5 MeV, but it is within the experimental error.
+Figure~\ref{highE} here, amplifies the results shown in the manuscript by focusing  on the high energy region.
+Note that we included our theoretical results, Srim, Casp5.2 and DPASS to this figure.
+The difference is actually small, being less than $3\%$ for the two values above 2 MeV.
+We included the following sentence in the text........................Pedro, PLEASE CHECK THIS}
+
+%-----------------------------------------------------------------------
+\begin{figure}[!t]
+\centering
+\includegraphics[width=12.cm]{Fig04.eps}
+\caption{stopping at high impact energies, theoretical curves and symbols for data explained inside the figure}
+\label{highE}
+\end{figure}
+%-----------------------------------------------------------------------
 
@@
 \vspace{0.1cm}
-{\color{red}Reply here.}
+{\color{red} We thank the referee for this suggestion that improves this new version.
+We have added CasP5.2 and DPASS curves in Figure 4, we have also added a mention in the introduction (last paragraph, lines 68-73), and the following comment in page 6, lines 315-322:}
+
+{\small ''The stopping maximum is a very sensitive region for any full theoretical description, with no parameters at all, and this is quite visible in a linear-scale plot like Fig.
+[4].
+However, the impact energy for the maximum seem to agree between our curve and DPASS, though $10 \%$ below.
+Instead, CasP predicts a stopping maximum $10 \%$ above and a completely different shape at lower energies.''}
 
@@
 \vspace{0.1cm}
-{\color{red}Reply here.}
+{\color{red} We agree with this comment by the referee.
+In a ion-target system with only two set of measurements more data is needed.
+In the energy region 80-500 keV only the data by Sirotinin group is available and cannot be compared with others.
+SRIM is not a test for this data, SRIM includes this data in the adjust (more detail about this can be found in the answer 4 to the first referee report).
+We change the comments in page 6 lines 328-336 and lines 364-366 as follows:}
+
+{\small ''Future experiments  would be important to
+complete this study on the stopping of Hf for protons.
+In the energy region $80-500$ keV only the measurements by Sirotinin group in 1984 are available.
+SRIM clearly includes these values in the adjust of the code, so it does not represent a test for this data.
+Undoubtedly, new measurements are needed in three specific energy regions: below 500 keV, around the stopping maximum (i.e.
+$50-200$ keV) and at low impact energies''}
+
+{\small ''Future experiments for impact energies 
+below 500 keV, around the stopping maximum and in the low energy region would be important to complete this study.''}
 
@@
 \vspace{0.1cm}
-Our calculations are performed for gas phase Hf atoms. We have clarify
-this in the manuscript (lines 211-214).
+{\color{red} The referee is correct, our calculations are performed for gas phase Hf atoms.
+We have added the following paragraph in this version of the manuscript (page 3, lines 207-215) to express this clearly.}
+
+{\small ''In the comparison with the experimental values in Fig.2 (a) it can be noted that the sign of the binding energy deviations is inverted for the outer $5s$-$4f\pm$ electrons, with the experimental binding energies being less bounded than our theoretical ones.
+Small differences for the outer shells are expected since
+the experimental values correspond to hafnium in solid-state, while our theoretical  calculations correspond to the element in the gas phase.''}
 
@@
 \vspace{0.1cm}
-{\color{red}Reply here.}
+{\color{red} Thank you for the comment that helps us to clarify this point.
+Our SLPA calculations use the relativistic wave functions and binding energies for the bound electron contribution to the stopping power.The three curves in Fig.
+3 and the curve in Fig.
+4 are relativistic in this sense.
+We have added an explicit mention about the use of relativistic wave functions and binding energies in page 4 lines 200-201, and also this comment in page 4, lines 287-289:}
+
+{\small ''Bound $1s$-$4f$ electrons (relativistic wave functions and binding energies) are calculated with the SLPA and shown separately in Fig. 3''}
 
@@
 \vspace{0.1cm}
-{\color{red}Te mando adjunto las funciones de onda y energías de ligaduras
-del caso no relativista (NR). Habría que ver como queda.}
+{\color{red}  The SLPA calculations for the bound electrons uses the Hartree-Fock wave funtions and binding energies, which are available in the tables by Clementi-Roetti or by Bunge for elements up to Xe (Z=54).
+The errors in the non-relativistic calculations worsen for higher Z atoms, and this is the reason why these tables do not extend any further than Z=54.
+The fail of non-relativistic values of binding energies is shown in Fig.
+2.
+We have added a comment on this in page 3, lines 214-228.
+}
+
+\vspace{0.1cm}
+{\small ``For more detail, Fig. 2 (b) shows  the relative errors
+with respect to the experimental  values.
+This figure shows clearly that the relativistic corrections are very important to describe
+the atomic structure of hafnium, even for the outer shells.
+It turns out that the errors committed in the non-relativistic calculations of the inner shell orbitals 
+propagate, through the Hartree-Fock approximation, to the outer shells.The importance of full-relativistic calculations for the outer shells has already been noted before for Au, Pb and Bi and W.''}
+
+{\color{red} When dealing with stopping in targets with Z>54 we need to face our own calculation of the wave functions and binding energies.
+And we do it within the relativistic approximation to the best of our possibilities.
+However, one can wonder if such discrepancy in the non-relativistic results affects the stopping power.
+First, it will only affect the bound electron contribution to the total stopping.
+Second, at high energies (i.e.
+1 MeV) the results converge to Bethe limit for any density of electrons as far as the correct number of target electrons is included.
+Having clarified this, we include in the figure \ref{NR} here, the total stopping using the non-relativistic values we calculate using the autostructure code.
+As can be noted, at sufficiently high energies it is the same, but differs around the stopping and at lower impact energies.
+We have included in this figure and in Fig.
+4 of the manuscript the DPASS results by Sigmund.}
+
+%-----------------------------------------------------------------------
+\begin{figure}[!t]
+\centering
+\includegraphics[width=12.cm]{Fig05.eps}
+\caption{Total stopping, present results using relativistic and non-relativistic solutions for hafnium atom. Curves and symbols explained inside the figure}
+\label{NR}
+\end{figure}
+%-----------------------------------------------------------------------
 
@@
 \vspace{0.1cm}
-{\color{red}Reply here.}
+{\color{red} Thank you for this suggestion. We have added the curves for FEG and bound electrons to Fig. 3, and the corresponding comments in the text and in the figure caption.}
 
@@
 \vspace{0.1cm}
-{\color{red}Reply here.}
+{\color{red} Good suggestion, thank you.
+Sirotinin group used the backscattering method for these measurements.
+We have added a sentence in the introduction, last paragraph, and in the second paragraph of section IV related to Fig.
+4.}
+
+{\small ''Present measurements using transmission method are in good agreement with the previous data by 
+Sirotinin, measured in backscattering geometry''}
 
@@
 \vspace{0.1cm}
-{\color{red}Reply here.}
+{\color{red} The stopping maximum is very sensitive to the theoretical model and to the experimental method.
+Sirotinin paper mention "possible reasons for the disagreement of the measurement results near the E(E) curve maximum".
+%The original paper by Sirotinin says: "The possible reasons for the disagreement of the measurement results near the E(E) curve maximum are mainly related to the inaccuracy in determining thickness, to nonuniformity of the thickness, and the neglect of the effect of the chemical composition and structure of the thin films used in the measurements"
+For targets with many different sets of data (i.e.
+H in Ag, Al, Au, Fe, Cu, Si, and others) it is usual to note great dispersion of values around the maximum.
+In the case of Hf, only one set of data is available and it is perfectly matched by SRIM semi-empirical curve.
+SRIM clearly includes these values in the adjust of the code, so it does not represent a test for this data.
+One of the goals of this work is to call the attention on possible differences between theory and the existing values, and to remark the necessity of more experimental data in certain energy regions, mainly, the stopping maximum and below.
+We have changed the comments in page 6 lines 328-336 and lines 364-366 as follows:}
+
+{\small ''Future experiments  would be important to
+complete this study on the stopping of Hf for protons.
+In the energy region $80-500$ keV only the measurements by Sirotinin group in 1984 are available.
+SRIM clearly includes these values in the adjust of the code, so it does not represent a test for this data.
+Undoubtedly, new measurements are needed in three specific energy regions: below 500 keV, around the stopping maximum (i.e.
+$50-200$ keV) and at low impact energies''}
+
+{\small ''Future experiments for impact energies around the stopping maximum and in the low energy region would be important to complete this study.''}
+
+{\color{red} In relation with the referee comment about the maximum of the stopping, in Fig.
+4 we have added the comparison with other two theoretical results: CASP and PASS.
+The following comment has been included in page 6, lines 315-322:}
+
+{\small ''The stopping maximum is a very sensitive region for any full theoretical description, with no parameters at all, and this is quite visible in a linear-scale plot like Fig.
+[4].
+However, the impact energy for the maximum seem to agree between our curve and DPASS, although DPASS is $10 \%$ below ours.
+Instead, CasP predicts a stopping maximum $10 \%$ above our value and a completely different shape at lower energies.''}
 
@@
 \vspace{0.1cm}
-We thank the referee for suggesting the missing reference. We have now
-include it in the revised manuscript.
+{\color{red} We thank the referee for suggesting this missing reference. We have
+included it in the revised manuscript.}
 
@@
 \textsl{- Y-axis labels in Figs. 3 an 4 are unreadable }
-{\color{red} ¿Cómo quedó el preprint? El pdf que genera mi latex tiene 
-bien los labels, así que debe un problema del generador de latex del PRA...}
 
-\st{\textsl{- Page 3,2nd column, 2nd full paragraph: "Too assess"}}
+{\color{red} We have changed to larger prints  and have written explicitly "stopping cross section" -Y- and "Impact energy" -X- instead of S(E) and E. }
 
-\st{\textsl{- Caption of Fig.2: "hollow circles" -$>$ "open circles" or
-"unfilled circles"}}
+\textsl{- Page 3,2nd column, 2nd full paragraph: "Too assess"}
+{\color{red} corrected}
+
+\textsl{- Caption of Fig.2: "hollow circles" -$>$ "open circles" or
+"unfilled circles"}
+{\color{red} corrected}
 

### 2020-04-20 Stopping_Hf_abril20.tex
checked with grammarly

--- 2020-04-11 sp_hafnium.tex
@@
-\documentclass[aps,pra,reprint,groupedaddress,showpacs,showkeys]{revtex4-1}
+\documentclass[aps,pra,reprint,superscriptaddress]{revtex4-1}
+%\documentclass[aps,pra,reprint,groupedaddress,showpacs,showkeys]{revtex4-1}
 \usepackage{amssymb}
@@
 
-%\def\clau#1{\textcolor{ForestGreen}{#1}}
-%\def\minor#1{\textcolor{green}{#1}}
+\def\clau#1{\textcolor{red}{#1}}
+\def\duda#1{\textcolor{green}{#1}}
+\def\ale#1{\textcolor{blue}{#1}}
 
@@
 %Title of paper
-\title{Stopping power of hafnium by proton impact and \\
-the importance of relativistic $4f$ electrons}
+\title{The stopping power of \clau{hydrogen in} hafnium and the importance of relativistic $4f$ electrons}
 %Authors and affiliations
@@
 Argentina}
-%------------------------------
+
+%---------------------------Chile
+\author{\clau{P. A. Miranda}}
+\affiliation{Departamento de F\'isica, Facultad de Ciencias Naturales, 
+Matem\'atica y del Medio Ambiente. Universidad Tecnol\'ogica Metropolitana. 
+7800002, Chile}
+%-------------------------Lisboa
+\author{\clau{E. Alves}}
+\affiliation{Centro de Ci\^{e}ncias e Tecnologias Nucleares, Instituto 
+Superior T\'ecnico, Universidade de Lisboa, 2696-953 Sacav\'{e}m, 
+Portugal}
+\affiliation{Instituto de Plasma e Fus\~{a}o Nuclear, Instituto Superior 
+T\'ecnico, Universidade de Lisboa, 2696-953 Sacav\'{e}m, Portugal}
+
+%------------------------------Argentina
 \author{A. M. P. Mendez}
@@
 Argentina}
-%---------------------------Chile
 
-\author{P.A. Miranda}
-\affiliation{Departamento de F\'isica, Facultad de Ciencias Naturales, 
-Matem\'atica y del Medio Ambiente. Universidad Tecnol\'ogica Metropolitana. 
-7800002, Chile}
-%----------------------------
+%---------------------------- Chile
 \author{R. Correa}
@@
 7800002, Chile}
+
 %-------------------------Lisboa
-\author{E. Alves}
-\affiliation{Centro de Ci\^{e}ncias e Tecnologias Nucleares, Instituto 
-Superior T\'ecnico, Universidade de Lisboa, 2696-953 Sacav\'{e}m, 
-Portugal}
-\affiliation{Instituto de Plasma e Fus\~{a}o Nuclear, Instituto Superior 
-T\'ecnico, Universidade de Lisboa, 2696-953 Sacav\'{e}m, Portugal}
-%---------------------------
 \author{N. Catarino}
@@
 %---------------------------
-\author{R.C. da Silva}
+\author{R. C. da Silva}
 \affiliation{Centro de Ci\^{e}ncias e Tecnologias Nucleares, Instituto 
@@
 while more attention has been recently given to  studies involving 
-hafnium oxide due to its practical use~\cite{Abril,Behar,Primetzhofer,Roth}. 
+hafnium oxide due to its practical use~\ale{\cite{Abril,Behar,Primetzhofer,Roth}}. 
 It is well known that significant attention has been paid in recent 
@@
 
-The theoretical approach implemented in this work uses the shell-wise 
-local plasma approximation (SLPA)~\cite{mon13} to describe the energy 
+The theoretical approach implemented in this work uses the shell-wise local plasma approximation (SLPA)~\cite{mon13} to describe the energy 
 transferred to the bound $1s$-$4f$ electrons, and two different models 
-for the FEG: the screened potential with cusp condition model (SPCC)~\cite{mon17}, 
+for the FEG: \clau{in low energy region,}
+the screened potential with cusp condition model (SPCC)~\cite{mon17}, 
 which is non-linear binary formalism; and the Mermin-Lindhard dielectric 
@@
 while the theoretical method is explained in Section~\ref{theory}. 
-Present theoretical and experimental values are finally compared to the 
-available experimental values~\cite{Sirotinin}, the ICRU-49 
-calculations~\cite{ICRU49} and the SRIM-2013 package~\cite{Ziegler01}. 
+Present theoretical and experimental values are finally compared to the \clau{only experimental values by Sirotinin group~\cite{Sirotinin} measured with backscattering method, the theoretical results by  Grande and Schiwietz~\cite{Grande,casp52}, and by Sigmund and Schinner~\cite{DPASS20}, and also with the semi-empirical values from the SRIM-2013 package~\cite{Ziegler01} and the ICRU-49 tabulation~\cite{ICRU49}}.
 Conclusions and discussions are given in Section~\ref{conclusion}. All 
@@
 not only binary but also collective excitations (plasmons) 
-occur~\cite{mon17}. Moreover, at high energies, also the bound electrons 
+\ale{occur}~\cite{mon17}. Moreover, at high energies, also the bound electrons 
 contribute to the stopping power. The method used in this work combines 
@@
 Hafnium ($Z=72$, [Xe] $4f^{14}\,6s^2\,5d_{3/2}^1\,5d_{5/2}^1$), belongs 
-to the first groups of transition metals, with four electrons as FEG 
-(theoretical $r_S=2.14$ a.u.) and $1s$-$4f$ electrons bound. We compared 
+to the first groups of transition metals, with \ale{four} electrons as FEG 
+(\st{theoretical} $r_S=2.14$ a.u.) and $1s$-$4f$ electrons bound. We compared 
 the computed $r_S$ with the experimental value obtained from measured 
@@
 \includegraphics[width=11.cm]{bindener.eps}
-\caption{(a) Binding energies of Hf. Non-relativistic and relativistic 
-calculations are given with filled symbols. Experimental measurements 
+\caption{(a) Binding energies of Hf. \clau{Present relativistic and available non-relativistic~\cite{badnell97}  
+values} are given with filled symbols. Experimental measurements 
 for solids~\cite{williams1995} are depicted with open circles. 
@@
 For the stopping power due to bound electrons, the SLPA~\cite{mon17,mon13} 
-is employed. It is worth mentioning that the only inputs for the SLPA 
+is employed. It is worth \ale{mentioning} that the only inputs for the SLPA 
 are the space-dependent densities of each shell in the ground state, 
 and their binding energies. Collective processes and screening among 
-electrons are included. Since hafnium is a relativistic target, the 
+electrons are included. \ale{Since} hafnium is a relativistic target, the 
 wave functions and binding energies must be obtained by solving the 
-many-electron Dirac Hamiltonian. Details of these calculations have been 
-published in \cite{mendez2019}, while Slater-type orbital expansions are 
+\ale{many-electron} Dirac Hamiltonian. Details of these calculations 
+\clau{and a table of binding energies} have been published in 
+Ref.~\cite{mendez2019}, while Slater-type orbital expansions are 
 given in \cite{Hf_arxiv}.
 
-To assess the importance of the description of bound electrons, 
-Fig.~\ref{Binding_E} shows (a) non-relativistic~\cite{badnell97} and 
-our relativistic binding energies $E_{nl\pm}$, with $\pm=j\pm1/2$; and 
-(b) the relative errors with respect to the experimental data on solid 
-Hf~\cite{williams1995} from $1s$ to $4f_{\pm}$. 
-%\st{We notice that not only the most inner shells require relativistic calculations, but also the outer $5p$ and $4f$ shells.
-Furthermore, this figure shows very clearly the disability of non-relativistic calculations to describe the experimental data, which surprisingly worsens from the inner to the outer shells.}
-One would expect the relativistic contributions to only affect 
-significantly the inner shells. The relativistic binding energies of the 
-electrons belonging to these shells are 2.5\% from the experimental values,
-while the non-relativistic ones are within $\sim$10\%. Moreover, we notice 
-the relativistic corrections greatly influences the outer $5p$ and $4f$ 
-shells. This effect may be a consequence of the orthogonality with the 
-relativistic inner shells. Also, the sign of the binding energy deviations 
-is inverted near the FEG limit, i.e. the electrons go from being more 
-bounded --than the experimental values-- to being less bounded. 
-This change of sign can be attributed to the fact that the experimental 
-values correspond to hafnium in solid-state, while our theoretical 
-calculations correspond to the element in the gas phase.
+\ale{To} assess the importance of  \clau{a full relativistic} description of 
+bound electrons, Fig.~\ref{Binding_E} (a) shows our \st{relativistic} binding 
+energies, $E_{nl\pm}$, with $\pm=j\pm1/2$; the 
+non-relativistic values~\cite{badnell97}; \clau{and the experimental data on 
+solid Hf~\cite{williams1995}, which is available only for $1s$ to $4f_{\pm}$ 
+subshells, as expected.} 
+We notice that not only the most inner shells require relativistic calculations, but also the outer $5p$ and $4f$ shells.
+Furthermore, this figure shows very clearly the disability of non-relativistic calculations to describe the experimental data, which surprisingly worsens from the inner to the outer shells.
 
-Our relativistic binding energies present spin-orbit split. However, in 
-total stopping power, where the initial state of the excited electron is 
-not measured, the quantum uncertainty in energy $\Delta E$ melts this 
-split. The criteria $\Delta E\Delta t\geq\hbar/2$ merges the energies 
+\clau{In the comparison with the experimental values in Fig.~\ref{Binding_E} (a) it can be noted that}
+\ale{\st{Also,} the sign of the binding energy deviations is inverted}  \clau{for the outer $5s$-$4f\pm$ electrons, with the experimental binding energies being less bounded than our theoretical ones.
+Small differences for the outer shells are expected since }
+\ale{\st{can be attributed to the fact that} the experimental
+values correspond to hafnium in solid-state, while our theoretical  calculations correspond to the element in the gas phase.}
+%i.e. the electrons go from being more bounded --than the experimental values-- to being less bounded.
+
+\clau{For more detail, Fig.~\ref{Binding_E} (b) shows}  the relative errors %\clau{of theoretical (relativistic and non relativistic)} 
+with respect to the experimental  values.
+%\ale{\st{The relativistic binding energies of the electrons belonging to the deep K, L and M shells deviate less than $2.5\%$ from the experimental values, while the non-relativistic ones are within $\sim10\%$.
+Moreover, we notice that the relativistic corrections greatly influences the outer shells.
+This effect may be consequence of the orthogonality with the relativistic inner shells.}}
+%\clau{ \st{This figure shows clearly that the relativistic corrections are decisive to describe the outer shells.
+The solution of the Dirac equation from inner to outer shells, and the orthogonality of the wave functions, involves all the shells.
+}}
+\duda{This figure shows clearly that \st{for this target,} the relativistic corrections are very important to describe
+the atomic structure of hafnium, even for the outer shells.
+It turns out that the errors committed in the non-relativistic calculations of the inner shell orbitals 
+propagate, through the Hartree-Fock approximation, to the outer shells.} \clau{The importance of full-relativistic calculations for the outer shells has already been noted for Au, Pb and Bi and W~\cite{mon09}.}
+
+\clau{For the contribution of bound electrons to the total stopping cross sections}, the SLPA considers independent contributions of each subshell.
+Our relativistic binding energies present spin-orbit split.
+However, in total stopping power, where the initial state of the excited electron is not measured, the quantum uncertainty in energy $\Delta E$ melts this split.
+The criteria $\Delta E\Delta t\geq\hbar/2$ merges the energies
 $E_{nl+}-E_{nl-}$ for sufficiently small values of $\Delta t$ (the 
@@
 we can expect all target electrons to respond together to the ion 
-passage~\cite{lindhard53,chu72}. Following previous work~\cite{mon09}, 
+passage~\cite{lindhard53,chu72}. Following previous works~\cite{mon09},
 the collisional time is estimated as $\Delta t\approx\langle r_i\rangle/v$, 
@@
 impact velocity, respectively. 
-
-%-----------------------------------------------------------------------
-\begin{figure*}[!t]
-\centering
-\includegraphics[width=13.cm]{Fig02.eps}
-\caption{Theoretical total stopping cross sections of protons in hafnium 
-adding FEG and bound $1s$-$4f$ contributions, as mentioned in the inset. 
-The vertical grey dashed-line indicates the energy above which plasmon 
-excitation is possible. The SLPA total stopping values for $1s$-$4f$ 
-bound electrons with and without $5p$-$4f$ screening are displayed with 
-solid and dotted lines, respectively.}
-\label{slpa4f}
-\end{figure*}
-%-----------------------------------------------------------------------
-
-The SLPA calculates the contribution of each subshell of bound electrons 
-to the total stopping cross sections. We found that for every sub-shell 
-of Hf, at the impact energy at which this sub-shell began to contribute, 
-the spin-orbit split was unresolved. Then, the $nl$-electrons should be 
+\clau{In the case of hafnium}, we found that for every sub-shell of electrons, the spin-orbit split is unresolved in the energy region this sub-shell contributes.
+Therefore, the $nl$-electrons should be
 considered together, responding to the ion passage as a single gas of 
@@
 together for impact energies $E>40$ keV (inter-shell screening).
+\clau{As already mentioned, at higher energies inter-shell screening is possible for other subshells (i.e.
+4p-4d for impact energies above 0.9 MeV), but its weight in the total stopping is minor for deeper shells.}
 
@@
 \centering
-\includegraphics[width=13.0cm]{Fig03.eps}
-\caption{Stopping power cross section of hafnium for protons. Symbols: 
+\includegraphics[width=13.cm]{Fig02_new3.eps}
+\caption{(colour online) Theoretical stopping cross sections of protons in hafnium.
+\clau{Blue dash-dotted-line,the non-perturbative SPCC for the FEG; blue solid-line, ML results for the FEG (includes plasmon excitation); red
+solid and dotted-lines, the SLPA results for bound electrons with and without $5p$-$4f$ screening, respectively. Black curves, total  stopping
+adding the FEG and bound $1s$-$4f$ contributions: dash-dotted-line, SPCC (FEG) + SLPA (bound); solid-line, ML (FEG) + SLPA (bound) with 4f-5p screening; dotted-line, ML (FEG) + SLPA (bound) without 4f-5p screening.}
+The vertical grey dashed-line indicates the energy \clau{of 37 keV} above which plasmon 
+excitation is possible.} 
+\label{slpa4f}
+\end{figure*}
+%-----------------------------------------------------------------------
+
+\clau{Finally, in all our calculations~\cite{mon17} we assumed the projectile to be proton and not neutral hydrogen.
+When an ion moves inside a metal, the FEG screens the nucleus, so the binding energies will be smaller than outside the metal, and this effect is more important a low impact velocities $v$.
+In the case of hydrogen the difference is drastic, i.e.
+for H inside Hf ($rs=2.07$), the $1s$-bound state is almost null at $v<2$~\cite{suppression}.
+It is worth to mention that this assumption agrees with Ziegler SRIM code~\cite{Ziegler01}, but differs from CasP code~\cite{Grande} that predicts neutral hydrogen at very low velocities \st{based on measurements of charge states outside the metal}.}
+
+%-----------------------------------------------------------------------
+\begin{figure*}[!t]
+\centering
+\includegraphics[width=13.0cm]{Fig03_new2.eps}
+\caption{(colour online) Stopping power cross section of hafnium for protons. Symbols: 
 solid circles, present values; open circles, previous data~\cite{Sirotinin}. 
 Curves: Black solid-line, present full theoretical results with 
-$4f$-$5p$ screening; violet dashed-line, values from ICRU-49~\cite{ICRU49}, 
-and green dotted line, SRIM-2013~\cite{Ziegler01}.}
+$4f$-$5p$ screening; \clau{pink dash-dot line, theoretical CasP5.2~\cite{Grande,casp52} values; orange dash-double-dot line, theoretical DPASS~\cite{DPASS20} results; } green dotted-line, semi-empirical SRIM-2013~\cite{Ziegler01}; violet dashed-line, ICRU49~\cite{ICRU49} tabulated values.}
 \label{F03}
@@
 In Fig.~\ref{slpa4f}, we display the present theoretical stopping cross
-section of Hf for protons with and without the $5p$-$4f$ screening. We 
-added the FEG and the bound electron contributions as explained above. 
-The minimum energy for plasmon excitation was estimated to $37$ keV. We 
+section of Hf for protons \clau{using the relativistic wave functions and binding energies, but} with and without the $5p$-$4f$ screening.
+\clau{We show separately the FEG and bound electron contributions and the total stopping as the addition of both of them.}
+\st{added the FEG and the bound electron contributions as explained above. }
+The minimum energy for plasmon excitation was estimated $\sim 37$ keV. We
 used the non-perturbative SPCC model for impact energies $E \leq 37$~keV, 
-and the perturbative ML calculation above this energy. Clearly, 
+and the perturbative ML calculation above this energy.
+\clau{Bound $1s$-$4f$ electrons (relativistic wave functions and binding energies) are calculated with the SLPA and shown separately in Fig.~\ref{slpa4f} with and without the $4f$-$5p$ screening.
+Below $\sim 40$ keV the difference between both calculations is negligible}.
+Clearly,
 considering $5p$-$4f$ electrons as a single group of 20 electrons with 
@@
 the separate $5p$ and $4f$ contributions. Notice that this shell 
-correction can only be considered within a many electron model such as 
-the SLPA.
+correction can \st{only} be considered \duda{self-consistently? naturally? intrinsically?} within a many electron model such as the SLPA, .
 
@@
 
-The present data and theoretical results are displayed in 
-Table~\ref{table01}, and Fig.~\ref{F03}. As can be seen in 
-Table~\ref{table01}, an overall relative uncertainty of around 5\% was 
+The present data \st{and theoretical results are} is displayed in 
+Table~\ref{table01}. %, and Fig.~\ref{F03}. 
+As can be \clau{observed,} %seen in Table~\ref{table01},
+an overall relative uncertainty of around 5\% was 
 achieved for the experimental stopping power values, which are mainly 
@@
 \caption{Stopping power values S$_{\mathrm{exp}}$ of hafnium for protons measured 
-in this work. $\Delta$E/E values are also shown.
-\label{table01}}
+in this work. $\Delta$E/E values are also shown.}
+\label{table01}
 
@@
 
-In Fig.~\ref{F03}, we have good agreement between the present 
-theoretical results and the measurements displayed in Table~\ref{table01}. 
-Our theoretical approach also agrees with previous data by 
+Fig.~\ref{F03} \clau{synthesizes the results of the present work.} \st{, we have good} \clau{The} agreement between the present 
+theoretical results and the \clau{new} measurements displayed in Table~\ref{table01} \clau{is clear}.
+\clau{Present measurements using transmission method are in good agreement with the previous data by
+Sirotinin~\cite{Sirotinin}, measured in backscattering geometry}. Our theoretical approach also agrees with the data by 
 Sirotinin~\cite{Sirotinin}, except for the lowest energy measurement at 
-80 keV. It is interesting that our full theoretical curve differs from 
+80 keV.
+\clau{We have also included in this figure the theoretical curves from Grande and Schiwietz CasP5.2 code~\cite{Grande,casp52} and from Sigmund and Schinner DPASS code~\cite{DPASS20}, both available online; also the semi-empirical results from SRIM-2013~\cite{Ziegler01} and the ICRU49 tables~\cite{ICRU49}}.
+It is interesting that our full theoretical curve differs from
 SRIM-2013 for impact energies below 100 keV. We obtain a stopping 
@@
 following the up-to-now only set of data \cite{Sirotinin}, SRIM-2013
-suggests a lower stopping maximum  at impact energy of 115 kev. It is 
-worth mentioning that similar theoretical results using the experimental 
-$r_S=2.07$ a.u. instead of the theoretical $r_S=2.14$ a.u. give the 
-stopping maximum at the same impact energy but $4\%$ higher. Future 
-experiments for impact energies below 100 keV would be important to 
-complete this study.
+suggests a lower stopping maximum  at impact energy of 115 kev. 
+\clau{The stopping maximum is a very sensitive region for any full theoretical description, with no parameters at all, and this is quite visible in a linear-scale plot like Fig.~\ref{F03}.
+However, the impact energy for the maximum seem to agree between our curve and DPASS, although is $10 \%$ below.
+Instead, CasP predicts a stopping maximum $10 \%$ above our value and a completely different shape at lower energies.}
+It is 
+worth mentioning that \clau{our model gives} similar \st{theoretical} results using the experimental 
+$r_S=2.07$ a.u. instead of the theoretical $r_S=2.14$ a.u. \st{give} \clau{with} the 
+stopping maximum at the same impact energy but $4\%$ higher.
+
+Future experiments \st{for impact energies below 100 keV} would be important to
+complete this study \clau{on the stopping of Hf for protons.
+In the energy region $80-500$ keV only the measurements by Sirotinin group~\cite{Sirotinin} in 1984 are available.
+SRIM-2013 clearly includes these values in the adjust of the code, so it does not represent a test for this data.
+Undoubtedly, new measurements are needed in three specific energy regions: below 500 keV, around the stopping maximum (i.e.
+$50-200$ keV) and at low impact energies.
+}
 
@@
 around the stopping maximum and above. Present theoretical stopping 
-cross sections cover an extensive energy range from 1 keV/amu-10 meV/amu.
+cross sections cover an extensive energy range from 1 keV/amu-10 MeV/amu.
 
@@
 energies the difference between our full-theoretical results and SRIM 
-is substantial. To the best of our knowledge, these are the first 
+is substantial.
+\clau{We compare our theoretical results with other two models given by the DPASS and CasP5.2 codes.
+Differences can be noted at intermediate to low impact energies, but  they also support an stopping maximum at lower energy than SRIM predictions}.
+
+To the best of our knowledge, these are the first 
 theoretical calculations of stopping in Hf taking into account the 
-atomic relativistic effect. Future experiments for impact energies 
-below 100 keV would be important to complete this study.
+atomic relativistic effect \clau{and the screening among electrons in a consistent way from very low to high impact energies}. Future experiments for impact energies 
+\clau{around the stopping maximum and in the low energy region} would be important to \st{complete  this study} \clau{have a more complete knowledge of the stopping of protons in hafnium}.
 
@@
 \bibitem{Roth}
-D. Roth, B. Bruckner, G. Undeutsch, V. Paneta, A. I. Mardare, 
+\ale{D. Roth, B. Bruckner, G. Undeutsch, V. Paneta, A. I. Mardare, 
 C. L. McGahan, M. Dosmailov, J. I. Juaristi, M. Alducin, 
 J. D. Pedarnig, R. F. Haglund, Jr., D. Primetzhofer, and P. Bauer
-Phys. Rev. Lett. \textbf{119}, 163401 (2017).
+Phys. Rev. Lett. \textbf{119}, 163401 (2017).}
 
@@
 
-\bibitem{Roth17} 
-D. Roth, B. Bruckner, M. V. Moro, S. Gruber, D. Goebl, J. I. Juaristi, 
-M. Alducin,R. Steinberger, J. Duchoslav, D. Primetzhofer, and P. Bauer, 
-Phys. Rev. Lett. \textbf{118}, 103401 (2017).
+\bibitem{Roth17} D.
+Roth, B.
+Bruckner, M.
+V.
+Moro, S.
+Gruber, D.
+Goebl, J.
+I.
+Juaristi, M.
+Alducin,R.
+Steinberger, J.
+Duchoslav, D.
+Primetzhofer, and P.
+Bauer, Phys.
+Rev.
+Lett.
+\textbf{118}, 103401 (2017).
 
@@
 
-\bibitem{ICRU49} 
-ICRU report 49, 
-Stopping Powers and Ranges for Protons and Alpha Particles, 
-International Commission on Radiation Units and Measurements, 1993.
+\bibitem{Grande} \clau{G. Schiwietz, P. L. Grande, Nucl. Instrum. Meth. B \textbf{175-177}, 125 (2001); CasP code, available from www.casp-program.org}
+
+\bibitem{casp52} \clau{G. Schiwietz and P. L. Grande,
+Nucl. Instr. and Meth. B \textbf{273}, 1-5 (2012); P.L. Grande and G. Schiwietz,
+Phys. Rev. A \textbf{58}, 3796 (1998).}
+
+\bibitem{DPASS20} \clau{A.
+Schinner, P.
+Sigmund, Nucl.
+Instrum.
+Meth.
+B \textbf{460}, 19 (2019); P.Sigmund, A.
+Schinner, Eur.
+Phys.
+J.
+D \textbf{12}, 425 (2000).
+DPASS code, available from https://www.sdu.dk/en/DPASS/}
 
 \bibitem{Ziegler01} 
-J.F. Ziegler, J.P. Biersack, SRIM2013, 
+J.F. Ziegler, J.P. Biersack, M. D. Ziegler, \clau{\textit{SRIM, The Stopping and Range of Ions in Matter}, (SRIM Co. Maryland, USA, 2008);} SRIM2013, 
 Computer Program and Manual. Available from www.srim.org.
+
+\bibitem{ICRU49} 
+ICRU report 49, \textit{Stopping Powers and Ranges for Protons and Alpha Particles},
+International Commission on Radiation Units and Measurements (1993).
 
@@
 \bibitem{badnell97} 
-N.R. Badnell, 
-Comput. Phys. Commun., \textbf{182}, 1528-1535 (2011).
+N. R. Badnell, Comput. Phys. Commun. \textbf{182}, 1528 (2011).
 
@@
 G. Williams in http://xdb.lbl.gov/Section1/Sec\_1-1.html
+
+\bibitem{mon09} 
+C.C. Montanari, D.M. Mitnik, C.D. Archubi, and J.E. Miraglia, 
+\clau{Phys. Rev. A \textbf{70}, 032903 (2009); and} Phys. Rev. A \textbf{80}, 012901 (2009).
 
@@
 
-\bibitem{mon09} 
-C.C. Montanari, D.M. Mitnik, C.D. Archubi, and J.E. Miraglia, 
-Phys. Rev. A \textbf{80}, 012901 (2009).
-
 \end{thebibliography}

### 2020-04-20 april_version/Stopping_Hf_abril20.tex
checked with grammarly

--- 2020-04-11 sp_hafnium.tex
@@
 %Title of paper
-\title{Stopping power of hafnium by proton impact and \\
-the importance of relativistic $4f$ electrons}
+\title{The stopping power of hydrogen in hafnium and the importance of relativistic $4f$ electrons}
 %Authors and affiliations
@@
 Argentina}
-%------------------------------
+
+%---------------------------Chile
+\author{P. A. Miranda}
+\affiliation{Departamento de F\'isica, Facultad de Ciencias Naturales, 
+Matem\'atica y del Medio Ambiente. Universidad Tecnol\'ogica Metropolitana. 
+7800002, Chile}
+%-------------------------Lisboa
+\author{E. Alves}
+\affiliation{Centro de Ci\^{e}ncias e Tecnologias Nucleares, Instituto 
+Superior T\'ecnico, Universidade de Lisboa, 2696-953 Sacav\'{e}m, 
+Portugal}
+\affiliation{Instituto de Plasma e Fus\~{a}o Nuclear, Instituto Superior 
+T\'ecnico, Universidade de Lisboa, 2696-953 Sacav\'{e}m, Portugal}
+
+%------------------------------Argentina
 \author{A. M. P. Mendez}
@@
 while more attention has been recently given to  studies involving 
-hafnium oxide due to its practical use~\cite{Abril,Behar,Primetzhofer,Roth}. 
+hafnium oxide due to its practical use~\cite{Abril,Behar,Primetzhofer,Roth}}. 
 It is well known that significant attention has been paid in recent 
@@
 
-The theoretical approach implemented in this work uses the shell-wise 
-local plasma approximation (SLPA)~\cite{mon13} to describe the energy 
+The theoretical approach implemented in this work uses the shell-wise local plasma approximation (SLPA)~\cite{mon13} to describe the energy 
 transferred to the bound $1s$-$4f$ electrons, and two different models 
-for the FEG: the screened potential with cusp condition model (SPCC)~\cite{mon17}, 
+for the FEG: in low energy region,
+the screened potential with cusp condition model (SPCC)~\cite{mon17}, 
 which is non-linear binary formalism; and the Mermin-Lindhard dielectric 
@@
 Present theoretical and experimental values are finally compared to the 
-available experimental values~\cite{Sirotinin}, the ICRU-49 
-calculations~\cite{ICRU49} and the SRIM-2013 package~\cite{Ziegler01}. 
+only experimental values by Sirotinin group~\cite{Sirotinin} measured 
+with backscattering method, the theoretical results by  Grande and 
+Schiwietz~\cite{Grande,casp52}, and by Sigmund and Schinner~\cite{DPASS20}, 
+and also with the semi-empirical values from the SRIM-2013 
+package~\cite{Ziegler01} and the ICRU-49 tabulation~\cite{ICRU49}}. 
 Conclusions and discussions are given in Section~\ref{conclusion}. All 
@@
 not only binary but also collective excitations (plasmons) 
-occur~\cite{mon17}. Moreover, at high energies, also the bound electrons 
+occur}~\cite{mon17}. Moreover, at high energies, also the bound electrons 
 contribute to the stopping power. The method used in this work combines 
@@
 Hafnium ($Z=72$, [Xe] $4f^{14}\,6s^2\,5d_{3/2}^1\,5d_{5/2}^1$), belongs 
-to the first groups of transition metals, with four electrons as FEG 
-(theoretical $r_S=2.14$ a.u.) and $1s$-$4f$ electrons bound. We compared 
+to the first groups of transition metals, with four} electrons as FEG 
+($r_S=2.14$ a.u.) and $1s$-$4f$ electrons bound. We compared 
 the computed $r_S$ with the experimental value obtained from measured 
@@
 \includegraphics[width=11.cm]{bindener.eps}
-\caption{(a) Binding energies of Hf. Non-relativistic and relativistic 
-calculations are given with filled symbols. Experimental measurements 
+\caption{(a) Binding energies of Hf. Present relativistic and available non-relativistic~\cite{badnell97}  
+values are given with filled symbols. Experimental measurements 
 for solids~\cite{williams1995} are depicted with open circles. 
@@
 For the stopping power due to bound electrons, the SLPA~\cite{mon17,mon13} 
-is employed. It is worth mentioning that the only inputs for the SLPA 
+is employed. It is worth mentioning} that the only inputs for the SLPA 
 are the space-dependent densities of each shell in the ground state, 
 and their binding energies. Collective processes and screening among 
-electrons are included. Since hafnium is a relativistic target, the 
+electrons are included. Since} hafnium is a relativistic target, the 
 wave functions and binding energies must be obtained by solving the 
-many-electron Dirac Hamiltonian. Details of these calculations have been 
-published in \cite{mendez2019}, while Slater-type orbital expansions are 
+many-electron} Dirac Hamiltonian. Details of these calculations 
+and a table of binding energies have been published in 
+Ref.~\cite{mendez2019}, while Slater-type orbital expansions are 
 given in \cite{Hf_arxiv}.
 
-To assess the importance of the description of bound electrons, 
-Fig.~\ref{Binding_E} shows (a) non-relativistic~\cite{badnell97} and 
-our relativistic binding energies $E_{nl\pm}$, with $\pm=j\pm1/2$; and 
-(b) the relative errors with respect to the experimental data on solid 
-Hf~\cite{williams1995} from $1s$ to $4f_{\pm}$. 
-%\st{We notice that not only the most inner shells require relativistic calculations, but also the outer $5p$ and $4f$ shells.
-Furthermore, this figure shows very clearly the disability of non-relativistic calculations to describe the experimental data, which surprisingly worsens from the inner to the outer shells.}
-One would expect the relativistic contributions to only affect 
-significantly the inner shells. The relativistic binding energies of the 
-electrons belonging to these shells are 2.5\% from the experimental values,
-while the non-relativistic ones are within $\sim$10\%. Moreover, we notice 
-the relativistic corrections greatly influences the outer $5p$ and $4f$ 
-shells. This effect may be a consequence of the orthogonality with the 
-relativistic inner shells. Also, the sign of the binding energy deviations 
-is inverted near the FEG limit, i.e. the electrons go from being more 
-bounded --than the experimental values-- to being less bounded. 
-This change of sign can be attributed to the fact that the experimental 
-values correspond to hafnium in solid-state, while our theoretical 
-calculations correspond to the element in the gas phase.
+To} assess the importance of a full relativistic description of 
+bound electrons, Fig.~\ref{Binding_E} (a) shows our binding 
+energies, $E_{nl\pm}$, with $\pm=j\pm1/2$; the 
+non-relativistic values~\cite{badnell97}; and the experimental data on 
+solid Hf~\cite{williams1995}, which is available only for $1s$ to $4f_{\pm}$ 
+subshells, as expected.
+We notice that not only the most inner shells require relativistic calculations, but also the outer $5p$ and $4f$ shells.
+Furthermore, this figure shows very clearly the disability of non-relativistic calculations to describe the experimental data, which surprisingly worsens from the inner to the outer shells.
 
-Our relativistic binding energies present spin-orbit split. However, in 
-total stopping power, where the initial state of the excited electron is 
-not measured, the quantum uncertainty in energy $\Delta E$ melts this 
-split. The criteria $\Delta E\Delta t\geq\hbar/2$ merges the energies 
+In the comparison with the experimental values in Fig.~\ref{Binding_E} (a) 
+it can be noted that the sign of the binding energy deviations is inverted 
+for the outer $5s$-$4f\pm$ electrons, with the experimental binding energies 
+being less bounded than our theoretical ones. Small differences for the 
+outer shells are expected since the experimental values correspond to 
+hafnium in solid-state, while our theoretical calculations correspond to 
+the element in the gas phase.
+
+For more detail, Fig.~\ref{Binding_E} (b) shows the relative errors 
+with respect to the experimental  values. This figure shows clearly 
+that the relativistic corrections are very 
+important to describe the atomic structure of hafnium, even for the 
+outer shells. It turns out that the errors committed in the 
+non-relativistic calculations of the inner shell orbitals 
+propagate, through the Hartree-Fock approximation, to the outer shells.
+The importance of full-relativistic calculations for the outer 
+shells has already been noted for Au, Pb and Bi and W~\cite{mon09}.
+
+For the contribution of bound electrons to the total stopping 
+cross sections, the SLPA considers independent contributions of each 
+subshell. Our relativistic binding energies present spin-orbit split. 
+However, in total stopping power, where the initial state of the 
+excited electron is not measured, the quantum uncertainty in energy 
+$\Delta E$ melts this split. The criteria $\Delta E\Delta t\geq\hbar/2$ 
+merges the energies 
 $E_{nl+}-E_{nl-}$ for sufficiently small values of $\Delta t$ (the 
@@
 impact velocity, respectively. 
-
-%-----------------------------------------------------------------------
-\begin{figure*}[!t]
-\centering
-\includegraphics[width=13.cm]{Fig02.eps}
-\caption{Theoretical total stopping cross sections of protons in hafnium 
-adding FEG and bound $1s$-$4f$ contributions, as mentioned in the inset. 
-The vertical grey dashed-line indicates the energy above which plasmon 
-excitation is possible. The SLPA total stopping values for $1s$-$4f$ 
-bound electrons with and without $5p$-$4f$ screening are displayed with 
-solid and dotted lines, respectively.}
-\label{slpa4f}
-\end{figure*}
-%-----------------------------------------------------------------------
-
-The SLPA calculates the contribution of each subshell of bound electrons 
-to the total stopping cross sections. We found that for every sub-shell 
-of Hf, at the impact energy at which this sub-shell began to contribute, 
-the spin-orbit split was unresolved. Then, the $nl$-electrons should be 
+In the case of hafnium, we found that for every sub-shell of electrons, 
+the spin-orbit split is unresolved in the energy region this sub-shell 
+contributes. Therefore, the $nl$-electrons should be 
 considered together, responding to the ion passage as a single gas of 
@@
 together for impact energies $E>40$ keV (inter-shell screening).
+As already mentioned, at higher energies inter-shell screening is 
+possible for other subshells (i.e. 4p-4d for impact energies above 
+0.9 MeV), but its weight in the total stopping is minor for deeper shells.
 
@@
 \centering
-\includegraphics[width=13.0cm]{Fig03.eps}
-\caption{Stopping power cross section of hafnium for protons. Symbols: 
+\includegraphics[width=13.cm]{Fig02_new3.eps}
+\caption{(colour online) Theoretical stopping cross sections of protons in hafnium. 
+Blue dash-dotted-line,the non-perturbative SPCC for the FEG; blue solid-line, 
+ML results for the FEG (includes plasmon excitation); red
+solid and dotted-lines, the SLPA results for bound electrons with and 
+without $5p$-$4f$ screening, respectively. Black curves, total  stopping
+adding the FEG and bound $1s$-$4f$ contributions: dash-dotted-line, 
+SPCC (FEG) + SLPA (bound); solid-line, ML (FEG) + SLPA (bound) with 
+4f-5p screening; dotted-line, ML (FEG) + SLPA (bound) without 4f-5p screening.
+The vertical grey dashed-line indicates the energy of 37 keV above which plasmon 
+excitation is possible.} 
+\label{slpa4f}
+\end{figure*}
+%-----------------------------------------------------------------------
+
+Finally, in all our calculations~\cite{mon17} we assumed the projectile 
+to be proton and not neutral hydrogen. When an ion moves inside a metal,
+ the FEG screens the nucleus, so the binding energies will be smaller 
+than outside the metal, and this effect is more important a low impact 
+velocities $v$. In the case of hydrogen the difference is drastic, i.e. 
+for H inside Hf ($rs=2.07$), the $1s$-bound state is almost null at 
+$v<2$~\cite{suppression}. It is worth to mention that this assumption 
+agrees with Ziegler SRIM code~\cite{Ziegler01}, but differs from CasP 
+code~\cite{Grande} that predicts neutral hydrogen at very low velocities.
+
+%-----------------------------------------------------------------------
+\begin{figure*}[!t]
+\centering
+\includegraphics[width=13.0cm]{Fig03_new2.eps}
+\caption{(colour online) Stopping power cross section of hafnium for protons. Symbols: 
 solid circles, present values; open circles, previous data~\cite{Sirotinin}. 
 Curves: Black solid-line, present full theoretical results with 
-$4f$-$5p$ screening; violet dashed-line, values from ICRU-49~\cite{ICRU49}, 
-and green dotted line, SRIM-2013~\cite{Ziegler01}.}
+$4f$-$5p$ screening; pink dash-dot line, theoretical 
+CasP5.2~\cite{Grande,casp52} values; orange dash-double-dot line, 
+theoretical DPASS~\cite{DPASS20} results; green dotted-line, 
+semi-empirical SRIM-2013~\cite{Ziegler01}; violet dashed-line, 
+ICRU49~\cite{ICRU49} tabulated values.}
 \label{F03}
@@
 In Fig.~\ref{slpa4f}, we display the present theoretical stopping cross
-section of Hf for protons with and without the $5p$-$4f$ screening. We 
-added the FEG and the bound electron contributions as explained above. 
-The minimum energy for plasmon excitation was estimated to $37$ keV. We 
+section of Hf for protons using the relativistic wave functions and
+ binding energies, but with and without the $5p$-$4f$ screening. 
+We show separately the FEG and bound electron contributions and the 
+total stopping as the addition of both of them.
+The minimum energy for plasmon excitation was estimated $\sim 37$ keV. We
 used the non-perturbative SPCC model for impact energies $E \leq 37$~keV, 
-and the perturbative ML calculation above this energy. Clearly, 
+and the perturbative ML calculation above this energy. Bound $1s$-$4f$ 
+electrons (relativistic wave functions and binding energies) are 
+calculated with the SLPA and shown separately in Fig.~\ref{slpa4f} 
+with and without the $4f$-$5p$ screening. Below $\sim 40$ keV the 
+difference between both calculations is negligible. Clearly, 
 considering $5p$-$4f$ electrons as a single group of 20 electrons with 
@@
 the separate $5p$ and $4f$ contributions. Notice that this shell 
-correction can only be considered within a many electron model such as 
-the SLPA.
+correction can be considered naturally within a many electron model 
+such as the SLPA, .
 
@@
 
-The present data and theoretical results are displayed in 
-Table~\ref{table01}, and Fig.~\ref{F03}. As can be seen in 
-Table~\ref{table01}, an overall relative uncertainty of around 5\% was 
+The present data is displayed in 
+Table~\ref{table01}. %, and Fig.~\ref{F03}. 
+As can be observed, %seen in Table~\ref{table01},
+an overall relative uncertainty of around 5\% was 
 achieved for the experimental stopping power values, which are mainly 
@@
 \centering
-\caption{Stopping power values S$_{\mathrm{exp}}$ of hafnium for protons measured 
-in this work. $\Delta$E/E values are also shown.
-\label{table01}}
+\caption{Stopping power values S$_{\mathrm{exp}}$ of hafnium for protons 
+measured in this work. $\Delta$E/E values are also shown.}
+\label{table01}
 
@@
 
-In Fig.~\ref{F03}, we have good agreement between the present 
-theoretical results and the measurements displayed in Table~\ref{table01}. 
-Our theoretical approach also agrees with previous data by 
+Fig.~\ref{F03} synthesizes the results of the present work.
+The agreement between the present theoretical results and the new
+measurements displayed in Table~\ref{table01} is clear. Present 
+measurements using transmission method are in good agreement with the 
+previous data by Sirotinin~\cite{Sirotinin}, measured in backscattering 
+geometry. Our theoretical approach also agrees with the data by 
 Sirotinin~\cite{Sirotinin}, except for the lowest energy measurement at 
-80 keV. It is interesting that our full theoretical curve differs from 
-SRIM-2013 for impact energies below 100 keV. We obtain a stopping 
-maximum around $40 \times 10^{-15}$ eV cm$^2$/atom  at 65 keV. Instead, 
-following the up-to-now only set of data \cite{Sirotinin}, SRIM-2013 
-suggests a lower stopping maximum  at impact energy of 115 kev. It is 
-worth mentioning that similar theoretical results using the experimental 
-$r_S=2.07$ a.u. instead of the theoretical $r_S=2.14$ a.u. give the 
-stopping maximum at the same impact energy but $4\%$ higher. Future 
-experiments for impact energies below 100 keV would be important to 
-complete this study.
+80 keV. We have also included in this figure the theoretical curves from 
+Grande and Schiwietz CasP5.2 code~\cite{Grande,casp52} and from Sigmund 
+and Schinner DPASS code~\cite{DPASS20}, both available online; also the 
+semi-empirical results from SRIM-2013~\cite{Ziegler01} and the ICRU49 
+tables~\cite{ICRU49}. It is interesting that our full theoretical curve 
+differs from SRIM-2013 for impact energies below 100 keV. We obtain a 
+stopping maximum around $40 \times 10^{-15}$ eV cm$^2$/atom  at 65 keV. 
+Instead, following the up-to-now only set of data \cite{Sirotinin}, SRIM-2013
+suggests a lower stopping maximum  at impact energy of 115 kev. 
+The stopping maximum is a very sensitive region for any full theoretical 
+description, with no parameters at all, and this is quite visible in a 
+linear-scale plot like Fig.~\ref{F03}. However, the impact energy for 
+the maximum seem to agree between our curve and DPASS, although is 
+$10 \%$ below. Instead, CasP predicts a stopping maximum $10 \%$ above 
+our value and a completely different shape at lower energies. It is 
+worth mentioning that our model gives similar results using the experimental 
+$r_S=2.07$ a.u. instead of the theoretical $r_S=2.14$ a.u. with the 
+stopping maximum at the same impact energy but $4\%$ higher.
+
+Future experiments would be important to complete this study on the 
+stopping of Hf for protons. In the energy region $80-500$ keV only 
+the measurements by Sirotinin group~\cite{Sirotinin} in 1984 are 
+available. SRIM-2013 clearly includes these values in the adjust of 
+the code, so it does not represent a test for this data. Undoubtedly, 
+new measurements are needed in three specific energy regions: 
+below 500 keV, around the stopping maximum (i.e. $50-200$ keV) and at 
+low impact energies.
 
@@
 energies the difference between our full-theoretical results and SRIM 
-is substantial. To the best of our knowledge, these are the first 
+is substantial. We compare our theoretical results with other two 
+models given by the DPASS and CasP5.2 codes. Differences can be noted 
+at intermediate to low impact energies, but  they also support an 
+stopping maximum at lower energy than SRIM predictions.
+
+To the best of our knowledge, these are the first 
 theoretical calculations of stopping in Hf taking into account the 
-atomic relativistic effect. Future experiments for impact energies 
-below 100 keV would be important to complete this study.
+atomic relativistic effect and the screening among electrons in a 
+consistent way from very low to high impact energies. Future 
+experiments for impact energies around the stopping maximum and in 
+the low energy region would be important to have a more complete 
+knowledge of the stopping of protons in hafnium.
 
@@
 
-\bibitem{ICRU49} 
-ICRU report 49, 
-Stopping Powers and Ranges for Protons and Alpha Particles, 
-International Commission on Radiation Units and Measurements, 1993.
+\bibitem{Grande} 
+G. Schiwietz, P. L. Grande, 
+Nucl. Instrum. Meth. B \textbf{175-177}, 125 (2001); 
+CasP code, available from www.casp-program.org
+
+\bibitem{casp52} 
+G. Schiwietz and P. L. Grande,
+Nucl. Instr. and Meth. B \textbf{273}, 1-5 (2012); 
+P.L. Grande and G. Schiwietz,
+Phys. Rev. A \textbf{58}, 3796 (1998).
+
+\bibitem{DPASS20}
+A. Schinner, P. Sigmund, 
+Nucl. Instrum. Meth. B \textbf{460}, 19 (2019); 
+P.Sigmund, A. Schinner, 
+Eur. Phys. J. D \textbf{12}, 425 (2000). 
+DPASS code, available from https://www.sdu.dk/en/DPASS/
 
 \bibitem{Ziegler01} 
-J.F. Ziegler, J.P. Biersack, SRIM2013, 
-Computer Program and Manual. Available from www.srim.org.
+J.F. Ziegler, J.P. Biersack, M. D. Ziegler, 
+\textit{SRIM, The Stopping and Range of Ions in Matter}, 
+(SRIM Co. Maryland, USA, 2008); 
+SRIM2013, Computer Program and Manual. Available from www.srim.org.
+
+\bibitem{ICRU49} 
+ICRU report 49, \textit{Stopping Powers and Ranges for Protons and Alpha Particles},
+International Commission on Radiation Units and Measurements (1993).
 
@@
 G. Williams in http://xdb.lbl.gov/Section1/Sec\_1-1.html
+
+\bibitem{mon09} 
+C.C. Montanari, D.M. Mitnik, C.D. Archubi, and J.E. Miraglia, 
+Phys. Rev. A \textbf{70}, 032903 (2009); 
+and Phys. Rev. A \textbf{80}, 012901 (2009).
 

### 2020-04-20 feb_submission/Paper_SPHfP_feb2020_prb.tex
checked with grammarly

--- 2020-03-11 Paper_SPHfP_feb2020_prb.tex
@@
 The measurements were performed at the Laboratory of Accelerators and X-Ray Diffraction in Lisbon, by using the transmission method on self-supporting stopping material.
-%for (0.6-2.5) MeV protons.
 The overall uncertainty of around $5\%$ was established over the protons energy range  (0.6-2.5) MeV.
-The theoretical developments involved fully relativistic atomic structure calculations for %this Z=72 target
- Hf, which required the solution of the Dirac equation.
+The theoretical developments involved fully relativistic atomic structure calculations for Hf, which required the solution of the Dirac equation.
 The shell-wise local plasma approximation (SLPA) was used to describe the energy transferred to the bound $1s$-$4f$ electrons, and the outer four electrons were considered as a free electron gas (FEG).
-%We contemplated two different models for the FEG: the screened potential with cusp condition (SPCC) for energies below that of the plasmon excitation, and the dielectric formalism, for energies around the stopping maximum and above.
- We found the relativistic description of the $4f$-shell and the screening between $4f$ and $5p$ electrons are decisive around the stopping maximum. %for energies above 40 keV.
- Present theoretical and experimental results are in very good agreement in the energy region of the new measurements. However, the present theoretical stopping cross sections show %interesting 
- substantial differences with the most used semi-empirical models (SRIM2013 and ICRU-49) at intermediate to low energies.
-Our calculations suggest the stopping maximum %around 40 eV$/(10^{15}$ atom/cm$^2)$ at 70 keV, which is
- higher and shifted to lower energies than these previous predictions. Future measurements below 100 keV would be decisive for a better knowledge  around the maximum and below.
+We found the relativistic description of the $4f$-shell and the screening between $4f$ and $5p$ electrons are decisive around the stopping maximum.
+Present theoretical and experimental results are in very good agreement in the energy region of the new measurements.
+However, the present theoretical stopping cross sections show substantial differences with the most used semi-empirical models (SRIM2013 and ICRU-49) at intermediate to low energies.
+Our calculations suggest the stopping maximum higher and shifted to lower energies than these previous predictions.
+Future measurements below 100 keV would be decisive for a better knowledge  around the maximum and below.
 \end{abstract}
@@
 
-For impact energies above a few keV/amu, mono-energetic charged particles penetrating a foil of any material lose their energy through a series of consecutive inelastic collisions, mainly with target electrons \cite{Chu01,Sigmund}.
-The information given by the energy loss process is essential not only to have a better knowledge of the physics behind the fundamental interactions, but also because it plays an important role in many applied fields such as materials science, nuclear physics, ionic implantation and radiotherapy \cite{Sigmund,Schardt}.
-Experimental data on ion mean energy loss per unit path $S(E)$ is of crucial relevance to check the reliability of semi-empirical models and to determine some key parameters
-%included in theoretical approaches of stopping power %\cite{Diwan,Damache01,Damache02,Damache03,Damache04}. 
-\cite{Diwan,Damache04,Damache02}. The experimental data available is often rather scarce, which is %particularly 
-troublesome when the material under study corresponds to an element of low occurrence on the Earth's upper crust, such as hafnium.
+For impact energies above a few keV/amu, mono-energetic charged particles penetrating a foil of any material lose their energy through a series of consecutive inelastic collisions, mainly with target electrons~\cite{Chu01,Sigmund}.
+The information given by the energy loss process is essential not only to have a better knowledge of the physics behind the fundamental interactions, but also because it plays an important role in many applied fields such as materials science, nuclear physics, ionic implantation and radiotherapy~\cite{Sigmund,Schardt}.
+Experimental data on ion mean energy loss per unit path $S(E)$ is of crucial relevance to check the reliability of semi-empirical models and to determine some key parameters~\cite{Diwan,Damache04,Damache02}.
+The experimental data available is often rather scarce, which is troublesome when the material under study corresponds to an element of low occurrence on the Earth's upper crust, such as hafnium.
 
-So far, only one experimental work has been published regarding the stopping power cross section of pure hafnium for protons \cite{Sirotinin}, while more attention has been recently given to studies involving hafnium oxide due to its practical use \cite{Abril,Behar,Primetzhofer}.
-It is well known that significant attention has been paid in recent years to transition metal-oxides such as HfO$_2$ because of their potential as alternative gate dielectrics to replace SiO$_2$ for the future generation of nano-electronics with less than 45 nm gate length \cite{Choi,Robertson}.
-Some important physical properties of the above mentioned metal-oxide films depend on its thickness, which is often measured by using Rutherford Backscattering Spectrometry \cite{Alfassi01, Tesmer01}, a method that relies heavily on the determination of the stopping power of ion beams in the material of interest.
+So far, only one experimental work has been published regarding the stopping power cross section of pure hafnium for protons~\cite{Sirotinin}, while more attention has been recently given to studies involving hafnium oxide due to its practical use~\cite{Abril,Behar,Primetzhofer}.
+It is well known that significant attention has been paid in recent years to transition metal-oxides such as HfO$_2$ because of their potential as alternative gate dielectrics to replace SiO$_2$ for the future generation of nano-electronics with less than 45 nm gate length~\cite{Choi,Robertson}.
+Some important physical properties of the above mentioned metal-oxide films depend on its thickness, which is often measured by using Rutherford Backscattering Spectrometry~\cite{Alfassi01,Tesmer01}, a method that relies heavily on the determination of the stopping power of ion beams in the material of interest.
 
 In this study, we report experimental stopping power cross sections over the incident energy range (0.6-2.5) MeV for protons crossing self-supported Hf thin-film by using the transmission method.
-We aim not only to upgrade stopping power data compilations \cite{HPaul03,mondim17} but also to provide useful information about the processes governing the slowing down of protons in multi-electronic targets.
+We aim not only to upgrade stopping power data compilations~\cite{HPaul03,mondim17} but also to provide useful information about the processes governing the slowing down of protons in multi-electronic targets.
 In the rare earth metals, the $4f$ electrons play a relevant role in the stopping power, being the first shell of bound electrons below the conduction band.
-As already noted \cite{Roth17}, the free electron gas (FEG) shows unexpected %a peculiar
+As already noted~\cite{Roth17}, the free electron gas (FEG) shows unexpected %a peculiar
 behavior in these elements, which casts %that states 
@@
 
-The theoretical approach implemented in this work uses the shell-wise local plasma approximation (SLPA) \cite{mon13} to describe the energy transferred to the bound $1s$-$4f$ electrons, and two different models for the FEG: the screened potential with cusp condition model (SPCC) \cite{mon17}, which is non-linear binary formalism%,  for energies below that of the plasmon excitation
-; and the Mermin-Lindhard dielectric formalism (ML) \cite{Mermin}, for energies around the stopping maximum and above. Our model requires 
-the relativistic wave functions and binding energies of Hf, and considers 4 electrons per atom in the  FEG \cite{mendez2019}. 
+The theoretical approach implemented in this work uses the shell-wise local plasma approximation (SLPA)~\cite{mon13} to describe the energy transferred to the bound $1s$-$4f$ electrons, and two different models for the FEG: the screened potential with cusp condition model (SPCC)~\cite{mon17}, which is non-linear binary formalism%,  for energies below that of the plasmon excitation
+; and the Mermin-Lindhard dielectric formalism (ML)~\cite{Mermin}, for energies around the stopping maximum and above. Our model requires 
+the relativistic wave functions and binding energies of Hf, and considers 4 electrons per atom in the FEG~\cite{mendez2019}. 
 %On the other hand,
@@
 
-The experimental details and data are given in Section \ref{experiment}, while the theoretical method is explained in Section \ref{theory}.
-Present theoretical and experimental values are finally compared to the available experimental values \cite{Sirotinin},  the ICRU-49 calculations \cite{ICRU49} and the SRIM-2013 package \cite{Ziegler01}.
-Conclusions and discussions are given in Section \ref{conclusion}.
-All the present data can be found at the Zenodo platform \cite{zenodo}.
+The experimental details and data are given in Section~\ref{experiment}, while the theoretical method is explained in Section~\ref{theory}.
+Present theoretical and experimental values are finally compared to the available experimental values~\cite{Sirotinin}, the ICRU-49 calculations~\cite{ICRU49} and the SRIM-2013 package~\cite{Ziegler01}.
+Conclusions and discussions are given in Section~\ref{conclusion}.
+All the present data can be found at the Zenodo platform~\cite{zenodo}.
 
@@
 \includegraphics[width=9cm]{Fig01.eps}
-\caption{RBS spectrum for $E_{\mathrm{avg}} = 921.1$ keV protons on hafnium sample which is subsequently used to determine the energy loss in the foil.}
+\caption{RBS spectrum for $E_{\mathrm{avg}}=921.1$ keV protons on hafnium sample which is subsequently used to determine the energy loss in the foil.}
 \label{F01}
@@
 \subsection{Accelerator and scattering Chamber}
-The procedure used in this work to obtain stopping power data is essentially the same as described in Ref. \cite{Miranda01}. The present measurements were made at the IST/LATR
+The procedure used in this work to obtain stopping power data is essentially the same as described in Ref.~\cite{Miranda01}. The present measurements were made at the IST/LATR
 (Laboratory of Accelerators and X-Ray Diffraction) in Lisbon.
@@
 \subsection{Target}
-The stopping material under analysis was a hafnium foil with nominal thickness of 1.0 $\mu$m and 99.95\% purity which was supplied by Lebow Company \cite{Lebow}.
+The stopping material under analysis was a hafnium foil with nominal thickness of 1.0 $\mu$m and 99.95\% purity which was supplied by Lebow Company~\cite{Lebow}.
 However, a more precise thickness value was achieved by measuring the energy loss of alpha particles coming from a calibrated ($^{239}$Pu, $^{241}$Am, $^{244}$Cm) source.
-From the alpha spectra with and without the Hf foil interposed, the characteristic energy shift $\delta$E was measured and then combined with the stopping power for 5.486 MeV alphas on hafnium (55.69 eV/$10^{15}$ at/cm$^2$) found in Ref.\cite{Ziegler01} to obtain an areal density of ($4.13 \pm 0.21$)$\times 10^{19}$ at/cm$^2$ which corresponds to a thickness of $0.920\pm0.046 \mu$m.
+From the alpha spectra with and without the Hf foil interposed, the characteristic energy shift $\delta$E was measured and then combined with the stopping power for 5.486 MeV alphas on hafnium (55.69 eV/$10^{15}$ at/cm$^2$) found in Ref.~\cite{Ziegler01} to obtain an areal density of ($4.13 \pm 0.21$)$\times 10^{19}$ at/cm$^2$ which corresponds to a thickness of $0.920\pm0.046 \mu$m.
 %---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
@@
 Once the beam impinges on the Au/SiO$_2$ sample, protons are back-scattered towards a Si surface barrier detector located at 140$^{\circ}$ relative to the initial beam direction.
-Fig.\ref{F01} shows two particle spectra where ion energies $E_2$ and $E_1$ are associated to a placed and removed hafnium sample, respectively.
-Both energy distributions were fitted by Gaussian functions to obtain the mean energy and width (FWHM) of the peaks \cite{Sun01}, and from the difference between these two peak positions in the spectrum, the total energy loss $\Delta E = (E_1 - E_2)$ in the foil was calculated.
-As established in previous studies \cite{Miranda01, Damache02}, the experimental stopping power cross sections $\varepsilon (E) $ are determined at some mean energy $E_{\mathrm{avg}}$ by measuring the ion energy losses $\Delta E$ within the investigated Hf foil, which has a mean thickness denoted by $\Delta x$.
+Fig.~\ref{F01} shows two particle spectra where ion energies $E_2$ and $E_1$ are associated to a placed and removed hafnium sample, respectively.
+Both energy distributions were fitted by Gaussian functions to obtain the mean energy and width (FWHM) of the peaks~\cite{Sun01}, and from the difference between these two peak positions in the spectrum, the total energy loss $\Delta E = (E_1 - E_2)$ in the foil was calculated.
+As established in previous studies~\cite{Miranda01,Damache02}, the experimental stopping power cross sections $\varepsilon (E) $ are determined at some mean energy $E_{\mathrm{avg}}$ by measuring the ion energy losses $\Delta E$ within the investigated Hf foil, which has a mean thickness denoted by $\Delta x$.
 In this way, only when the energy loss fraction $\Delta E/E_{\mathrm{avg}}$ across the Hf foil is not exceeding 20\%
-it is possible to define the stopping cross section by \cite{Raisanen01,Schulz01}:
+it is possible to define the stopping cross section by~\cite{Raisanen01,Schulz01}:
 \begin{equation}\label{eq:stcross}
- \varepsilon (E) = \frac{S(E)}{N} = -\frac{dE}{N dx} \approx -\frac{\Delta E}{N \Delta x},
+ \varepsilon(E)=\frac{S(E)}{N}=-\frac{dE}{N\,dx}\approx-\frac{\Delta E}{N\Delta x},
 \end{equation}
 where $N$ denotes the atomic number density (atoms cm$^{-3}$) of the material under study.
-When this condition was not fulfilled, a small correction to the mean energies $E_{\mathrm{avg}}$ was applied in order to account for the nonlinear dependence on ion energy of stopping powers \cite{Chilton,Rajatora}.
+When this condition was not fulfilled, a small correction to the mean energies $E_{\mathrm{avg}}$ was applied in order to account for the nonlinear dependence on ion energy of stopping powers~\cite{Chilton,Rajatora}.
 
@@
 Above certain velocity (i.e.
-$v\geq 1.5 \ v_F$) , not only binary but also collective excitations (plasmons) are possible \cite{mon17}.
+$v\geq 1.5\,v_F$) , not only binary but also collective excitations (plasmons) are possible~\cite{mon17}.
 Moreover, at high energies, also the bound electrons contribute to the stopping power.
@@
 
-We used the SPCC model \cite{mon17} to describe the stopping power of low velocity charged particles in the FEG.
+We used the SPCC model~\cite{mon17} to describe the stopping power of low velocity charged particles in the FEG.
 It is a non-perturbative binary collisional approximation, thus valid at energies below that of plasmon excitations.
-The SPCC \cite{mon17} is based on a screened central potential with cusp condition of the electronic density close to the projectile.
-This model proved to give a good description of the induced electron density even for negative projectiles \cite{mon17}, and reproduces the low velocity proton-antiproton differences in the stopping power (Barkas effect).
+The SPCC~\cite{mon17} is based on a screened central potential with cusp condition of the electronic density close to the projectile.
+This model proved to give a good description of the induced electron density even for negative projectiles~\cite{mon17}, and reproduces the low velocity proton-antiproton differences in the stopping power (Barkas effect).
 The SPCC formalism only depends on the Wigner-Seitz radio, $r_S$, which is a measure of the electronic density of the FEG.
-For metals of well-known $r_S$, the SPCC describes correctly the low energy experimental stopping data \cite{mon17}, agreeing with the DFT results by Echenique and coworkers \cite{eche81,nagy89} at $v=0$.
+For metals of well-known $r_S$, the SPCC describes correctly the low energy experimental stopping data~\cite{mon17}, agreeing with the DFT results by Echenique and coworkers~\cite{eche81,nagy89} at $v=0$.
 
 %In the case of 
-Hafnium ($Z=72$, [Xe] $4f^{14} \ 6s^2 \ 5d_{3/2}^1 \ 5d_{5/2}^1$), belongs to the first groups of transition metals, with  4 electrons as FEG (theoretical $r_S=2.14$ a.u.) and $1s$-$4f$ electrons bound.
-We compared the computed $r_S$ with the experimental value obtained from measured energy loss function by Lynch \textit{et al} \cite{lynch75}.
-The experimental plasmon energy of Hf is $\hbar\omega_P \approx 15.8$ eV, with a width at half maximum $\delta \approx 4.4$ eV, and $r_S \approx 2.07$ a.u.
-\cite{lynch75}.
-The difference of less than $5\%$ between theoretical and experimental $r_S$ assess Hf as a canonical target \cite{mon17}.
+Hafnium ($Z=72$, [Xe] $4f^{14}\,6s^2\,5d_{3/2}^1\,5d_{5/2}^1$), belongs to the first groups of transition metals, with  4 electrons as FEG (theoretical $r_S=2.14$ a.u.) and $1s$-$4f$ electrons bound.
+We compared the computed $r_S$ with the experimental value obtained from measured energy loss function by Lynch \textit{et al}~\cite{lynch75}.
+The experimental plasmon energy of Hf is $\hbar\omega_P \approx 15.8$ eV, with a width at half maximum $\delta \approx 4.4$ eV, and $r_S \approx 2.07$ a.u.~\cite{lynch75}.
+The difference of less than $5\%$ between theoretical and experimental $r_S$ assess Hf as a canonical target~\cite{mon17}.
 
-Above certain impact velocity, the plasmon contribution is important (i.e. around and above the stopping maximum). An interesting value for our analysis is the minimum impact velocity to excite
-plasmons, $v_P$.
-In the dielectric formalism, this value can be obtained as $v_P\ \approx \ v_F[1+(3 \pi\ v_F)^{-1/2} ]$ \cite{suppression}.
-To describe the energy loss considering collective and binary excitation, we resort to the ML dielectric formalism \cite{Mermin}, which is a linear response, perturbative approximation, so it depends on the square of the ion charge.
+Above certain impact velocity, the plasmon contribution is important (i.e.
+around and above the stopping maximum).
+An interesting value for our analysis is the minimum impact velocity to excite plasmons, $v_P$.
+In the dielectric formalism, this value can be obtained as $v_P\,\approx\,v_F[1+(3\pi\,v_F)^{-1/2}]$~\cite{suppression}.
+To describe the energy loss considering collective and binary excitation, we resort to the ML dielectric formalism~\cite{Mermin}, which is a linear response, perturbative approximation, so it depends on the square of the ion charge.
 In this formalism, the response of target electrons to the ion passage is described through the quantum dielectric function, depending on the characteristic $r_S$ and $\delta$ parameters of the FEG.
@@
 
-For the stopping power due to bound electrons, the SLPA \cite{mon17,mon13} is employed.
+For the stopping power due to bound electrons, the SLPA~\cite{mon17,mon13} is employed.
 It is worth to mention that the only inputs for the SLPA are the space-dependent densities of each shell in the ground state, and their binding energies.
@@
 Hafnium is a relativistic target, therefore the wave functions and binding energies must be obtained by solving the many–electron Dirac Hamiltonian.
-Details of these calculations have been published in \cite{mendez2019}, and the Slater-type orbital expansions in \cite{Hf_arxiv}.
+Details of these calculations have been published in~\cite{mendez2019}, and the Slater-type orbital expansions in~\cite{Hf_arxiv}.
 
-Too assess the importance of the description of bound electrons, Fig.
-\ref{Binding_E} shows (a) non-relativistic \cite{badnell97} and our relativistic binding energies, $E_{nl\pm}$, with $\pm=j\pm1/2$; and (b) the relative errors with respect to the experimental data on solid Hf \cite{williams1995} from $1s$ to $4f_{\pm}$.
+Too assess the importance of the description of bound electrons, Fig.~\ref{Binding_E} shows (a) non-relativistic~\cite{badnell97} and our relativistic binding energies, $E_{nl\pm}$, with $\pm=j\pm1/2$; and (b) the relative errors with respect to the experimental data on solid Hf~\cite{williams1995} from $1s$ to $4f_{\pm}$.
 We notice that not only the most inner shells require relativistic calculations, but also the outer $5p$ and $4f$ shells.
-%Figure \ref{Binding_E} (b) displays the relative error of non relativistic and relativistic energies with respect to the experimental data in logarithmic scale. 
 Furthermore, this figure shows very clearly the disability of non-relativistic calculations to describe the experimental data, which surprisingly worsens from the inner to the outer shells.
-%It is clear in this figure the disability of non-relativistic binding energies to describe the experimental data, which surprisingly worsens from the inner to the outer-shells.
-%We used here the theoretical binding energies and wave functions of Hf calculated in \cite{mendez2019}. 
-%The present theoretical description of Hf considers a FEG of $r_S=2.14$ and 1s-4f electrons as bound.
 
-%The relativistic results for binding energies and wave functions in \cite{mendez2019} 
 Our relativistic binding energies present spin-orbit split.
 However, in total stopping power where the initial state of the excited electron is not measured, the quantum uncertainty in energy $\Delta E$ melts this split.
-The criteria $\Delta E\Delta t \geq \hbar /2$ merges the energies  $E_{nl+} - E_{nl-}$ for sufficiently small values of $\Delta t$ (the collisional mean-time).
-In fact, at sufficiently high impact velocity, we can expect all target electrons to respond together to the ion passage \cite{lindhard53,chu72}.
-Following previous work \cite{mon09}, the collisional time is estimated as $\Delta t \approx \langle r_i\rangle/v$, with $\langle r_i\rangle$ and $v$ being the orbital mean radio and impact velocity, respectively.
+The criteria $\Delta E\Delta t\geq\hbar/2$ merges the energies $E_{nl+}-E_{nl-}$ for sufficiently small values of $\Delta t$ (the collisional mean-time).
+In fact, at sufficiently high impact velocity, we can expect all target electrons to respond together to the ion passage~\cite{lindhard53,chu72}.
+Following previous work~\cite{mon09}, the collisional time is estimated as $\Delta t\approx\langle r_i\rangle/v$, with $\langle r_i\rangle$ and $v$ being the orbital mean radio and impact velocity, respectively.
 
@@
 %The FEG calculations employ the SPCC model at low energies (no plasmon excitation) and the ML formalism (includes plasmon excitation).
-The SLPA total stopping values for $1s$-$4f$ bound electrons with and without $5p-4f$ screening are displayed with solid and dotted lines, respectively.}
+The SLPA total stopping values for $1s$-$4f$ bound electrons with and without $5p$-$4f$ screening are displayed with solid and dotted lines, respectively.}
 \label{slpa4f}
@@
 For example, the $4f_{-}$ and $4f_{+}$ of Hf can only be resolved for impact energies $E<0.05$ keV, but the contribution of $4f$ to the total stopping is negligible for $E<40$ keV.
-Moreover, the $5p$ and $4f$ electrons of Hf are very close in energy ( $\Delta E_{5p-4f} \approx 1$ a.u.
-\cite{mendez2019}) %merges them
-and they react together for impact energies $E>40$ keV (inter-shell screening).
+Moreover, the $5p$ and $4f$ electrons of Hf are very close in energy ($\Delta E_{5p-4f} \approx 1$ a.u.~\cite{mendez2019}) and they react together for impact energies $E>40$ keV (inter-shell screening).
 
@@
 \caption{Stopping power cross section of hafnium for protons.
-Symbols: solid circles, present values; open circles, previous data \cite{Sirotinin}.
-Curves: Black solid-line, present full theoretical results with $4f$-$5p$ screening; violet dashed-line, values from ICRU-49 \cite{ICRU49}, and green dotted line, SRIM-2013 \cite{Ziegler01}.}
+Symbols: solid circles, present values; open circles, previous data~\cite{Sirotinin}.
+Curves: Black solid-line, present full theoretical results with $4f$-$5p$ screening; violet dashed-line, values from ICRU-49~\cite{ICRU49}, and green dotted line, SRIM-2013~\cite{Ziegler01}.}
 \label{F03}
@@
 
-In Fig.
-\ref{slpa4f}, we display the present theoretical stopping cross section of Hf for protons with and without the $5p$-$4f$ screening.
+In Fig.~\ref{slpa4f}, we display the present theoretical stopping cross section of Hf for protons with and without the $5p$-$4f$ screening.
 We added the FEG and the bound electron contributions as explained above.
@@
 
-The present data and theoretical results are displayed in Table \ref{table01}, and Fig.
-\ref{F03}.
-As can be seen in Table~\ref{table01}, an overall relative uncertainty of around 5$\%$ was achieved for the experimental stopping power values, which are mainly due to the uncertainty in the hafnium foil thickness.
-%A good agreement in the whole energy range is found between our new data and the previous measurements made by Sirotinin \textit{et al.} \cite{Sirotinin}, as depicted in Fig.~\ref{F03}.
-Also shown in the same figure are our theoretical stopping cross sections together with the semi-empirical SRIM-2013 curve \cite{Ziegler01} and the suggested values by the ICRU-49 report \cite{ICRU49}.
+The present data and theoretical results are displayed in Table~\ref{table01}, and Fig.~\ref{F03}.
+As can be seen in Table~\ref{table01}, an overall relative uncertainty of around 5\% was achieved for the experimental stopping power values, which are mainly due to the uncertainty in the hafnium foil thickness.
 
@@
 
-In Fig.
-\ref{F03}, we have good agreement between the present theoretical results and the measurements displayed in Table~\ref{table01}.
-Our theoretical approach also agrees with previous data by Sirotinin \cite{Sirotinin}, except for the lowest energy measurement at 80 keV.
+In Fig.~\ref{F03}, we have good agreement between the present theoretical results and the measurements displayed in Table~\ref{table01}.
+Our theoretical approach also agrees with previous data by Sirotinin~\cite{Sirotinin}, except for the lowest energy measurement at 80 keV.
 It is interesting that our full theoretical curve differs from SRIM-2013 for impact energies below 100 keV.
 We obtain a stopping maximum around $40 \times 10^{-15}$ eV cm$^2$/atom  at 65 keV.
-Instead, following the up-to-now only set of data \cite{Sirotinin}, SRIM-2013 suggests a lower stopping maximum  at impact energy of 115 kev.
+Instead, following the up-to-now only set of data~\cite{Sirotinin}, SRIM-2013 suggests a lower stopping maximum  at impact energy of 115 kev.
 It is worth mentioning that similar theoretical results using the experimental $r_S=2.07$ a.u.

### 2020-05-11 response.tex
checked english with grammarly

--- 2020-04-11 response.tex
@@
 \usepackage{xcolor}
+\usepackage[margin=0.8in]{geometry}
+\usepackage{graphicx}
 
 \setstcolor{red}
+\def\reviewer#1{\vspace{0.35cm}\textsl{#1}}
+\def\reply#1{\vspace{0.1cm}\textcolor{red}{#1}}
 
@@
 % ----------------------------------------------------------------------
-
-\textsl{1. Since only protons are considered as projectiles, the title 
+\reviewer{1. Since only protons are considered as projectiles, the title 
 needs to be modified accordingly.}
 
-\vspace{0.1cm}
-The title has been modified following the referee's recommendation to
-``Stopping power of hafnium by proton impact and the importance of 
-relativistic $4f$ electrons''.
+% {\color{red} The title has been modified following the referee's 
+% recommendation:}{\small ''The stopping power of hydrgen in hafnium 
+% and the importance of relativistic $4f$ electrons''}.
+\reply{The title has been modified following the referee's
+recommendation to ``The stopping power of hydrogen in hafnium and the
+importance of relativistic $4f$ electrons''}.
 
 % ----------------------------------------------------------------------
-\vspace{0.25cm}
-\textsl{2. It appears that calculations assume bare protons. According 
+\reviewer{2. It appears that calculations assume bare protons. According 
 to CasP, neutrals play a considerable role in charge equilibrium around 
@@
 
-\vspace{0.1cm}
-{\color{red}Reply here.}
+%\reply{It is correct, our calculations assume bare protons.
+We consider that there is no bound electron or very small probability because free electrons of metals screens the hydrogen and this screening is enough at low energies as to have no bound electron.
+CasP includes a very good fitting of experimental charge states and a powerful scaling rule about them.
+However, it is based on measurements of charge states outside the metal.
+For multi-charged ions almost no differences are expected and CasP charge state works very well.
+For H and He ions, differences are important.
+We add the following explanation about why considering just proton impact, and the reference to CasP and the paper of charge states by Schwietz and Grande.} \\ {\small ''Finally, in all our calculations we assumed the projectile to be proton and not neutral hydrogen.
+When an ion moves inside a metal, the FEG screens the nucleus, so the binding energies will be smaller than outside the metal, and this effect is more important a low impact velocities $v$.
+In the case of hydrogen the difference is drastic, i.e.
+for H inside Hf ($rs=2.07$), the $1s$-bound state is almost null at $v<2$.
+It is worth to mention that this assumption agrees with Ziegler SRIM code, but differs from CasP code that predicts neutral hydrogen at very low velocities.''}
+\reply{The referee is correct; our calculations assume bare protons. We
+considered that there is no bound electron (or insignificant probability
+of there being any bound electron) because the free electrons in the
+metal screen the hydrogen, and, at low energies, this screening is
+sufficient as to have no bound electron. CasP includes an excellent
+fitting of experimental charge states and a robust scaling rule about
+them. However, it is based on measurements of charge states outside the
+metal. For multi-charged ions, almost no differences are expected, and
+the CasP charge state works very well. For H and He ions, differences
+are significant.  We incorporated a new paragraph (lines 263-273) to
+explain why we are considering just proton impact, the reference to
+CasP, and the paper of charge states by Schwietz and Grande.}
 
 % ----------------------------------------------------------------------
-\vspace{0.25cm}
-\textsl{3. Thickness measurements by Rutherford backscattering rely 
+\reviewer{3. Thickness measurements by Rutherford backscattering rely 
 ‘heavily’ on the scattering cross section, while energy loss is a minor 
@@
 
-\vspace{0.1cm}
-{\color{red}Reply here.}
+\reply{Pedro Miranda...................................................}
 
 % ----------------------------------------------------------------------
-\vspace{0.25cm}
-\textsl{4. In order to get a more precise thickness value of the Hf 
+\reviewer{4. In order to get a more precise thickness value of the Hf 
 foil, the authors use the energy loss of 5.486 MeV alphas as a standard, 
@@
 
-\vspace{0.1cm}
-{\color{red}Reply here.}
+%\reply{The referee is correct, we have used the stopping of 5.486 MeV alphas and SRIM as reference to get the thickness.
+Although SRIM fits the experimental data available, it is not just a fitting, it is a mix of model and fitting (semiempirical).
+At high impact energies where the stopping is well known srim is quite reliable.
+Figure \ref{HeHf} here can clarify these assertions.
+a) In this figure we plotted together H and He in Hf, dividing by $Z^2$.
+Above 1 MeV/amu both srim curves and our theoretical results agree within $2\%$.
+The vertical pink line indicates the impact energy we use as reference.
+It is in the region where the srim predictions are reliable.
+b) srim is semiempirical, the comparison of H and He in the same target shows clearly how srim for H ions curves down and up to fit Sirotonin data.
+It is also clear how srim for He ion overvalue in order to fit the single measurement by Chu, Ziegler and others.
+Why do we say "overvalue"?
+Because the stopping of He ions divided by 4 can be equal or below the stopping of H ions.
+}
+\reply{The referee is correct; we have used the stopping of 5.486 MeV
+alphas and the SRIM software package as a reference to get the 
+thickness. Although SRIM fits the experimental data available, it is not
+just a fitting; the software implements a mix of modeling and fitting
+(semiempirical). At high impact energies, where the stopping is well
+known, SRIM is quite reliable. Figure \ref{HeHf} included in this
+reply can clarify these assertions. a) In this figure, we plotted H and 
+He in Hf, both divided by $Z^2$. Above 1 MeV/amu, SRIM and our
+theoretical results agree within $2\%$. The vertical pink line indicates
+the impact energy used as reference. In the region, the SRIM predictions
+are reliable. b) By comparing the SRIM results for hydrogen and helium
+in the same target, one can see that SRIM for H ions rapidly changes its
+concavity to fit Sirotonin data. Furthermore, SRIM for He ion 
+overestimates the scaled-with-$Z^2$ cross-section at 500 keV/amu in
+order to fit the single measurement by Chu, Ziegler, and collaborators. 
+We say the stopping is overestimated because the stopping of He ions 
+divided $Z^2$ can only be equal or below the stopping of H ions.}
 
-% ----------------------------------------------------------------------
-\vspace{0.25cm}
-\textsl{5. In connection with the discussion of relativity and 
+%-----------------------------------------------------------------------
+\begin{figure}[!t]
+\centering
+\includegraphics[width=12.cm]{Fig_HeHf.eps}
+\caption{Stopping cross section of H and He in hafnium at high impact energies.}
+\label{HeHf}
+\end{figure}
+
+%-----------------------------------------------------------------------
+\reviewer{5. In connection with the discussion of relativity and 
 outer-shell electrons (p.3) it might be instructive to mention that 
@@
 
-\vspace{0.1cm}
-The referee is correct. We included the following modifications in the text
-(lines 200-214):
-
-\vspace{0.1cm}
-{\small ``One would expect the relativistic contributions to only affect 
-significantly the inner shells. The relativistic binding energies of the 
-electrons belonging to these shells are 2.5\% from the experimental values,
-while the non-relativistic ones are within $\sim$10\%. Moreover, we notice 
-the relativistic corrections greatly influences the outer $5p$ and $4f$ 
-shells. This effect may be a consequence of the orthogonality with the 
-relativistic inner shells. Also, the sign of the binding energy deviations 
-is inverted near the FEG limit, i.e. the electrons go from being more 
-bounded --than the experimental values-- to being less bounded. 
-This change of sign can be attributed to the fact that the experimental 
-values correspond to hafnium in solid-state, while our theoretical 
-calculations correspond to the element in the gas phase.''}
+%\reply{We included two new paragraphs (lines 211-230) to make reference to the referee's comments on the binding energy calculations.}
+\reply{The referee is correct. We included two paragraphs in Section III
+(lines 211-230) to refer to her/his comments.}
 
 % ----------------------------------------------------------------------
-\vspace{0.25cm}
-\textsl{6. The importance of 4f-5p screening is documented clearly 
+\reviewer{6. The importance of 4f-5p screening is documented clearly 
 (p.3), but the reader will ask why this does not affect other shells. 
@@
 
-\vspace{0.1cm}
-{\color{red}Reply here.}
+%\reply{The mixing criteria is explained on page 3: we consider the quantum uncertainty in energy so bound electrons with similar energies and mean radio $<r>$ response as a single density of electrons including screening among them.
+We recognize that a more detailed study is needed, we are developing such work as a systematic for different targets involving 4f electrons.
+What we have done here for Hf is to check the importance of the contribution to the total stopping.
+Only the 5p-4f mixing have influence in the total stopping because they are the second most important contribution after the FEG.
+Some other shells may be mixed, and we calculate them, but the total stopping changes in very small $\%$.
+We added lines 262-264 to page 4.} {\small '' As already mentioned, at higher energies inter-shell screening is possible for other subshells (i.e.
+4p-4d for impact energies above 0.9 MeV), but its weight in the total stopping is minor for deeper shells.'' }
+\reply{The mixing criterion is explained on page 3; we considered
+the quantum uncertainty in energy so that bound electrons with similar
+energy values and mean radio $\langle r\rangle$ response as a single
+density of electrons, including screening among them. We acknowledge
+that a more detailed study is required. We are currently working on
+developing a systematic approach for different targets involving 4f
+electrons. So far, we have corroborated the importance of the
+contribution of the 4f shell to the total stopping in Hf. We have also
+determined that the 5p-4f mixing influences the total cross section
+mainly because they are the second most significant contribution after
+the FEG. Other shells may be mixed, and we have computed them, but the
+total stopping varies very little. Reference to this analysis is given
+in lines 258-262.}
 
 % ----------------------------------------------------------------------
-\vspace{0.25cm}
-\textsl{7. In view of the large contribution of the outermost shell, I 
-suggest to include the core contribution in figure 3.}
+\reviewer{7. In view of the large contribution of the outermost shell, 
+I suggest to include the core contribution in figure 3.}
 
-\vspace{0.1cm}
-{\color{red}Reply here.}
+%\reply{We thank the referee for this suggestion, actually we had this figure, but we doubted on including all the curves.
+Now Fig.
+3 includes separate FEG and bound electron contribution, and the total curves too.}
+\reply{We thank the referee for her/his suggestion. At first, we thought 
+of showing all contributions separately but opted for a simpler version.
+Upon recommendation, we have included in Fig. 3 of the manuscript the
+FEG and bound electron contribution curves.}
 
 % ----------------------------------------------------------------------
-\vspace{0.25cm}
-\textsl{8. In addition to comparisons with experiments and empirical 
+\reviewer{8. In addition to comparisons with experiments and empirical 
 models, it would be instructive to see comparisons with theoretical 
@@
 
-\vspace{0.1cm}
-{\color{red}Reply here.}
+%\reply{We thank this suggestion.
+We had already compared with Casp5.2 and DPASS for ourselves before submitting.
+In this version we added CasP5.2 and DPASS to figure 4, we added a mention in the introduction(last paragraph, lines 68-73), and the following comment in page 6, lines 315-322:} {\small ''The stopping maximum is a very sensitive region for any full theoretical description, with no parameters at all, and this is quite visible in a linear-scale plot like Fig.
+[4].
+However, the impact energy for the maximum seem to agree between our curve and DPASS, though $10 \%$ below.
+Instead, CasP predicts a stopping maximum $10 \%$ above and a completely different shape at lower energies.''}
+\reply{We compared our results with Casp5.2 and DPASS before submission,
+but we were discouraged to include them by the number of curves in the
+figure. However, we have now managed to include CasP5.2 and DPASS to
+figure 4. We comment about these changes in the Introduction 
+(lines 71-72), and in Section IV (lines 308-311).}
 
@@
 % ----------------------------------------------------------------------
+\reviewer{1. Can the authors discuss how possible impurities in the 
+sample and -especially- on both sides of the target would affect the 
+final stopping data? The authors must discuss it in the manuscript.}
 
-\textsl{1. Can the authors discuss how possible impurities in the sample and -especially- on both sides of the target would affect the final
-stopping data? The authors must discuss it in the manuscript.}
-
-\vspace{0.1cm}
-{\color{red}Reply here.}
+\reply{Pedro Miranda...................................................}
 
 % ----------------------------------------------------------------------
-\vspace{0.25cm}
-\textsl{2. In the same context, the authors have to clarify whether potential
-surface roughness and/or target non-uniformity might affect the final
-results obtained in transmission method.}
+\reviewer{2. In the same context, the authors have to clarify whether
+potential surface roughness and/or target non-uniformity might affect 
+the final results obtained in transmission method.}
 
-\vspace{0.1cm}
-{\color{red}Reply here.}
+\reply{Pedro Miranda...................................................}
 
 % ----------------------------------------------------------------------
-\vspace{0.25cm}
-\textsl{3. What is accuracy of the primary beam energy of the accelerator?
-Moreover: when the target is placed and removed from the front of the
-detector, does the detector itself need to unbiased and the chamber
-opened? If so, might there be any offset ($\sim$keV) in the detection
-system that would affect the energy distribution peaks, hence the
-Gaussian position from the fits (Fig. 1)?}
+\reviewer{3. What is accuracy of the primary beam energy of the 
+accelerator? Moreover: when the target is placed and removed from the 
+front of the detector, does the detector itself need to unbiased and the 
+chamber opened? If so, might there be any offset ($\sim$keV) in the 
+detection system that would affect the energy distribution peaks, hence 
+the Gaussian position from the fits (Fig. 1)?}
 
-\vspace{0.1cm}
-{\color{red}Reply here.}
+\reply{Pedro Miranda...................................................}
 
 % ----------------------------------------------------------------------
-\vspace{0.25cm}
-\textsl{4. The experimental data above $\sim$1.5 MeV is systematically 
+\reviewer{4. The experimental data above $\sim$1.5 MeV is systematically 
 below the data from Ref. [7], hence SRIM semi-empirical approach, and 
 also below the authors’ theory output. At these energies (and higher), 
-the agreement between experiment and theory is expected to be better than
-a few percent (that is why you used stopping power for 5.486 MeV alpha
-on Hf to obtain the target thickness). Can the authors explain why
+the agreement between experiment and theory is expected to be better 
+than a few percent (that is why you used stopping power for 5.486 MeV 
+alpha on Hf to obtain the target thickness). Can the authors explain why
 there is such systematic difference for the present measurements using
@@
 
-\vspace{0.1cm}
-{\color{red}Reply here.}
+% \reply{There is a small systematic difference between our experiments and theory for impact energies above 1.5 MeV, but it is within the experimental error.
+Figure 2 of this reply amplifies the results shown in the manuscript by focusing on the high energy region.
+Note that we included our theoretical results, SRIM, Casp5.2, and DPASS in the figure.
+The difference is small, being less than 3\% for the two values above 2 MeV.
+We included comments regarding this subject in the manuscript.
+<=== Pedro, PLEASE CHECK THIS}
+\reply{There is a small systematic difference between our experiments and theory for impact energies above 1.5 MeV, but it is within the experimental error.
+Figure 2 of this reply amplifies the results shown in the manuscript by focusing on the high energy region.
+Note that we included our theoretical results, SRIM, Casp5.2, and DPASS in the figure.
+The difference is small, being less than 3\% for the two values above 2 MeV.
+We included comments regarding this subject in the manuscript.}
 
-% ----------------------------------------------------------------------
-\vspace{0.25cm}
-\textsl{5. The authors claim an overall uncertainty of 5\%, but it is not clear
-how they obtained this value, or whether only statistical
+%-----------------------------------------------------------------------
+\begin{figure}[!t]
+\centering
+\includegraphics[width=12.cm]{Fig04.eps}
+\caption{Stopping at high impact energies. Theoretical curves and
+symbols for data are explained inside the figure.}
+\label{highE}
+\end{figure}
+
+%-----------------------------------------------------------------------
+\reviewer{5. The authors claim an overall uncertainty of 5\%, but it is 
+not clear how they obtained this value, or whether only statistical
 contributions were taken into account. Nevertheless, uncertainties of
@@
 
-\vspace{0.1cm}
-{\color{red}Reply here.}
+\reply{Pedro Miranda...................................................}
 
 % ----------------------------------------------------------------------
-\vspace{0.25cm}
-\textsl{6. Can the present theory in the manuscript be compared to other
-models (e.g., CasP and DPASS), especially towards low energies? Aiming
-thus a constructive discussion on the overall-agreement among
-different models to predict SCS of “challenge” materials, as
-transition metals (f-subshell)?}
+\reviewer{6. Can the present theory in the manuscript be compared to 
+other models (e.g., CasP and DPASS), especially towards low energies? 
+Aiming thus a constructive discussion on the overall-agreement among
+different models to predict SCS of “challenge” materials, as transition 
+metals (f-subshell)?}
 
-\vspace{0.1cm}
-{\color{red}Reply here.}
+% \reply{We thank the referee for this suggestion that improves this new version.
+We have added CasP5.2 and DPASS curves in Figure 4, we have also added a mention in the introduction (last paragraph, lines 68-73), and the following comment in page 6, lines 315-322:} \\ {\small ''The stopping maximum is a very sensitive region for any full theoretical description, with no parameters at all, and this is quite visible in a linear-scale plot like Fig.
+[4].
+However, the impact energy for the maximum seem to agree between our curve and DPASS, though $10 \%$ below.
+Instead, CasP predicts a stopping maximum $10 \%$ above and a completely different shape at lower energies.''}
+\reply{We thank the referee for her/his suggestion. We have included the 
+CasP5.2 and DPASS curves in Fig. 4 of the manuscript. We made the 
+corresponding modification in the Introduction (lines 71-72), and 
+Section IV (lines 308-311).}
 
 % ----------------------------------------------------------------------
-\vspace{0.25cm}
-\textsl{7. In the last sentence of Sec. IV, the authors point out for a need
-of more experimental data below 100 keV H+ but, in my opinion, more
-data are needed even at higher energies, e.g., nearby (and slightly
+\reviewer{7. In the last sentence of Sec. IV, the authors point out for 
+a need of more experimental data below 100 keV H+ but, in my opinion, 
+more data are needed even at higher energies, e.g., nearby (and slightly
 above) the Bragg peak... Please enhance discussions about it.}
 
-\vspace{0.1cm}
-{\color{red}Reply here.}
+% \reply{We agree with this comment by the referee.
+In a ion-target system with only two set of measurements more data is needed.
+In the energy region 80-500 keV only the data by Sirotinin group is available and cannot be compared with others.
+SRIM is not a test for this data, SRIM includes this data in the adjust (more detail about this can be found in the answer 4 to the first referee report).
+We change the comments in page 6 lines 328-336 and lines 364-366 as follows:} \\ {\small ''Future experiments  would be important to complete this study on the stopping of Hf for protons.
+In the energy region $80-500$ keV only the measurements by Sirotinin group in 1984 are available.
+SRIM clearly includes these values in the adjust of the code, so it does not represent a test for this data.
+Undoubtedly, new measurements are needed in three specific energy regions: below 500 keV, around the stopping maximum (i.e.
+$50-200$ keV) and at low impact energies''} \\ {\small ''Future experiments for impact energies below 500 keV, around the stopping maximum and in the low energy region would be important to complete this study.''}
+\reply{We agree with the referee's comment.
+In an ion-target system with only two sets of measurements, more data is needed.
+In the energy region 80-500 keV, only the data by Sirotinin et al.
+is available and cannot be compared with others.
+SRIM is not a test for this data since it adjusts these measurements itself (a more detailed discussion about the matter can be found in reply 4 of the first referee report).
+We changed our manuscript accordingly in Section IV (lines 331-339) and Section V (lines 374-377).}
 
@@
 % ----------------------------------------------------------------------
+\reviewer{1. Regarding the theoretical model that describes the
+contribution to the stopping power of the bound electrons, the authors
+emphasize the importance of including relativistic effects in the
+description of the binding energies and densities of the Hf 1s-4f
+states. However, it is not clear if the calculations performed by the
+authors are done for gas phase Hf atoms or for the solid Hf as in the
+experiments of Ref. [38]. In case of being for the former, the slightly
+worse agreement found for the 4f states as compared to the experimental
+values could be related to this point.}
 
-\textsl{
-- Regarding the theoretical model that describes the contribution to
-the stopping power of the bound electrons, the authors emphasize the
-importance of including relativistic effects in the description of the
-binding energies and densities of the Hf 1s-4f states. However, it is
-not clear if the calculations performed by the authors are done for
-gas phase Hf atoms or for the solid Hf as in the experiments of Ref.
-[38]. In case of being for the former, the slightly worse agreement
-found for the 4f states as compared to the experimental values could
-be related to this point.}
-
-\vspace{0.1cm}
-Our calculations are performed for gas phase Hf atoms. We have clarify
-this in the manuscript (lines 211-214).
+%\reply{The referee is correct, our calculations are performed for gas phase Hf atoms.
+We have explicitly clarify the difference between the experimental and theoretical binding energies in page 3, lines 215-219.}
+\reply{The referee is correct; our calculations are performed for gas-
+phase Hf atoms. We have now explicitly clarified the difference between 
+the experimental and theoretical binding energies in the manuscript 
+(lines 215-219).}
 
 % ----------------------------------------------------------------------
-\vspace{0.25cm}
-\textsl{
-1. The results from the three different theoretical calculations shown
-in Fig. 3 are not completely clear. The authors emphasize the
-importance of considering both screening effects and relativistic
-effects for the 4f and 5p electrons, but it is not clear if later in
-the text the authors use indistinctly the terminology ``relativistic''
-and ``screening'' to refer to the same issue. For instance, are the ML
-results shown in Fig. 3 calculated with the relativistic binding
-energies and densities, but with and without including screening
-corrections, or do the two ML curves differ in whether the
-relativistic corrections are or not included? It is neither clear what
-description of the bound electrons is used in the curve labeled as
-``SPCC(FEG)+SLPA(bound)''.}
+\reviewer{2. The results from the three different theoretical
+calculations shown in Fig. 3 are not completely clear. The authors
+emphasize the importance of considering both screening effects and
+relativistic effects for the 4f and 5p electrons, but it is not clear if
+later in the text the authors use indistinctly the terminology
+``relativistic'' and ``screening'' to refer to the same issue. For
+instance, are the ML results shown in Fig. 3 calculated with the
+relativistic binding energies and densities, but with and without
+including screening corrections, or do the two ML curves differ in
+whether the relativistic corrections are or not included? It is neither
+clear what description of the bound electrons is used in the curve
+labeled as ``SPCC(FEG)+SLPA(bound)''.}
 
-\vspace{0.1cm}
-{\color{red}Reply here.}
+%\reply{Thank you for the comment that helps us to clarify this point.
+Our SLPA calculations use the relativistic wave functions and binding energies for the bound electron contribution to the stopping power.
+The three curves in Fig.
+3 and the curve in Fig.
+4 are relativistic in this sense.
+We have added an explicit mention about the use of relativistic wave functions and binding energies in page 4 lines 200-201, and also this comment in page 4, lines 287-289:} \\ {\small ''Bound $1s$-$4f$ electrons (relativistic wave functions and binding energies) are calculated with the SLPA and shown separately in Fig.
+3''}
+\reply{We thank the referee for her/his comment since it will help us 
+clarify this point. Our SLPA calculations use relativistic wave 
+functions and binding energies for the bound electron contribution to 
+the stopping power. In this sense, the three curves in figure 3 and the 
+solid curve in Fig. 4 are called relativistic. We have added explicit 
+mention of the use of relativistic wave functions and binding energies 
+on page 4 lines 244-246 and lines 283-286.}
 
 % ----------------------------------------------------------------------
-\vspace{0.25cm}
-\textsl{
-2. If all the results of Fig. 3 were calculated for the relativistic
-description, it would be also interesting to show the results obtained
-with the non-relativistic description in order to demonstrate the
-importance of this correction.}
+\reviewer{3. If all the results of Fig. 3 were calculated for the 
+relativistic description, it would be also interesting to show the
+results obtained with the non-relativistic description in order to
+demonstrate the importance of this correction.}
 
-\vspace{0.1cm}
-{\color{red}Te mando adjunto las funciones de onda y energías de ligaduras
-del caso no relativista (NR). Habría que ver como queda.}
+%\reply{The SLPA calculations for the bound electrons uses the Hartree-Fock wave funtions and binding energies, which are available in the tables by Clementi-Roetti or by Bunge for elements up to Xe (Z=54).
+The errors in the non-relativistic calculations worsen for higher Z atoms, and this is the reason why these tables do not extend any further than Z=54.
+The fail of non-relativistic values of binding energies is shown in Fig.
+2.
+We have added a comment on this in page 3, lines 214-228.} \\ {\small ``For more detail, Fig.
+2 (b) shows  the relative errors with respect to the experimental  values.
+This figure shows clearly that the relativistic corrections are very important to describe the atomic structure of hafnium, even for the outer shells.
+It turns out that the errors committed in the non-relativistic calculations of the inner shell orbitals propagate, through the Hartree-Fock approximation, to the outer shells.The importance of full-relativistic calculations for the outer shells has already been noted before for Au, Pb and Bi and W.''} \\ {\color{red} When dealing with stopping in targets with Z>54 we need to face our own calculation of the wave functions and binding energies.
+And we do it within the relativistic approximation to the best of our possibilities.
+However, one can wonder if such discrepancy in the non-relativistic results affects the stopping power.
+First, it will only affect the bound electron contribution to the total stopping.
+Second, at high energies (i.e.
+1 MeV) the results converge to Bethe limit for any density of electrons as far as the correct number of target electrons is included.
+Having clarified this, we include in the figure \ref{NR} here, the total stopping using the non-relativistic values we calculate using the autostructure code.
+As can be noted, at sufficiently high energies it is the same, but differs around the stopping and at lower impact energies.
+We have included in this figure and in Fig.
+4 of the manuscript the DPASS results by Sigmund.}
+\reply{The SLPA calculations for the bound electrons uses the Hartree-
+Fock wave functions and binding energies, which are available in the 
+tables by Clementi-Roetti or by Bunge for elements up to Xe (Z=54). The 
+errors in the non-relativistic calculations worsen for higher Z atoms, 
+and this is the reason why these tables do not extend any further than 
+Z=54. The fail of the non-relativistic description of the binding 
+energies is shown in Fig. 2. We have added a comment on this on page 3, 
+lines 220-230. When dealing with stopping in targets with Z>54, we need 
+to compute the wave functions and binding energies ourselves. And we do 
+it within the relativistic approximation to the best of our 
+possibilities. However, one can wonder if such a discrepancy in the non-
+relativistic results affects the stopping power. First, it will only 
+affect the bound electron contribution to the total stopping. Second, at 
+high energies (i.e., 1 MeV), the results converge to the Bethe limit for 
+any density of electrons as far as the correct number of target 
+electrons is included. Having clarified this, we added in figure 3 of 
+this reply, the total stopping using the non-relativistic values we 
+calculated using the AutoStructure code. As can be noted, at 
+sufficiently high energy values, the results agree, but they start to 
+differ around the stopping and at lower impact energies. We have
+included in this figure, and in Fig. 4 of the manuscript, the DPASS
+results by Sigmund.}
 
+%-----------------------------------------------------------------------
+\begin{figure}[!t]
+\centering
+\includegraphics[width=12.cm]{Fig05.eps}
+\caption{Total stopping cross section. Present results using
+relativistic and non-relativistic solutions for hafnium atom. 
+Curves and symbols explained inside the figure.}
+\label{NR}
+\end{figure}
 % ----------------------------------------------------------------------
-\vspace{0.25cm}
-\textsl{
-3. Furthermore, it would also be meaningful to show the results from
+\reviewer{
+4. Furthermore, it would also be meaningful to show the results from
 the ML(FEG) calculation alone in order to identify better the
@@
 
-\vspace{0.1cm}
-{\color{red}Reply here.}
+%\reply{Thank you for this suggestion. We have added the curves for FEG and bound electrons to Fig. 3, and the corresponding comments in the text and in the figure caption.}
+\reply{Thank you for this suggestion. We have added the curves for FEG and bound electrons to Fig. 3 of the manuscript, and the corresponding comments in the text and figure caption.}
 
 % ----------------------------------------------------------------------
-\vspace{0.25cm}
-\textsl{
-4. When comparing the new and existing experimental data, clarify what
+\reviewer{
+5. When comparing the new and existing experimental data, clarify what
 experimental technique was used in Ref. 7 to measure the stopping
@@
 
-\vspace{0.1cm}
-{\color{red}Reply here.}
+\reply{Good suggestion, thank you. Sirotinin et al. used the 
+backscattering method for these measurements. We have made reference to 
+the method implemented in the Introduction (line 70) and Section IV 
+(line 305).}
 
 % ----------------------------------------------------------------------
-\vspace{0.25cm}
-\textsl{
-5. The data by Sirotonin et al. plotted in Fig. 4 seem to suggest that
+\reviewer{
+6. The data by Sirotonin et al. plotted in Fig. 4 seem to suggest that
 the stopping power maximum would be around 100-200 keV, while the
@@
 
-\vspace{0.1cm}
-{\color{red}Reply here.}
+\reply{The stopping maximum is very sensitive to the theoretical model and to the experimental method.
+Sirotinin paper mention "possible reasons for the disagreement of the measurement results near the E(E) curve maximum".
+%The original paper by Sirotinin says: "The possible reasons for the disagreement of the measurement results near the E(E) curve maximum are mainly related to the inaccuracy in determining thickness, to nonuniformity of the thickness, and the neglect of the effect of the chemical composition and structure of the thin films used in the measurements"
+For targets with many different sets of data (i.e.
+H in Ag, Al, Au, Fe, Cu, Si, and others) it is usual to note great dispersion of values around the maximum.
+In the case of Hf, only one set of data is available and it is perfectly matched by SRIM semi-empirical curve.
+SRIM clearly includes these values in the adjust of the code, so it does not represent a test for this data.
+One of the goals of this work is to call the attention on possible differences between theory and the existing values, and to remark the necessity of more experimental data in certain energy regions, mainly, the stopping maximum and below.
+We have changed the comments in page 6 lines 328-336 and lines 364-366 as follows:} \\ {\small ''Future experiments  would be important to complete this study on the stopping of Hf for protons.
+In the energy region $80-500$ keV only the measurements by Sirotinin group in 1984 are available.
+SRIM clearly includes these values in the adjust of the code, so it does not represent a test for this data.
+Undoubtedly, new measurements are needed in three specific energy regions: below 500 keV, around the stopping maximum (i.e.
+$50-200$ keV) and at low impact energies''} \\ {\small ''Future experiments for impact energies around the stopping maximum and in the low energy region would be important to complete this study.''} \\ {\color{red} In relation with the referee comment about the maximum of the stopping, in Fig.
+4 we have added the comparison with other two theoretical results: CASP and PASS.
+The following comment has been included in page 6, lines 315-322:} \\ {\small ''The stopping maximum is a very sensitive region for any full theoretical description, with no parameters at all, and this is quite visible in a linear-scale plot like Fig.
+[4].
+However, the impact energy for the maximum seem to agree between our curve and DPASS, although DPASS is $10 \%$ below ours.
+Instead, CasP predicts a stopping maximum $10 \%$ above our value and a completely different shape at lower energies.''}
 
 % ----------------------------------------------------------------------
-\vspace{0.25cm}
-\textsl{
-6. When reviewing on the experimental work performed on Hf oxides in
+\reviewer{
+7. When reviewing on the experimental work performed on Hf oxides in
 the introduction, note that there is a quite recent work by Roth and
@@
 
-\vspace{0.1cm}
-We thank the referee for suggesting the missing reference. We have now
-include it in the revised manuscript.
+\reply{We thank the referee for suggesting this missing reference. We have included it in the revised manuscript.}
 
 % ----------------------------------------------------------------------
-\vspace{0.25cm}
-\textsl{Finally, there are few misprints in the generated pdf file:}
+\reviewer{Finally, there are few misprints in the generated pdf file:}
 
-\textsl{- Y-axis labels in Figs. 3 an 4 are unreadable }
-{\color{red} ¿Cómo quedó el preprint? El pdf que genera mi latex tiene 
-bien los labels, así que debe un problema del generador de latex del PRA...}
+\reviewer{- Y-axis labels in Figs. 3 an 4 are unreadable }
 
-\st{\textsl{- Page 3,2nd column, 2nd full paragraph: "Too assess"}}
+\reply{We have changed to larger prints  and have written explicitly "stopping cross section" -Y- and "Impact energy" -X- instead of S(E) and E. }
 
-\st{\textsl{- Caption of Fig.2: "hollow circles" -$>$ "open circles" or
-"unfilled circles"}}
+\reviewer{- Page 3,2nd column, 2nd full paragraph: "Too assess"}
+
+\reply{Corrected.}
+
+\reviewer{- Caption of Fig.2: "hollow circles" -$>$ "open circles" or
+"unfilled circles"}
+
+\reply{Corrected.}
+
+\reply{We have also made some improvements in the English redaction of the manuscript.}
 

### 2020-05-11 sp_hafnium.tex
checked english with grammarly

@@
 %Title of paper
-\title{Stopping power of hafnium by proton impact and \\
+\title{The stopping power of hydrogen in hafnium and \\
 the importance of relativistic $4f$ electrons}
@@
 experimentally and theoretically. The measurements were performed at 
-the Laboratory of Accelerators and X-Ray Diffraction in Lisbon, by 
+the Laboratory of Accelerators and X-Ray Diffraction in Lisbon by 
 using the transmission method on self-supporting stopping material. 
@@
 as a free electron gas (FEG). We found the relativistic description of 
-the $4f$-shell and the screening between $4f$ and $5p$ electrons are 
+the $4f$-shell and the screening between $4f$ and $5p$ electrons to be 
 decisive around the stopping maximum. Present theoretical and 
-experimental results are in very good agreement in the energy region 
-of the new measurements. However, the present theoretical stopping 
+experimental results are in excellent agreement in the energy region 
+of the new measurements. However, our theoretical stopping 
 cross sections show substantial differences with the most used 
 semi-empirical models (SRIM2013 and ICRU-49) at intermediate to low 
-energies. Our calculations suggest the stopping maximum higher and 
+energies. Our calculations suggest the stopping maximum to be higher and 
 shifted to lower energies than these previous predictions. Future 
-measurements below 100 keV would be decisive for a better knowledge 
-around the maximum and below.
+measurements below 100 keV would be necessary for a better understanding 
+of the stopping power around the maximum and below.
 \end{abstract}
@@
 electrons~\cite{Chu01,Sigmund}. The information given by the energy 
-loss process is essential not only to have a better knowledge of the 
-physics behind the fundamental interactions, but also because it plays 
-an important role in many applied fields such as materials science, 
-nuclear physics, ionic implantation and radiotherapy~\cite{Sigmund,Schardt}. 
+loss process is essential not only to have a better comprehension of the 
+physics behind the fundamental interactions but also because it plays 
+a vital role in many applied fields such as materials science, 
+nuclear physics, ionic implantation, and radiotherapy~\cite{Sigmund,Schardt}. 
 Experimental data on ion mean energy loss per unit path $S(E)$ is of 
@@
 length~\cite{Choi,Robertson}. Some important physical properties of the 
-above mentioned metal-oxide films depend on its thickness, which is 
-often measured by using Rutherford Backscattering Spectrometry~\cite{Alfassi01,Tesmer01}, 
-a method that relies heavily on the determination of the stopping power 
-of ion beams in the material of interest.
+above mentioned metal-oxide films depend on their thickness, which is 
+often measured by using Rutherford Backscattering 
+Spectrometry~\cite{Alfassi01,Tesmer01}. This method relies heavily on 
+the determination of the stopping power of ion beams in the material of 
+interest.
 
-In this study, we report experimental stopping power cross sections 
-over the incident energy range (0.6-2.5) MeV for protons crossing 
-self-supported Hf thin-film by using the transmission method. We aim 
-not only to upgrade stopping power data compilations~\cite{HPaul03,mondim17} 
-but also to provide useful information about the processes governing 
-the slowing down of protons in multi-electronic targets. In the rare 
-earth metals, the $4f$ electrons play a relevant role in the stopping 
-power, being the first shell of bound electrons below the conduction 
-band. As already noted \cite{Roth17}, the free electron gas (FEG) shows 
-unexpected behavior in these elements, which casts doubts on its proper 
-description. In the case of Hf, we found the contribution of the 
-$4f$-shell decisive even at impact energies around the stopping maximum, 
-as shown later.
+In this study, we report experimental stopping power cross sections over 
+the incident energy range (0.6-2.5) MeV for protons crossing 
+self-supported Hf thin-film by using the transmission method. We aim not 
+only to upgrade stopping power data compilations~\cite{HPaul03,mondim17} 
+but also to provide useful information about the processes governing the 
+slowing down of protons in multi-electronic targets. In the rare earth 
+metals, the $4f$ electrons play an essential role in the stopping power 
+since they belong to the first shell of bound electrons below the 
+conduction band. As already noted \cite{Roth17}, the free electron gas 
+(FEG) shows unexpected behavior in these elements, which casts doubts 
+on its proper description. In the case of Hf, we found the contribution 
+of the $4f$-shell to be decisive even at impact energies around the 
+stopping maximum, as will be shown later.
 
@@
 local plasma approximation (SLPA)~\cite{mon13} to describe the energy 
-transferred to the bound $1s$-$4f$ electrons, and two different models 
-for the FEG: the screened potential with cusp condition model (SPCC)~\cite{mon17}, 
-which is non-linear binary formalism; and the Mermin-Lindhard dielectric 
-formalism (ML)~\cite{Mermin}, for energies around the stopping maximum 
-and above. Our model requires the relativistic wave functions and 
-binding energies of Hf, and considers 4 electrons per atom in the 
-FEG~\cite{mendez2019}. Hf has the extra interest of the filled 
-$4f$-subshell (with 14 electrons) as the main contribution below the 
-FEG, causing the stopping cross sections to be very sensitive to a good 
-description of this shell. The screening among the $4f$ and $5p$ 
-electrons has been considered and found to play a major role within the 
-SLPA calculations.
+transferred to the bound $1s$-$4f$ electrons and two different models 
+for the FEG; in low energy region, the screened potential with cusp 
+condition model (SPCC)~\cite{mon17}, which is non-linear binary 
+formalism, and the Mermin-Lindhard dielectric formalism 
+(ML)~\cite{Mermin} for energies around the stopping maximum and above. 
+Our model requires the relativistic wave functions and binding energies 
+of Hf and considers four electrons per atom in the FEG~\cite{mendez2019}. 
+Hafnium is particularly interesting since the filled $4f$-subshell (with 
+14 electrons) is the main contributor below the FEG, causing the 
+stopping cross sections to be very sensitive to a good description of 
+this shell. The screening among the $4f$ and $5p$ electrons has been 
+considered and found to play a significant role within the SLPA 
+calculations.
 
@@
 Present theoretical and experimental values are finally compared to the 
-available experimental values~\cite{Sirotinin}, the ICRU-49 
-calculations~\cite{ICRU49} and the SRIM-2013 package~\cite{Ziegler01}. 
+only experimental values measured by Sirotinin {\it et al.}~\cite{Sirotinin}  
+with the backscattering method, the theoretical results by Grande and 
+Schiwietz~\cite{Grande,casp52}, and by Sigmund and Schinner~\cite{DPASS20}, 
+and also with the semi-empirical values from the SRIM-2013 
+package~\cite{Ziegler01} and the ICRU-49 tabulation~\cite{ICRU49}. 
 Conclusions and discussions are given in Section~\ref{conclusion}. All 
@@
 \caption{RBS spectrum for $E_{\mathrm{avg}}=921.1$ keV protons on 
-hafnium sample which is subsequently used to determine the energy loss 
-in the foil.}
+the hafnium sample, which is subsequently used to determine the energy 
+loss in the foil.}
 \label{F01}
@@
 accelerator to deliver $^1$H$^+$ primary ion beams through a series of 
-electrostatic lenses and collimators onto a thin Au/SiO$_2$ sample which 
-is used as a scattering center. This sample was placed in the center of 
-a RBS/C scattering chamber, where a high vacuum (pressure of 
+electrostatic lenses and collimators onto a thin Au/SiO$_2$ sample, 
+which is used as a scattering center. This sample was placed in the 
+center of an RBS/C scattering chamber, where a high vacuum (pressure of 
 $\sim$10$^{-6}$ Torr) was maintained during the measurements. The beam 
@@
 \subsection{Target}
-The stopping material under analysis was a hafnium foil with nominal 
-thickness of 1.0 $\mu$m and 99.95\% purity which was supplied by Lebow 
+The stopping material under analysis was a hafnium foil with a nominal 
+thickness of 1.0 $\mu$m and 99.95\% purity, which was supplied by Lebow 
 Company~\cite{Lebow}. However, a more precise thickness value was 
@@
 found in Ref.~\cite{Ziegler01} to obtain an areal density of 
-($4.13 \pm 0.21$)$\times 10^{19}$ at/cm$^2$ which corresponds to a 
+($4.13 \pm 0.21$)$\times 10^{19}$ at/cm$^2$, which corresponds to a 
 thickness of $0.920\pm0.046 \mu$m.
@@
 Once the beam impinges on the Au/SiO$_2$ sample, protons are 
-back-scattered towards a Si surface barrier detector located at 
+backscattered towards a Si surface barrier detector located at 
 140$^{\circ}$ relative to the initial beam direction. Fig.~\ref{F01} 
-shows two particle spectra where ion energies $E_2$ and $E_1$ are 
-associated to a placed and removed hafnium sample, respectively. Both 
+shows two particle spectra, where the ion energies $E_2$ and $E_1$ are 
+associated with a placed and removed hafnium sample, respectively. Both 
 energy distributions were fitted by Gaussian functions to obtain the 
@@
 loss fraction $\Delta E/E_{\mathrm{avg}}$ across the Hf foil is not 
-exceeding 20\% it is possible to define the stopping cross section by 
+exceeding 20\%, it is possible to define the stopping cross section by 
 \cite{Raisanen01,Schulz01}:
@@
 correction to the mean energies $E_{\mathrm{avg}}$ was applied in order 
-to account for the nonlinear dependence on ion energy of stopping 
+to account for the non-linear dependence on ion energy of stopping 
 powers \cite{Chilton,Rajatora}.
@@
 conduction band, which is well approximated by a free electron gas (FEG) 
-of Fermi velocity $v_F$. Above certain velocity (i.e. $v\geq 1.5\,v_F$), 
-not only binary but also collective excitations (plasmons) 
-occur~\cite{mon17}. Moreover, at high energies, also the bound electrons 
-contribute to the stopping power. The method used in this work combines 
-a description for the interaction with the valence (or conduction) 
-electrons as a FEG, and a different one for the interaction with the 
-bound electrons.
+of Fermi velocity $v_F$. Above a particular velocity value (i.e., 
+$v\geq 1.5\,v_F$), not only binary but also collective excitations 
+(plasmons) occur~\cite{mon17}. Moreover, at high energies, also the 
+bound electrons contribute to the stopping power. The method used in 
+this work combines a FEG description for the interaction with the 
+valence (or conduction) electrons and a different one for the 
+interaction with the bound electrons.
 
@@
 to the projectile. This model proved to give a good description of the 
-induced electron density even for negative projectiles~\cite{mon17}, 
+induced electron density even for negative projectiles~\cite{mon17} 
 and reproduces the low velocity proton-antiproton differences in the 
@@
 Wigner-Seitz radio, $r_S$, which is a measure of the electronic density 
-of the FEG. For metals of well-known $r_S$, the SPCC describes correctly 
-the low energy experimental stopping data~\cite{mon17}, agreeing with 
+of the FEG. For metals of well-known $r_S$, the SPCC describes the low 
+energy experimental stopping data correctly~\cite{mon17}, agreeing with 
 the DFT results by Echenique and coworkers~\cite{eche81,nagy89} at $v=0$.
 
-Hafnium ($Z=72$, [Xe] $4f^{14}\,6s^2\,5d_{3/2}^1\,5d_{5/2}^1$), belongs 
+Hafnium ($Z=72$, [Xe] $4f^{14}\,6s^2\,5d_{3/2}^1\,5d_{5/2}^1$) belongs 
 to the first groups of transition metals, with four electrons as FEG 
-(theoretical $r_S=2.14$ a.u.) and $1s$-$4f$ electrons bound. We compared 
-the computed $r_S$ with the experimental value obtained from measured 
-energy loss function by Lynch \textit{et al} \cite{lynch75}. The 
+($r_S=2.14$ a.u.) and $1s$-$4f$ electrons bound. We compared the 
+computed $r_S$ with the experimental value obtained from the measured 
+energy loss function by Lynch \textit{et al.}~\cite{lynch75}. The 
 experimental plasmon energy of Hf is $\hbar\omega_P \approx 15.8$ eV, 
-with a width at half maximum $\delta \approx 4.4$ eV, and 
-$r_S \approx 2.07$ a.u. \cite{lynch75}. The difference of less than 
+with a width at half maximum $\delta\approx 4.4$ eV, and 
+$r_S\approx 2.07$ a.u.~\cite{lynch75}. The difference of less than 
 $5\%$ between theoretical and experimental $r_S$ assess Hf as a 
@@
 
-Above certain impact velocity, the plasmon contribution is important 
-(i.e. around and above the stopping maximum). An interesting value for 
+Above certain impact velocity, the plasmon contribution is essential
+(i.e., around and above the stopping maximum). A value of interest for 
 our analysis is the minimum impact velocity to excite plasmons, $v_P$. 
@@
 electrons to the ion passage is described through the quantum dielectric 
-function, depending on the characteristic $r_S$ and $\delta$ parameters 
+function, which depends on the characteristic $r_S$ and $\delta$ parameters 
 of the FEG.
@@
 \includegraphics[width=11.cm]{bindener.eps}
-\caption{(a) Binding energies of Hf. Non-relativistic and relativistic 
-calculations are given with filled symbols. Experimental measurements 
-for solids~\cite{williams1995} are depicted with open circles. 
-(b) Corresponding relative errors respect to experimental data.}
+\caption{(a) Binding energies of Hf. Present relativistic and available 
+non-relativistic~\cite{badnell97} values are given with filled symbols. 
+Experimental measurements for solids~\cite{williams1995} are depicted 
+with open circles. (b) Corresponding relative errors with respect to 
+experimental data.}
 \label{Binding_E}
@@
 and their binding energies. Collective processes and screening among 
-electrons are included. Since hafnium is a relativistic target, the 
-wave functions and binding energies must be obtained by solving the 
-many-electron Dirac Hamiltonian. Details of these calculations have been 
-published in \cite{mendez2019}, while Slater-type orbital expansions are 
-given in \cite{Hf_arxiv}.
+electrons are included. Since hafnium is a relativistic target, the wave 
+functions and binding energies must be obtained by solving the many-electron 
+Dirac Hamiltonian. Details of these calculations and a table of binding 
+energies have been published in Ref.~\cite{mendez2019}, while Slater-type 
+orbital expansions are given in Ref.~\cite{Hf_arxiv}.
 
-To assess the importance of the description of bound electrons, 
-Fig.~\ref{Binding_E} shows (a) non-relativistic~\cite{badnell97} and 
-our relativistic binding energies $E_{nl\pm}$, with $\pm=j\pm1/2$; and 
-(b) the relative errors with respect to the experimental data on solid 
-Hf~\cite{williams1995} from $1s$ to $4f_{\pm}$. 
-%\st{We notice that not only the most inner shells require relativistic calculations, but also the outer $5p$ and $4f$ shells.
-Furthermore, this figure shows very clearly the disability of non-relativistic calculations to describe the experimental data, which surprisingly worsens from the inner to the outer shells.}
-One would expect the relativistic contributions to only affect 
-significantly the inner shells. The relativistic binding energies of the 
-electrons belonging to these shells are 2.5\% from the experimental values,
-while the non-relativistic ones are within $\sim$10\%. Moreover, we notice 
-the relativistic corrections greatly influences the outer $5p$ and $4f$ 
-shells. This effect may be a consequence of the orthogonality with the 
-relativistic inner shells. Also, the sign of the binding energy deviations 
-is inverted near the FEG limit, i.e. the electrons go from being more 
-bounded --than the experimental values-- to being less bounded. 
-This change of sign can be attributed to the fact that the experimental 
+To assess the importance of a fully relativistic description of bound 
+electrons, Fig.~\ref{Binding_E}~(a) shows our binding energies, 
+$E_{nl\pm}$, with $\pm=j\pm1/2$; non-relativistic values~\cite{badnell97}; 
+and experimental data on solid-state Hf~\cite{williams1995}, which is 
+available only for $1s$ to $4f_{\pm}$ subshells, as expected. We notice that not only the most inner shells 
+require relativistic calculations, but also the outer $5p$ and $4f$ 
+shells. Furthermore, this figure shows very clearly the disability of 
+non-relativistic calculations to describe the experimental data, which 
+surprisingly worsens from the inner to the outer shells.
+
+From the comparison with the experimental values in Fig.~\ref{Binding_E}~(a),
+it can be noted that the sign of the binding energy deviations is 
+inverted for the outer $5s$ and $4f$ electrons, with the experimental 
+binding energies being less bounded than our theoretical ones. Small 
+differences for the outer shells are expected since the experimental
 values correspond to hafnium in solid-state, while our theoretical 
@@
 
-Our relativistic binding energies present spin-orbit split. However, in 
-total stopping power, where the initial state of the excited electron is 
-not measured, the quantum uncertainty in energy $\Delta E$ melts this 
-split. The criteria $\Delta E\Delta t\geq\hbar/2$ merges the energies 
+More detail of the theoretical binding energies is given in 
+Fig.~\ref{Binding_E}~(b), where relative errors with respect to the 
+experimental values are shown. This figure shows clearly that the 
+relativistic corrections are critical to describing the atomic structure 
+of hafnium, even for the outer shells. It turns out that the errors 
+committed in the non-relativistic calculations of the inner shell 
+orbitals propagate, through the Hartree-Fock approximation, to the outer 
+shells. The importance of fully relativistic calculations for the outer 
+shells has already been noted for Au, Pb, Bi, and W~\cite{mon09}.
+
+For the contribution of bound electrons to the total stopping cross 
+sections, the SLPA considers independent contributions of each subshell. 
+Our relativistic binding energies present spin-orbit split. However, 
+in total stopping power, where the initial state of the excited electron 
+is not measured, the quantum uncertainty in energy $\Delta E$ melts this 
+split. The criterion $\Delta E\Delta t\geq\hbar/2$ merges the energies 
 $E_{nl+}-E_{nl-}$ for sufficiently small values of $\Delta t$ (the 
@@
 we can expect all target electrons to respond together to the ion 
-passage~\cite{lindhard53,chu72}. Following previous work~\cite{mon09}, 
+passage~\cite{lindhard53,chu72}. Following previous works~\cite{mon09},
 the collisional time is estimated as $\Delta t\approx\langle r_i\rangle/v$, 
 with $\langle r_i\rangle$ and $v$ being the orbital mean radio and 
-impact velocity, respectively.
-
-%-----------------------------------------------------------------------
-\begin{figure*}[!t]
-\centering
-\includegraphics[width=13.cm]{Fig02.eps}
-\caption{Theoretical total stopping cross sections of protons in hafnium 
-adding FEG and bound $1s$-$4f$ contributions, as mentioned in the inset. 
-The vertical grey dashed-line indicates the energy above which plasmon 
-excitation is possible. The SLPA total stopping values for $1s$-$4f$ 
-bound electrons with and without $5p$-$4f$ screening are displayed with 
-solid and dotted lines, respectively.}
-\label{slpa4f}
-\end{figure*}
-%-----------------------------------------------------------------------
-
-The SLPA calculates the contribution of each subshell of bound electrons 
-to the total stopping cross sections. We found that for every sub-shell 
-of Hf, at the impact energy at which this sub-shell began to contribute, 
-the spin-orbit split was unresolved. Then, the $nl$-electrons should be 
-considered together, responding to the ion passage as a single gas of 
-electrons with density $\delta_{nl}(r)$ and a mean binding energy 
-$E_{nl}$. This feature is important within the SLPA calculations because 
+impact velocity, respectively. In the case of hafnium, we found that for 
+every sub-shell of electrons, the spin-orbit split is unresolved in the 
+energy region this sub-shell contributes. Therefore, the $nl$-electrons 
+should be considered together, responding to the ion passage as a single 
+gas of electrons with density $\delta_{nl}(r)$ and mean binding energy 
+$E_{nl}$. This feature is vital within the SLPA calculations because 
 it accounts for the screening among electrons of the \textit{same} 
@@
 $5p$ and $4f$ electrons of Hf are very close in energy 
-($\Delta E_{5p-4f} \approx 1$ a.u.~\cite{mendez2019}) and they react 
-together for impact energies $E>40$ keV (inter-shell screening).
+($\Delta E_{5p-4f} \approx 1$ a.u.~\cite{mendez2019}), and they react 
+together at impact energies $E>40$ keV (inter-shell screening). As 
+already mentioned, at higher energies, inter-shell screening is possible 
+for other subshells (i.e., $4p$-$4d$ for impact energies above 0.9 MeV), 
+but its weight in the total stopping is minor for deeper shells.
 
@@
 \centering
-\includegraphics[width=13.0cm]{Fig03.eps}
-\caption{Stopping power cross section of hafnium for protons. Symbols: 
-solid circles, present values; open circles, previous data~\cite{Sirotinin}. 
-Curves: Black solid-line, present full theoretical results with 
-$4f$-$5p$ screening; violet dashed-line, values from ICRU-49~\cite{ICRU49}, 
-and green dotted line, SRIM-2013~\cite{Ziegler01}.}
+\includegraphics[width=13.cm]{Fig02_new3.eps}
+\caption{(color online) Theoretical stopping cross sections of protons 
+in hafnium. 
+Blue dash-dotted-line, the non-perturbative SPCC for the FEG; 
+blue solid-line, ML results for the FEG (includes plasmon excitation); 
+red solid and dotted-lines, the SLPA results for bound electrons with 
+and without $5p$-$4f$ screening, respectively. 
+Black curves, total stopping adding the FEG and bound $1s$-$4f$ 
+contributions: 
+dash-dotted-line, SPCC (FEG) + SLPA (bound); 
+solid-line, ML (FEG) + SLPA (bound) with 4f-5p screening; 
+dotted-line, ML (FEG) + SLPA (bound) without 4f-5p screening.
+The vertical grey dashed-line indicates the energy of 37 keV
+above which plasmon excitation is possible.} 
+\label{slpa4f}
+\end{figure*}
+%-----------------------------------------------------------------------
+
+Finally, in all our calculations~\cite{mon17}, we assumed the projectile 
+to be proton and not neutral hydrogen. When an ion moves inside a metal, 
+the FEG screens the nucleus, so the binding energies will be smaller 
+than outside the metal, and this effect is more critical at low impact 
+velocities $v$. In the case of hydrogen, the difference is drastic, 
+i.e., for H inside Hf ($rs=2.07$), the $1s$-bound state is almost null at 
+$v<2$~\cite{suppression}. It is worth to mention that this assumption 
+agrees with Ziegler SRIM code~\cite{Ziegler01} but differs from CasP 
+code~\cite{Grande}, that predicts neutral hydrogen at very low velocities.
+
+%-----------------------------------------------------------------------
+\begin{figure*}[!t]
+\centering
+\includegraphics[width=13.0cm]{Fig03_new2.eps}
+\caption{(color online) Stopping power cross section of hafnium for 
+protons. Symbols: solid circles, present values; open circles, previous 
+data~\cite{Sirotinin}. Curves: 
+Black solid-line, present full theoretical results with $4f$-$5p$ screening; 
+pink dash-dot line, theoretical CasP5.2~\cite{Grande,casp52} values; 
+orange dash-double-dot line, theoretical DPASS~\cite{DPASS20} results; 
+green dotted-line, semi-empirical SRIM-2013~\cite{Ziegler01}; 
+violet dashed-line, ICRU49~\cite{ICRU49} tabulated values.}
 \label{F03}
@@
 In Fig.~\ref{slpa4f}, we display the present theoretical stopping cross
-section of Hf for protons with and without the $5p$-$4f$ screening. We 
-added the FEG and the bound electron contributions as explained above. 
-The minimum energy for plasmon excitation was estimated to $37$ keV. We 
-used the non-perturbative SPCC model for impact energies $E \leq 37$~keV, 
-and the perturbative ML calculation above this energy. Clearly, 
-considering $5p$-$4f$ electrons as a single group of 20 electrons with 
-screening among them gives lower stopping values than the addition of 
-the separate $5p$ and $4f$ contributions. Notice that this shell 
-correction can only be considered within a many electron model such as 
-the SLPA.
+section of Hf for protons using the relativistic wave functions 
+and binding energies, but with and without the $5p$-$4f$ screening. 
+We show the FEG and bound electron contributions separately and the 
+total stopping as the addition of both of them. The minimum energy for 
+plasmon excitation was estimated at approximately $37$ keV. We used the 
+non-perturbative SPCC model for impact energies $E \leq 37$~keV, and the 
+perturbative ML calculation above this value. Bound $1s$-$4f$ electrons 
+(relativistic wave functions and binding energies) are calculated with 
+the SLPA and shown separately in Fig.~\ref{slpa4f} with and without the 
+$4f$-$5p$ screening. Below $\sim 40$ keV, the difference between both 
+calculations is negligible. Considering $5p$-$4f$ electrons 
+as a single group of 20 electrons with screening among them gives lower 
+stopping values than the addition of the separate $5p$ and $4f$ 
+contributions. Notice that this shell correction can be considered 
+naturally within a many-electron model, such as the SLPA.
 
@@
 
-The present data and theoretical results are displayed in 
-Table~\ref{table01}, and Fig.~\ref{F03}. As can be seen in 
-Table~\ref{table01}, an overall relative uncertainty of around 5\% was 
-achieved for the experimental stopping power values, which are mainly 
-due to the uncertainty in the hafnium foil thickness.
+The present data are displayed in Table~\ref{table01}. As can be 
+observed, an overall relative uncertainty 
+of around 5\% was achieved for the experimental stopping power values, 
+which are mainly due to the uncertainty in the hafnium foil thickness.
 
@@
 
-In Fig.~\ref{F03}, we have good agreement between the present 
-theoretical results and the measurements displayed in Table~\ref{table01}. 
-Our theoretical approach also agrees with previous data by 
-Sirotinin~\cite{Sirotinin}, except for the lowest energy measurement at 
-80 keV. It is interesting that our full theoretical curve differs from 
-SRIM-2013 for impact energies below 100 keV. We obtain a stopping 
-maximum around $40 \times 10^{-15}$ eV cm$^2$/atom  at 65 keV. Instead, 
-following the up-to-now only set of data \cite{Sirotinin}, SRIM-2013 
-suggests a lower stopping maximum  at impact energy of 115 kev. It is 
-worth mentioning that similar theoretical results using the experimental 
-$r_S=2.07$ a.u. instead of the theoretical $r_S=2.14$ a.u. give the 
-stopping maximum at the same impact energy but $4\%$ higher. Future 
-experiments for impact energies below 100 keV would be important to 
-complete this study.
+Fig.~\ref{F03} synthesizes the results of the present work. The 
+agreement between the present theoretical results and the new 
+measurements displayed in Table~\ref{table01} is excellent. Present 
+measurements using the transmission method are in good agreement with 
+the previous data by Sirotinin~\cite{Sirotinin}, which were measured 
+in backscattering geometry. Our theoretical approach also agrees with 
+the data by Sirotinin~\cite{Sirotinin}, except for the lowest energy 
+measurement at 80 keV. We have also included in this figure the 
+theoretical curves from the CasP5.2 code by Grande and 
+Schiwietz~\cite{Grande,casp52} and from the DPASS code by Sigmund and 
+Schinner~\cite{DPASS20}, both available online. Furthermore, we 
+incorporated the semi-empirical results from SRIM-2013~\cite{Ziegler01} 
+and the ICRU49 tables~\cite{ICRU49}. Interestingly, our full theoretical 
+curve differs from SRIM-2013 for impact energies below 100 keV. We 
+obtain a stopping maximum of approximately 
+$40\times 10^{-15}$ eV cm$^2$/atom at 65 keV. Instead, following the 
+up-to-now only set of data \cite{Sirotinin}, SRIM-2013 suggests a lower 
+stopping maximum at an impact energy of 115 keV. The stopping maximum is 
+a susceptible region for any full theoretical description without 
+parameters, and this is quite visible in a linear-scale plot like 
+Fig.~\ref{F03}. However, the impact energy for the maximum seems to 
+agree between our curve and DPASS, although it is $10 \%$ below. 
+Instead, CasP a maximum value of the stopping power $10 \%$ above our 
+value and a completely different shape at lower energies. It is worth 
+mentioning that our model gives similar results using the experimental 
+value $r_S=2.07$ a.u. rather than the theoretical one $r_S=2.14$ a.u.,
+with the stopping maximum at the same impact energy but $4\%$ higher.
+
+Future experiments for impact energies below 100 keV would be essential 
+to complete this study on the stopping of Hf for protons. In the energy 
+region $80-500$ keV, only the measurements by Sirotinin's 
+group~\cite{Sirotinin} from 1984 are available. SRIM-2013 includes these 
+values in the code fitting, so it does not represent a test for this 
+data. Undoubtedly, new measurements are needed in three specific 
+energy regions: below 500 keV, around the stopping maximum 
+(i.e., $50-200$ keV), and at low impact energies.
 
@@
 theoretical framework that involved the relativistic wave functions and 
-binding energies of Hf, and considered four electrons per atom in the free 
-electron gas. The shell-wise local plasma approximation was employed to 
-describe the energy transferred to the bound $1s$-$4f$ electrons, and 
-two different models for the FEG: the screened potential with cusp 
-condition (SPCC model) for energies below that of the plasmon 
+binding energies of Hf and considered four electrons per atom in the 
+free electron gas. The shell-wise local plasma approximation was 
+employed to describe the energy transferred to the bound $1s$-$4f$ 
+electrons, and two different models for the FEG: the screened potential 
+with cusp condition (SPCC model) for energies below that of the plasmon 
 excitation, and the Mermin-Lindhard dielectric formalism, for energies 
 around the stopping maximum and above. Present theoretical stopping 
-cross sections cover an extensive energy range from 1 keV/amu-10 meV/amu.
+cross sections cover an extensive energy range from 1 keV/amu to 
+10 MeV/amu.
 
@@
 the attention that around the stopping maximum and at lower impact 
-energies the difference between our full-theoretical results and SRIM 
-is substantial. To the best of our knowledge, these are the first 
-theoretical calculations of stopping in Hf taking into account the 
-atomic relativistic effect. Future experiments for impact energies 
-below 100 keV would be important to complete this study.
+energies, the difference between our full-theoretical results and SRIM 
+is substantial. We compare our theoretical results with two other models 
+given by the DPASS and CasP5.2 codes. Differences can be noted at 
+intermediate to low impact energies, but they also support a stopping 
+maximum at lower energy than SRIM predictions.
+
+To the best of our knowledge, these are the first theoretical 
+calculations of stopping in Hf to take into account relativistic effects 
+in the atomic structure and screening among electrons in a consistent 
+way, from very low to high impact energies. Future experiments for 
+impact energies around the stopping maximum and in the low energy region 
+would be essential to have a better understanding of the stopping of 
+protons in hafnium.
 
@@
 
-\bibitem{ICRU49} 
-ICRU report 49, 
-Stopping Powers and Ranges for Protons and Alpha Particles, 
-International Commission on Radiation Units and Measurements, 1993.
+\bibitem{Grande} 
+G. Schiwietz, P. L. Grande, 
+Nucl. Instrum. Meth. B \textbf{175-177}, 125-131 (2001); 
+CasP code, available from www.casp-program.org
+
+\bibitem{casp52} 
+G. Schiwietz and P. L. Grande,
+Nucl. Instr. and Meth. B \textbf{273}, 1-5 (2012); 
+P.L. Grande and G. Schiwietz, 
+Phys. Rev. A \textbf{58}, 3796 (1998).
+
+\bibitem{DPASS20} 
+A. Schinner, P. Sigmund, 
+Nucl. Instrum. Meth. B \textbf{460}, 19 (2019); 
+P.Sigmund, A. Schinner, 
+Eur. Phys. J. D \textbf{12}, 425 (2000). 
+DPASS code, available from https://www.sdu.dk/en/DPASS/
 
 \bibitem{Ziegler01} 
-J.F. Ziegler, J.P. Biersack, SRIM2013, 
-Computer Program and Manual. Available from www.srim.org.
+J.F. Ziegler, J.P. Biersack, M. D. Ziegler, 
+\textit{SRIM, The Stopping and Range of Ions in Matter}, 
+(SRIM Co. Maryland, USA, 2008); 
+SRIM2013, Computer Program and Manual. Available from www.srim.org.
+
+\bibitem{ICRU49} 
+ICRU report 49, \textit{Stopping Powers and Ranges for Protons and Alpha Particles},
+International Commission on Radiation Units and Measurements (1993).
 
@@
 G. Williams in http://xdb.lbl.gov/Section1/Sec\_1-1.html
+
+\bibitem{mon09} 
+C.C. Montanari, D.M. Mitnik, C.D. Archubi, and J.E. Miraglia, 
+Phys. Rev. A \textbf{70}, 032903 (2009); 
+Phys. Rev. A \textbf{80}, 012901 (2009).
 

### 2020-06-02 sp_hafnium.tex
latest changes

@@
 decisive around the stopping maximum. Present theoretical and 
-experimental results are in excellent agreement in the energy region 
+experimental results are in very good agreement in the energy region 
 of the new measurements. However, our theoretical stopping 
@@
 shifted to lower energies than these previous predictions. Future 
-measurements below 100 keV would be necessary for a better understanding 
-of the stopping power around the maximum and below.
+measurements around the maximum and below would be necessary for a better understanding 
+of the stopping power of hafnium.
 \end{abstract}
@@
 often measured by using Rutherford Backscattering 
-Spectrometry~\cite{Alfassi01,Tesmer01}. This method relies heavily on 
-the determination of the stopping power of ion beams in the material of 
+Spectrometry~\cite{Alfassi01,Tesmer01}. This method relies on 
+the determination of both the scattering cross section and also the stopping power of ion beams in the material of 
 interest.
@@
 \caption{RBS spectrum for $E_{\mathrm{avg}}=921.1$ keV protons on 
-the hafnium sample, which is subsequently used to determine the energy 
-loss in the foil.}
+hafnium sample which is subsequently used to determine the energy loss 
+in the foil.}
 \label{F01}
@@
 X-Ray Diffraction) in Lisbon. This facility uses a 2.5 MV Van de Graaff 
-accelerator to deliver $^1$H$^+$ primary ion beams through a series of 
-electrostatic lenses and collimators onto a thin Au/SiO$_2$ sample, 
-which is used as a scattering center. This sample was placed in the 
-center of an RBS/C scattering chamber, where a high vacuum (pressure of 
-$\sim$10$^{-6}$ Torr) was maintained during the measurements. The beam 
-current on the sample was kept at around 5.0 nA to attain sufficient 
-statistics in each particle spectrum. By using a beam spot of about 
-1.0 mm in diameter, a solid angle of 11.4 msr was attained. The overall 
-energy resolution (FWHM) of the detection system was about 15 keV 
-relative to 5.486 MeV alpha particles from a $^{241}$Am source.
+accelerator to deliver $^1$H$^+$ primary ion beams with a precision better 
+than $\pm 0.5$ keV through a series of electrostatic lenses and collimators 
+onto a thin Au/SiO$_2$ sample, which is used as a scattering center. The Hf 
+foil was mounted on a movable target holder and placed inside a RBS/C scattering 
+chamber to allow energy measurements of the direct beam and the beam transmitted 
+through the sample without breaking the high vacuum ($\sim$10$^{-6}$ Torr) inside 
+the scattering chamber. The beam current on the sample was kept at around 5.0 nA 
+to attain sufficient statistics in each particle spectrum. By using a beam spot 
+of about 1.0 mm in diameter, a solid angle of 11.4 msr was attained. The overall 
+energy resolution (FWHM) of the detection system was about 15 keV relative to 5.486 
+MeV alpha particles from a $^{241}$Am source.
 
@@
 thickness of 1.0 $\mu$m and 99.95\% purity, which was supplied by Lebow 
-Company~\cite{Lebow}. However, a more precise thickness value was 
-achieved by measuring the energy loss of alpha particles coming from a 
-calibrated ($^{239}$Pu, $^{241}$Am, $^{244}$Cm) source. From the alpha 
+Company~\cite{Lebow}. A more precise thickness value was achieved by measuring 
+the energy loss of alpha particles coming from a calibrated 
+($^{239}$Pu, $^{241}$Am, $^{244}$Cm) source. From the alpha 
 spectra with and without the Hf foil interposed, the characteristic 
@@
 ($4.13 \pm 0.21$)$\times 10^{19}$ at/cm$^2$, which corresponds to a 
-thickness of $0.920\pm0.046 \mu$m.
-
+thickness of $0.920\pm0.046 \mu$m. Target non-uniformity was investigated through 
+systematic measurements (at five different points over the sample area) of the energy 
+loss of alpha particles from the same radioactive source. The uncertainties originating
+from the non-uniformity of the sample was $\sim 2.5$\%. However, the primary source of
+uncertainty related to target thickness comes from estimates in the SRIM database for 
+alphas on hafnium ($\sim$ 4\%). Additionally, we consider estimates coming from surface 
+roughness ($\sim$ 1\%) and possible impurities ($\sim$ 1\%) in the foil; and finally, 
+statistical uncertainty ($\sim$ 0.6\%) related to the gaussian fits used to determine 
+the energy loss of alphas through the Hf target.
 %-----------------------------------------------------------------------
@@
 correction to the mean energies $E_{\mathrm{avg}}$ was applied in order 
-to account for the non-linear dependence on ion energy of stopping 
-powers \cite{Chilton,Rajatora}.
+to account for the non-linear dependence on ion energy of stopping powers 
+\cite{Chilton,Rajatora}. The uncertainty ($\sim$ 0.7\%) in the measured energy 
+loss $\Delta E$ of protons in the hafnium sample is mainly related to the 
+statistical uncertainty found in the gaussian fits mentioned above. If this value 
+is combined with the $\sim$ 4.9\% uncertainty in target thickness, then a 
+$\sim$ 5.0\% uncertainty in the measured cross section is obtained.
 
@@
 Experimental measurements for solids~\cite{williams1995} are depicted 
-with open circles. (b) Corresponding relative errors with respect to 
+with open circles. (b) Corresponding relative errors respect to 
 experimental data.}
@@
 
-More detail of the theoretical binding energies is given in 
+More detail about the theoretical binding energies is given in 
 Fig.~\ref{Binding_E}~(b), where relative errors with respect to the 
 experimental values are shown. This figure shows clearly that the 
-relativistic corrections are critical to describing the atomic structure 
+relativistic corrections are critical to describe the atomic structure 
 of hafnium, even for the outer shells. It turns out that the errors 
@@
 orbitals propagate, through the Hartree-Fock approximation, to the outer 
-shells. The importance of fully relativistic calculations for the outer 
+shells. 
+The non-relativistic $4f$ binding energy is four times the experimental one. Such an incorrect value leads to the underestimation of the $4f$-ionization and shifts  the threshold to higher energies.  
+The importance of fully relativistic calculations for the outer 
 shells has already been noted for Au, Pb, Bi, and W~\cite{mon09}.
@@
 together at impact energies $E>40$ keV (inter-shell screening). As 
-already mentioned, at higher energies, inter-shell screening is possible 
-for other subshells (i.e., $4p$-$4d$ for impact energies above 0.9 MeV), 
-but its weight in the total stopping is minor for deeper shells.
+already mentioned, at higher energies, inter-shell screening is also possible 
+for deeper subshells %(i.e., $4p$-$4d$ for impact energies above 0.9 MeV), 
+but their weight in the total stopping is minor.
 
@@
 stopping values than the addition of the separate $5p$ and $4f$ 
-contributions. Notice that this shell correction can be considered 
-naturally within a many-electron model, such as the SLPA.
+contributions. Notice that this shell correction can only be considered self-consistently within a many-electron model, such as the SLPA.
 
@@
 
-The present data are displayed in Table~\ref{table01}. As can be 
-observed, an overall relative uncertainty 
-of around 5\% was achieved for the experimental stopping power values, 
-which are mainly due to the uncertainty in the hafnium foil thickness.
+The present data are displayed in Table~\ref{table01}. As can be observed, 
+an overall relative uncertainty of around 5\% was achieved for the experimental 
+stopping power values, which are mainly due to the uncertainty in the hafnium foil 
+thickness.
 
@@
 up-to-now only set of data \cite{Sirotinin}, SRIM-2013 suggests a lower 
-stopping maximum at an impact energy of 115 keV. The stopping maximum is 
-a susceptible region for any full theoretical description without 
-parameters, and this is quite visible in a linear-scale plot like 
+stopping maximum at an impact energy of 115 keV.
+
+The stopping maximum is 
+a very sensitive region for any full theoretical description, and this is quite visible in a linear-scale plot like 
 Fig.~\ref{F03}. However, the impact energy for the maximum seems to 
 agree between our curve and DPASS, although it is $10 \%$ below. 
-Instead, CasP a maximum value of the stopping power $10 \%$ above our 
-value and a completely different shape at lower energies. It is worth 
+Instead, CasP maximum is $10 \%$ above ours, but 
+has a completely different shape at lower energies. It is worth 
 mentioning that our model gives similar results using the experimental 
@@
 with the stopping maximum at the same impact energy but $4\%$ higher.
-
-Future experiments for impact energies below 100 keV would be essential 
-to complete this study on the stopping of Hf for protons. In the energy 
-region $80-500$ keV, only the measurements by Sirotinin's 
-group~\cite{Sirotinin} from 1984 are available. SRIM-2013 includes these 
-values in the code fitting, so it does not represent a test for this 
-data. Undoubtedly, new measurements are needed in three specific 
-energy regions: below 500 keV, around the stopping maximum 
-(i.e., $50-200$ keV), and at low impact energies.
+Future experiments would be important for a more complete understanding of this case, mainly for proton energies around the stopping maximum (i.e.
+$30-300$ keV) and also below $25$ keV, in the region where a linear dependence with the velocity is expected.
 
@@
 To the best of our knowledge, these are the first theoretical 
-calculations of stopping in Hf to take into account relativistic effects 
-in the atomic structure and screening among electrons in a consistent 
-way, from very low to high impact energies. Future experiments for 
+calculations of stopping in Hf that cover from very low to high impact energies, taking into account relativistic effects in the atomic structure and screening among electrons in a consistent 
+way. Future experiments for 
 impact energies around the stopping maximum and in the low energy region 
